---
title: "Land Use Efficiency Metrics"
format: html
editor: visual
---

# 1 - Setup: paths and parameters

```{r, setup, include=FALSE}
options(scipen=999)

pacman::p_load(
  sf,
  data.table,
  tidycensus,
  tidyverse,
  openxlsx,
  tidygeocoder,
  terra,
  reticulate
)

# devtools::install_github('MAPC/mapcdatakeys')
library(mapcdatakeys)

reticulate::use_python("C:/Program Files/ArcGIS/Pro/bin/Python/envs/arcgispro-py3/python.exe")

# Paths
if (grepl('VM',Sys.info()[['nodename']])){
  base <- '//data-sync/Public/DataServices'
} else {
  base <- 'S:/Network Shares/K Drive/DataServices'
}


data_path <- file.path(base,'Projects/Current_Projects/Environment/EEA_MLCA/Data')
task_path = paste0(data_path,'/Task_3/')
deliv_path = paste0(task_path, 'Deliverables')
sp_path <- paste0(task_path,'Spatial/')
task_gdb = paste0(sp_path, 'LandUse_Efficiency_Metrics.gdb')
task_gdb2 = paste0(sp_path, 'LUE_Metrics.gdb')
task_gpkg = paste0(sp_path, 'LandUse_Efficiency_Metrics.gpkg')

parcel_root <- paste0(base,'/Datasets/Parcel_DB/Data/parcel_timeseries')
assess_path <- file.path(parcel_root,'_Assess_CSVs')
poly_gdb <- paste0(parcel_root,"/_Merged_polys/All_parcel_polys.gdb")
comp_gdb <- paste0(parcel_root,"/_Merged_polys/Assess_joined_parcel_polys.gdb")
delta_gdb <- paste0(parcel_root,"/_Merged_polys/Delta_parcel_polys.gdb")
year_gpkg <- paste0(parcel_root,"/_Merged_polys/parcels_by_year.gpkg")
muni_gpkg <- paste0(parcel_root,"/_Merged_polys/parcels_by_muni_timeseries.gpkg")


munis <- mapcdatakeys::all_muni_data_keys |> 
  select(muni_id,muni_name,cosub_5y14,cosub_5y23,cmsbt08,mpo) |> 
  mutate(cosub_5y14 = as.character(cosub_5y14),
         cosub_5y23 = as.character(cosub_5y23))

blk_2020 <- geog_xw_2020 |>
  select(bl20_id, muni_id) |> 
  left_join(munis, by = 'muni_id') |>
  select(bl20_id, muni_id, muni_name, cmsbt08, mpo) |>
  setDT()

knitr::opts_knit$set(root.dir = data_path)
```

# 2 - ACS data 2014

```{r}

munis <- mapcdatakeys::all_muni_data_keys |> 
  select(muni_id,muni_name,cosub_5y14,cosub_5y23,cmsbt08,mpo) |> 
  mutate(cosub_5y14 = as.character(cosub_5y14),
         cosub_5y23 = as.character(cosub_5y23))

# NEED TO SWITCH TO 2020 TRACT + MUNI GEOGRAPHIES for the 2014 data
# AND SWITCH BACK THE 2023 DATA to 2020 tracts as well
ct_xw <- fread(paste0(data_path, '/nhgis_bg2010_tr2020_25.csv')) |>
  mutate(bg2010ge = as.character(bg2010ge))


vars14 <- load_variables('acs5', year = 2014) |>
  setDT()

uv14 <- vars14[concept == 'UNITS IN STRUCTURE', name]
hv14 <- vars14[concept == 'HOUSEHOLD TYPE BY TENURE' &
                 label == 'Estimate!!Total', name]
pv14 <- vars14[concept == 'TOTAL POPULATION', name]

vs14 <- c(pv14, hv14, uv14)

# Tract-level
acs14.tract <- get_acs(
  year = 2014,
  state = 'MA',
  survey = 'acs5',
  variables = vs14,
  geography = 'block group',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'bg2010ge',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    bg2010ge,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  left_join(ct_xw, by = 'bg2010ge') |>
  mutate(
    total_pop = (total_pop * wt_pop),
    total_households = (total_households * wt_hh),
    total_units = (total_units * wt_hu),
    units_1 = (units_1 * wt_hu),
    units_2 = (units_2 * wt_hu),
    units_3_4 = (units_3_4 * wt_hu),
    units_5_9 = (units_5_9 * wt_hu),
    units_10p = (units_10p * wt_hu),
  ) |>
  rename(ct20_id = tr2020ge) |>
  group_by(ct20_id) |>
  summarize(
    total_pop = round(sum(total_pop)),
    total_households = round(sum(total_households)),
    total_units = round(sum(total_units)),
    units_1 = round(sum(units_1)),
    units_2 = round(sum(units_2)),
    units_3_4 = round(sum(units_3_4)),
    units_5_9 = round(sum(units_5_9)),
    units_10p = round(sum(units_10p))
  ) |>
  setorder(ct20_id) |>
  mutate(ct20_id = as.character(ct20_id)) |> 
  setDT()
setnames(acs14.tract, c('ct20_id',paste0(names(acs14.tract)[2:9],'_2014')))

# Muni-level
acs14.muni <- get_acs(
  year = 2014,
  state = 'MA',
  survey = 'acs5',
  variables = vs14,
  geography = 'county subdivision',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'cosub_5y14',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    cosub_5y14,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  left_join(munis, by = 'cosub_5y14') |>
  relocate(c(muni_id, muni_name, cmsbt08, mpo)) |>
  select(-c(cosub_5y14, cosub_5y23)) |>
  filter(!is.na(muni_id)) |>
  setorder(muni_id) |>
  setDT()

acs14.cmsbt <- acs14.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    total_pop_2014 = round(sum(total_pop)),
    total_households_2014 = round(sum(total_households)),
    total_units_2014 = round(sum(total_units)),
    units_1_2014 = round(sum(units_1)),
    units_2_2014 = round(sum(units_2)),
    units_3_4_2014 = round(sum(units_3_4)),
    units_5_9_2014 = round(sum(units_5_9)),
    units_10p_2014 = round(sum(units_10p))
  ) 
  
acs14.sw <- acs14.muni |> 
  summarize(
    total_pop_2014 = round(sum(total_pop)),
    total_households_2014 = round(sum(total_households)),
    total_units_2014 = round(sum(total_units)),
    units_1_2014 = round(sum(units_1)),
    units_2_2014 = round(sum(units_2)),
    units_3_4_2014 = round(sum(units_3_4)),
    units_5_9_2014 = round(sum(units_5_9)),
    units_10p_2014 = round(sum(units_10p))
  ) |> 
  mutate(mpo='STATEWIDE')

acs14.mpo <- acs14.muni |> 
  group_by(mpo) |> 
  summarize(
    total_pop_2014 = round(sum(total_pop)),
    total_households_2014 = round(sum(total_households)),
    total_units_2014 = round(sum(total_units)),
    units_1_2014 = round(sum(units_1)),
    units_2_2014 = round(sum(units_2)),
    units_3_4_2014 = round(sum(units_3_4)),
    units_5_9_2014 = round(sum(units_5_9)),
    units_10p_2014 = round(sum(units_10p))
  ) |> 
  rbind(acs14.sw)

setnames(acs14.muni, c(names(acs14.muni)[1:4],paste0(names(acs14.muni)[5:12],'_2014')))

```

# 3 - ACS data 2023

```{r}


vars23 <- load_variables('acs5', year = 2023) |>
  setDT()

uv23 <- vars23[concept == 'Units in Structure', name]
hv23 <- vars23[concept == 'Households by Type' &
                 label == 'Estimate!!Total:', name]
pv23 <- vars23[concept == 'Total Population', name]

vs23 <- c(pv23, hv23, uv23)

# Tract-level
acs23.tract <- get_acs(
  year = 2023,
  state = 'MA',
  survey = 'acs5',
  variables = vs23,
  geography = 'tract',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'ct20_id',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    ct20_id,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  setorder(ct20_id) |>
  mutate(ct20_id = as.character(ct20_id)) |> 
  setDT()
setnames(acs23.tract, c('ct20_id',paste0(names(acs23.tract)[2:9],'_2023')))

# Muni-level
acs23.muni <- get_acs(
  year = 2023,
  state = 'MA',
  survey = 'acs5',
  variables = vs23,
  geography = 'county subdivision',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'cosub_5y23',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    cosub_5y23,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  left_join(munis, by = 'cosub_5y23') |>
  relocate(c(muni_id, muni_name, cmsbt08, mpo)) |>
  select(-c(cosub_5y14, cosub_5y23)) |>
  filter(!is.na(muni_id)) |>
  setorder(muni_id) |>
  setDT()

acs23.cmsbt <- acs23.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    total_pop_2023 = round(sum(total_pop)),
    total_households_2023 = round(sum(total_households)),
    total_units_2023 = round(sum(total_units)),
    units_1_2023 = round(sum(units_1)),
    units_2_2023 = round(sum(units_2)),
    units_3_4_2023 = round(sum(units_3_4)),
    units_5_9_2023 = round(sum(units_5_9)),
    units_10p_2023 = round(sum(units_10p))
  ) 
  
acs23.sw <- acs23.muni |> 
  summarize(
    total_pop_2023 = round(sum(total_pop)),
    total_households_2023 = round(sum(total_households)),
    total_units_2023 = round(sum(total_units)),
    units_1_2023 = round(sum(units_1)),
    units_2_2023 = round(sum(units_2)),
    units_3_4_2023 = round(sum(units_3_4)),
    units_5_9_2023 = round(sum(units_5_9)),
    units_10p_2023 = round(sum(units_10p))
  ) |> 
  mutate(mpo='STATEWIDE')

acs23.mpo <- acs23.muni |> 
  group_by(mpo) |> 
  summarize(
    total_pop_2023 = round(sum(total_pop)),
    total_households_2023 = round(sum(total_households)),
    total_units_2023 = round(sum(total_units)),
    units_1_2023 = round(sum(units_1)),
    units_2_2023 = round(sum(units_2)),
    units_3_4_2023 = round(sum(units_3_4)),
    units_5_9_2023 = round(sum(units_5_9)),
    units_10p_2023 = round(sum(units_10p))
  ) |> 
  rbind(acs23.sw)

setnames(acs23.muni, c(names(acs23.muni)[1:4],paste0(names(acs23.muni)[5:12],'_2023')))

```

# 4 - LEHD data 2014 + 2023

```{r}

emp14 <- lehdr::grab_lodes(state = 'ma',
                           year = 2014,
                           lodes_type = 'wac') |>
  mutate(
    emp_2014_sect_1 = CNS04,
    emp_2014_sect_2 = CNS15 + CNS16,
    emp_2014_sect_3 = CNS10 + CNS11,
    emp_2014_sect_4 = CNS20,
    emp_2014_sect_5 = CNS09,
    emp_2014_sect_6 = CNS07 + CNS17 + CNS18,
    emp_2014_sect_7 = CNS05,
    emp_2014_sect_8 = CNS01 + CNS02 + CNS19,
    emp_2014_sect_9 = CNS12 + CNS13 + CNS14,
    emp_2014_sect_10 = CNS03 + CNS06 + CNS08
  ) |>
  select(
    w_geocode,
    C000,
    emp_2014_sect_1,
    emp_2014_sect_2,
    emp_2014_sect_3,
    emp_2014_sect_4,
    emp_2014_sect_5,
    emp_2014_sect_6,
    emp_2014_sect_7,
    emp_2014_sect_8,
    emp_2014_sect_9,
    emp_2014_sect_10
  ) |>
  rename(bl20_id = w_geocode, emp_2014 = C000) |>
  mutate(bl20_id = as.character(bl20_id)) |>
  setDT()

emp23 <- lehdr::grab_lodes(state = 'ma',
                           year = 2023,
                           lodes_type = 'wac') |>
  mutate(
    emp_2023_sect_1 = CNS04,
    emp_2023_sect_2 = CNS15 + CNS16,
    emp_2023_sect_3 = CNS10 + CNS11,
    emp_2023_sect_4 = CNS20,
    emp_2023_sect_5 = CNS09,
    emp_2023_sect_6 = CNS07 + CNS17 + CNS18,
    emp_2023_sect_7 = CNS05,
    emp_2023_sect_8 = CNS01 + CNS02 + CNS19,
    emp_2023_sect_9 = CNS12 + CNS13 + CNS14,
    emp_2023_sect_10 = CNS03 + CNS06 + CNS08
  ) |>
  select(
    w_geocode,
    C000,
    emp_2023_sect_1,
    emp_2023_sect_2,
    emp_2023_sect_3,
    emp_2023_sect_4,
    emp_2023_sect_5,
    emp_2023_sect_6,
    emp_2023_sect_7,
    emp_2023_sect_8,
    emp_2023_sect_9,
    emp_2023_sect_10
  ) |>
  rename(bl20_id = w_geocode, emp_2023 = C000) |>
  mutate(bl20_id = as.character(bl20_id)) |> 
  setDT()


emp.blk <- blk_2020 |> 
  mutate(bl20_id = as.character(bl20_id)) |> 
  left_join(emp14, by = 'bl20_id') |> 
  left_join(emp23, by='bl20_id') |> 
  setDT()

emp.ct <- emp.blk |> 
  mutate(ct20_id = substr(bl20_id,1,11)) |> 
  group_by(ct20_id) |> 
  summarize(
    emp2014 = sum(emp_2014, na.rm=T),
    emp2023 = sum(emp_2023, na.rm=T),
    emp_2014_sect_1 = sum(emp_2014_sect_1, na.rm=T),
    emp_2014_sect_2 = sum(emp_2014_sect_2, na.rm=T),
    emp_2014_sect_3 = sum(emp_2014_sect_3, na.rm=T),
    emp_2014_sect_4 = sum(emp_2014_sect_4, na.rm=T),
    emp_2014_sect_5 = sum(emp_2014_sect_5, na.rm=T),
    emp_2014_sect_6 = sum(emp_2014_sect_6, na.rm=T),
    emp_2014_sect_7 = sum(emp_2014_sect_7, na.rm=T),
    emp_2014_sect_8 = sum(emp_2014_sect_8, na.rm=T),
    emp_2014_sect_9 = sum(emp_2014_sect_9, na.rm=T),
    emp_2014_sect_10 = sum(emp_2014_sect_10, na.rm=T),
    emp_2023_sect_1 = sum(emp_2023_sect_1, na.rm=T),
    emp_2023_sect_2 = sum(emp_2023_sect_2, na.rm=T),
    emp_2023_sect_3 = sum(emp_2023_sect_3, na.rm=T),
    emp_2023_sect_4 = sum(emp_2023_sect_4, na.rm=T),
    emp_2023_sect_5 = sum(emp_2023_sect_5, na.rm=T),
    emp_2023_sect_6 = sum(emp_2023_sect_6, na.rm=T),
    emp_2023_sect_7 = sum(emp_2023_sect_7, na.rm=T),
    emp_2023_sect_8 = sum(emp_2023_sect_8, na.rm=T),
    emp_2023_sect_9 = sum(emp_2023_sect_9, na.rm=T),
    emp_2023_sect_10 = sum(emp_2023_sect_10, na.rm=T)
  ) |> 
  setDT()

emp.muni <- emp.blk |> 
  group_by(muni_id,muni_name) |> 
  summarize(
    emp2014 = sum(emp_2014, na.rm=T),
    emp2023 = sum(emp_2023, na.rm=T),
    emp_2014_sect_1 = sum(emp_2014_sect_1, na.rm=T),
    emp_2014_sect_2 = sum(emp_2014_sect_2, na.rm=T),
    emp_2014_sect_3 = sum(emp_2014_sect_3, na.rm=T),
    emp_2014_sect_4 = sum(emp_2014_sect_4, na.rm=T),
    emp_2014_sect_5 = sum(emp_2014_sect_5, na.rm=T),
    emp_2014_sect_6 = sum(emp_2014_sect_6, na.rm=T),
    emp_2014_sect_7 = sum(emp_2014_sect_7, na.rm=T),
    emp_2014_sect_8 = sum(emp_2014_sect_8, na.rm=T),
    emp_2014_sect_9 = sum(emp_2014_sect_9, na.rm=T),
    emp_2014_sect_10 = sum(emp_2014_sect_10, na.rm=T),
    emp_2023_sect_1 = sum(emp_2023_sect_1, na.rm=T),
    emp_2023_sect_2 = sum(emp_2023_sect_2, na.rm=T),
    emp_2023_sect_3 = sum(emp_2023_sect_3, na.rm=T),
    emp_2023_sect_4 = sum(emp_2023_sect_4, na.rm=T),
    emp_2023_sect_5 = sum(emp_2023_sect_5, na.rm=T),
    emp_2023_sect_6 = sum(emp_2023_sect_6, na.rm=T),
    emp_2023_sect_7 = sum(emp_2023_sect_7, na.rm=T),
    emp_2023_sect_8 = sum(emp_2023_sect_8, na.rm=T),
    emp_2023_sect_9 = sum(emp_2023_sect_9, na.rm=T),
    emp_2023_sect_10 = sum(emp_2023_sect_10, na.rm=T)
  ) |> 
  setDT()

emp.cmsbt <- emp.blk |> 
  group_by(cmsbt08) |> 
  summarize(
    emp2014 = sum(emp_2014, na.rm=T),
    emp2023 = sum(emp_2023, na.rm=T),
    emp_2014_sect_1 = sum(emp_2014_sect_1, na.rm=T),
    emp_2014_sect_2 = sum(emp_2014_sect_2, na.rm=T),
    emp_2014_sect_3 = sum(emp_2014_sect_3, na.rm=T),
    emp_2014_sect_4 = sum(emp_2014_sect_4, na.rm=T),
    emp_2014_sect_5 = sum(emp_2014_sect_5, na.rm=T),
    emp_2014_sect_6 = sum(emp_2014_sect_6, na.rm=T),
    emp_2014_sect_7 = sum(emp_2014_sect_7, na.rm=T),
    emp_2014_sect_8 = sum(emp_2014_sect_8, na.rm=T),
    emp_2014_sect_9 = sum(emp_2014_sect_9, na.rm=T),
    emp_2014_sect_10 = sum(emp_2014_sect_10, na.rm=T),
    emp_2023_sect_1 = sum(emp_2023_sect_1, na.rm=T),
    emp_2023_sect_2 = sum(emp_2023_sect_2, na.rm=T),
    emp_2023_sect_3 = sum(emp_2023_sect_3, na.rm=T),
    emp_2023_sect_4 = sum(emp_2023_sect_4, na.rm=T),
    emp_2023_sect_5 = sum(emp_2023_sect_5, na.rm=T),
    emp_2023_sect_6 = sum(emp_2023_sect_6, na.rm=T),
    emp_2023_sect_7 = sum(emp_2023_sect_7, na.rm=T),
    emp_2023_sect_8 = sum(emp_2023_sect_8, na.rm=T),
    emp_2023_sect_9 = sum(emp_2023_sect_9, na.rm=T),
    emp_2023_sect_10 = sum(emp_2023_sect_10, na.rm=T)
  ) |> 
  setDT()

emp.sw <- emp.blk |> 
  summarize(
    emp2014 = sum(emp_2014, na.rm=T),
    emp2023 = sum(emp_2023, na.rm=T),
    emp_2014_sect_1 = sum(emp_2014_sect_1, na.rm=T),
    emp_2014_sect_2 = sum(emp_2014_sect_2, na.rm=T),
    emp_2014_sect_3 = sum(emp_2014_sect_3, na.rm=T),
    emp_2014_sect_4 = sum(emp_2014_sect_4, na.rm=T),
    emp_2014_sect_5 = sum(emp_2014_sect_5, na.rm=T),
    emp_2014_sect_6 = sum(emp_2014_sect_6, na.rm=T),
    emp_2014_sect_7 = sum(emp_2014_sect_7, na.rm=T),
    emp_2014_sect_8 = sum(emp_2014_sect_8, na.rm=T),
    emp_2014_sect_9 = sum(emp_2014_sect_9, na.rm=T),
    emp_2014_sect_10 = sum(emp_2014_sect_10, na.rm=T),
    emp_2023_sect_1 = sum(emp_2023_sect_1, na.rm=T),
    emp_2023_sect_2 = sum(emp_2023_sect_2, na.rm=T),
    emp_2023_sect_3 = sum(emp_2023_sect_3, na.rm=T),
    emp_2023_sect_4 = sum(emp_2023_sect_4, na.rm=T),
    emp_2023_sect_5 = sum(emp_2023_sect_5, na.rm=T),
    emp_2023_sect_6 = sum(emp_2023_sect_6, na.rm=T),
    emp_2023_sect_7 = sum(emp_2023_sect_7, na.rm=T),
    emp_2023_sect_8 = sum(emp_2023_sect_8, na.rm=T),
    emp_2023_sect_9 = sum(emp_2023_sect_9, na.rm=T),
    emp_2023_sect_10 = sum(emp_2023_sect_10, na.rm=T)
  ) |> 
  mutate(mpo = 'STATEWIDE') |> 
  setDT()

emp.mpo <- emp.blk |> 
  group_by(mpo) |> 
  summarize(
    emp2014 = sum(emp_2014, na.rm=T),
    emp2023 = sum(emp_2023, na.rm=T),
    emp_2014_sect_1 = sum(emp_2014_sect_1, na.rm=T),
    emp_2014_sect_2 = sum(emp_2014_sect_2, na.rm=T),
    emp_2014_sect_3 = sum(emp_2014_sect_3, na.rm=T),
    emp_2014_sect_4 = sum(emp_2014_sect_4, na.rm=T),
    emp_2014_sect_5 = sum(emp_2014_sect_5, na.rm=T),
    emp_2014_sect_6 = sum(emp_2014_sect_6, na.rm=T),
    emp_2014_sect_7 = sum(emp_2014_sect_7, na.rm=T),
    emp_2014_sect_8 = sum(emp_2014_sect_8, na.rm=T),
    emp_2014_sect_9 = sum(emp_2014_sect_9, na.rm=T),
    emp_2014_sect_10 = sum(emp_2014_sect_10, na.rm=T),
    emp_2023_sect_1 = sum(emp_2023_sect_1, na.rm=T),
    emp_2023_sect_2 = sum(emp_2023_sect_2, na.rm=T),
    emp_2023_sect_3 = sum(emp_2023_sect_3, na.rm=T),
    emp_2023_sect_4 = sum(emp_2023_sect_4, na.rm=T),
    emp_2023_sect_5 = sum(emp_2023_sect_5, na.rm=T),
    emp_2023_sect_6 = sum(emp_2023_sect_6, na.rm=T),
    emp_2023_sect_7 = sum(emp_2023_sect_7, na.rm=T),
    emp_2023_sect_8 = sum(emp_2023_sect_8, na.rm=T),
    emp_2023_sect_9 = sum(emp_2023_sect_9, na.rm=T),
    emp_2023_sect_10 = sum(emp_2023_sect_10, na.rm=T)
  ) |> 
  rbind(emp.sw) |> 
  setDT()




```

# 5 - NLCD Developed Area calculation

```{r}
msf <- muni_sf(2020) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,mpo,cmsbt08), by = 'muni_id')
ct20 <- tract_sf(2020)

dev2014 <- st_read(dsn = task_gdb, layer='nlcd2014_developed_area') |> 
  st_transform(crs = st_crs(msf))

dev14_ct <- st_intersection(dev2014,ct20) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev14_ct, dsn = task_gpkg, layer = 'nlcd2014_developed_tract_inter', append=F)

dev14_ct <- dev14_ct |> 
  group_by(ct20_id) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev14_ct, paste0(sp_path, 'nlcd2014_dev_area_tract.rds'))

dev14_ms <- st_intersection(dev2014,msf) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev14_ms, dsn = task_gpkg, layer = 'nlcd2014_developed_muni_inter', append=F)

dev14_muni <- dev14_ms |> 
  group_by(muni_id, muni_name) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev14_muni, paste0(sp_path, 'nlcd2014_dev_area_muni.rds'))

dev14_cmsbt <- dev14_ms |> 
  group_by(cmsbt08) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev14_cmsbt, paste0(sp_path, 'nlcd2014_dev_area_commsubtype.rds'))

dev14_sw <- dev14_ms |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  mutate(mpo='STATEWIDE')

dev14_mpo <- dev14_ms |> 
  group_by(mpo) |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  rbind(dev14_sw)
saveRDS(dev14_mpo, paste0(sp_path, 'nlcd2014_dev_area_mpo.rds'))

# 2023 NLCD

dev2023 <- st_read(dsn = task_gdb, layer='nlcd2023_developed_area') |> 
    st_transform(crs = st_crs(msf))

dev23_ct <- st_intersection(dev2023,ct20) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev23_ct, dsn = task_gpkg, layer = 'nlcd2023_developed_tract_inter', append=F)

dev23_ct <- dev23_ct |> 
  group_by(ct20_id) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev23_ct, paste0(sp_path, 'nlcd2023_dev_area_tract.rds'))

dev23_ms <- st_intersection(dev2023,msf) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev23_ms, dsn = task_gpkg, layer = 'nlcd2023_developed_muni_inter', append=F)

dev23_muni <- dev23_ms |> 
  group_by(muni_id, muni_name) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev23_muni, paste0(sp_path, 'nlcd2023_dev_area_muni.rds'))

dev23_cmsbt <- dev23_ms |> 
  group_by(cmsbt08) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev23_cmsbt, paste0(sp_path, 'nlcd2023_dev_area_commsubtype.rds'))

dev23_sw <- dev23_ms |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  mutate(mpo='STATEWIDE')

dev23_mpo <- dev23_ms |> 
  group_by(mpo) |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  rbind(dev23_sw)
saveRDS(dev23_mpo, paste0(sp_path, 'nlcd2023_dev_area_mpo.rds'))


```

# 6 - Parcels land use codes and area

## 6a - Parcels combined for \<=2014

```{r}

alist <- fread(paste0(sp_path,'parcel_vintage_years_by_muni.csv'))

msf <- mapcdatakeys::muni_sf(2020) |> 
  select(muni_name,geometry)

m14 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2014')

na14 <- m14 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2014) |> 
  select(-muni_name)

m14 <- m14 |> 
  filter(!is.na(CITY)) |> 
  rbind(na14) |> 
  mutate(SUMYEAR=2014)

city2014 <- unique(m14$CITY)

m13 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2013') 
na13 <- m13 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2013) |> 
  select(-muni_name)

m13 <- m13 |> 
  filter(!is.na(CITY)) |> 
  rbind(na13)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m13)
city2014 <- unique(m14$CITY)

m12 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2012') 
na12 <- m12 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2012) |> 
  select(-muni_name)

m12 <- m12 |> 
  filter(!is.na(CITY)) |> 
  rbind(na12)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m12)
city2014 <- unique(m14$CITY)

m11 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2011') 
na11 <- m11 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2011) |> 
  select(-muni_name)

m11 <- m11 |> 
  filter(!is.na(CITY)) |> 
  rbind(na11)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m11)
city2014 <- unique(m14$CITY)


m10 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2010') 
na10 <- m10 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2010) |> 
  select(-muni_name)

m10 <- m10 |> 
  filter(!is.na(CITY)) |> 
  rbind(na11)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m10)

# Boston and Nantucket have minimum years > 2014
m16 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2016')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na16 <- m16 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2016) |> 
  select(-muni_name)
m16 <- m16 |> 
  filter(!is.na(CITY)) |> 
  rbind(na16) |> 
  filter(CITY == 'Nantucket') |> 
  mutate(SUMYEAR=2014) |> 
  rename(SHAPE=geom)

m20 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2020')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na20 <- m20 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2020) |> 
  select(-muni_name)
m20 <- m20 |> 
  filter(!is.na(CITY)) |> 
  rbind(na20) |> 
  filter(CITY == 'Boston') |> 
  mutate(SUMYEAR=2014) |> 
  rename(SHAPE=geom)

m14 <- rbind(m14,m20,m16) %>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

st_write(m14, dsn = delta_gdb, layer = 'parcels_combine_2014', append=F)

# Missing data somehow
# m14 <- st_read(dsn = delta_gdb, layer = 'parcels_combine_2014')
# ns <- names(m14)
# 
# p306 <- st_read(dsn = comp_gdb, layer = 'M306_parcel_2012_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2012,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Wales',
#          YEAR=2012,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p20 <- st_read(dsn = comp_gdb, layer = 'M020_parcel_2011_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2011,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Barnstable',
#          YEAR=2011,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p181 <- st_read(dsn = comp_gdb, layer = 'M181_parcel_2011_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2011,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Methuen',
#          YEAR=2011,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p38 <- st_read(dsn = comp_gdb, layer = 'M038_parcel_2012_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2012,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Boxford',
#          YEAR=2012,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p144 <- st_read(dsn = comp_gdb, layer = 'M144_parcel_2013_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2013,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Ipswich',
#          YEAR=2013,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p215 <- st_read(dsn = comp_gdb, layer = 'M215_parcel_2011_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2011,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Northborough',
#          YEAR=2011,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p49 <- st_read(dsn = comp_gdb, layer = 'M049_parcel_2012_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2012,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Cambridge',
#          YEAR=2012,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p258 <- st_read(dsn = comp_gdb, layer = 'M258_parcel_2011_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2011,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Salem',
#          YEAR=2011,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p201 <- st_read(dsn = comp_gdb, layer = 'M201_parcel_2012_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2012,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='New Bedford',
#          YEAR=2012,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# p153 <- st_read(dsn = comp_gdb, layer = 'M153_parcel_2013_with_assess_data') |> 
#   rename(LOC_ID=LOC_ID_2013,
#          LOT_SQFT = LOT_SIZE_G) |> 
#   mutate(CITY='Leominster',
#          YEAR=2013,
#          SUMYEAR=2014) |> 
#   select(ns)
# 
# m14m <- m14 |> 
#   filter(!CITY %in% c('Wales','Barnstable','Methuen','Boxford','Ipswich','Northborough','Cambridge','Salem','New Bedford','Leominster')) |> 
#   rbind(p306,p20,p181,p38,p144,p215,p49,p258,p201,p153)
# 
# st_write(m14m, dsn = delta_gdb, layer = 'parcels_combine_2014_fixed', append=F)

# Missing 2014: Wales 306, Barnstable 20, Methuen 181, Boxford 38, Ipswich 144, Northborough 215, Cambridge 49, Salem 258, New Bedford 201, Leominster 153



```

## 6b - Parcels combined for \>=2022

```{r}

msf <- mapcdatakeys::muni_sf(2020) |> 
  select(muni_name,geometry)

#2022
m22 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2022')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na22 <- m22 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2022) |> 
  select(-muni_name)

m22 <- m22 |> 
  filter(!is.na(CITY)) |> 
  rbind(na22) |> 
  mutate(SUMYEAR=2022)

city2022 <- unique(m22$CITY)

#2023
m23 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2023')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na23 <- m23 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2023) |> 
  select(-muni_name)

m23 <- m23 |> 
  filter(!is.na(CITY)) |> 
  rbind(na23) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m23)
city2022 <- unique(m22$CITY)

#2024
m24 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2024')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na24 <- m24 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2024) |> 
  select(-muni_name)

m24 <- m24 |> 
  filter(!is.na(CITY)) |> 
  rbind(na24) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m24)
city2022 <- unique(m22$CITY)

#2025
m25 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2025')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na25 <- m25 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2025) |> 
  select(-muni_name)

m25 <- m25 |> 
  filter(!is.na(CITY)) |> 
  rbind(na25) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m25)
city2022 <- unique(m22$CITY)

#2026
m26 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2026')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na26 <- m26 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2026) |> 
  select(-muni_name)

m26 <- m26 |> 
  filter(!is.na(CITY)) |> 
  rbind(na26) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m26)
city2022 <- unique(m22$CITY)

# Two towns with pre-2022 data only
m19 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2019')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na19 <- m19 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2019) |> 
  select(-muni_name)
m19 <- m19 |> 
  filter(!is.na(CITY)) |> 
  rbind(na19) |> 
  filter(CITY == 'Pelham') |> 
  mutate(SUMYEAR=2022)

m18 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2018')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na18 <- m18 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2018) |> 
  select(-muni_name)
m18 <- m18 |> 
  filter(!is.na(CITY)) |> 
  rbind(na18) |> 
  filter(CITY == 'Chatham') |> 
  mutate(SUMYEAR=2022)

# Final compilation
m22 <- rbind(m22,m19,m18)
m22 <- m22 %>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

st_write(m22, dsn = delta_gdb, layer = 'parcels_combine_2022', append=F)

# Missing munis

m22 <- st_read(dsn = delta_gdb, layer = 'parcels_combine_2022')
ns <- names(m22)

p201 <- st_read(dsn = comp_gdb, layer = 'M201_parcel_2023_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2023,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='New Bedford',
         YEAR=2023,
         SUMYEAR=2022) |> 
  select(ns)

p148 <- st_read(dsn = comp_gdb, layer = 'M148_parcel_2023_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2023,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='Lanesborough',
         YEAR=2023,
         SUMYEAR=2022) |> 
  select(ns)

p153 <- st_read(dsn = comp_gdb, layer = 'M153_parcel_2023_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2023,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='Leominster',
         YEAR=2023,
         SUMYEAR=2022) |> 
  select(ns)

p141 <- st_read(dsn = comp_gdb, layer = 'M141_parcel_2023_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2023,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='Hudson',
         YEAR=2023,
         SUMYEAR=2022) |> 
  select(ns)

p16 <- st_read(dsn = comp_gdb, layer = 'M016_parcel_2023_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2023,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='Attleboro',
         YEAR=2023,
         SUMYEAR=2022) |> 
  select(ns)

p292 <- st_read(dsn = comp_gdb, layer = 'M292_parcel_2023_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2023,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='Swansea',
         YEAR=2023,
         SUMYEAR=2022) |> 
  select(ns)

p346 <- st_read(dsn = comp_gdb, layer = 'M346_parcel_2024_with_assess_data') |> 
  rename(LOC_ID=LOC_ID_2024,
         LOT_SQFT = LOT_SIZE_G) |> 
  mutate(CITY='Winthrop',
         YEAR=2024,
         SUMYEAR=2022) |> 
  select(ns)

m22m <- m22 |> 
  filter(!CITY %in% c('New Bedford','Lanesborough','Leominster','Hudson','Attleboro','Swansea','Winthrop')) |> 
  rbind(p201,p148,p153,p141,p16,p292,p346)

st_write(m22m, dsn = delta_gdb, layer = 'parcels_combine_2022_fixed', append=F)
# Missing 2022: New BEdford 201, LanesBorough 148, Leominster 153, Hudson 141. Attleboro 16, Swansea 292, Winthrop 346



```

# 7 - Calculate residential and commercial parcel area

```{r}

ct20 <- mapcdatakeys::tract_sf(2020)
msf <- mapcdatakeys::muni_sf(2020) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,cmsbt08), by = 'muni_id')

res.codes <- c(
  '010',
  '012',
  '013',
  '014',
  '016',
  '017',
  '018',
  '019',
  '061',
  '071',
  '081',
  '091',
  10,
  12,
  13,
  14,
  16,
  17,
  18,
  19,
  61,
  71,
  81,
  91,
  945,
  959,
  970,
  973,
  100:131,
  133:139,
  150:199
)
com.codes <- c(
  '013',
  '014',
  '023',
  '030',
  '031',
  '032',
  '034',
  '035',
  '036',
  '037',
  '038',
  '039',
  '040',
  '041',
  '042',
  '043',
  '045',
  '046',
  '047',
  '048',
  '049',
  '063',
  '064',
  '073',
  '074',
  '083',
  '093',
  '094',
  13,
  14,
  23,
  30,
  31,
  32,
  34,
  35,
  36,
  37,
  38,
  39,
  40,
  41,
  42,
  43,
  45,
  46,
  47,
  48,
  49,
  63,
  64,
  73,
  74,
  83,
  93,
  94,
  140,
  220,
  300:393,
  400:442,
  917,
  927,
  940:943,
  955:957,
  960:969
)

# luc.tbl <- data.table(res_codes = as.character(res.codes), com_codes = as.character(com.codes))
# 2014 layers
m14 <- st_read(dsn = delta_gdb, layer = 'parcels_combine_2014_fixed') |>
  mutate(
    parcel_type = case_when(
      MIN_LUCA %in% as.character(res.codes) ~ 'RES',
      MIN_LUCA %in% as.character(com.codes) ~ 'COM'    
    )
  )
saveRDS(m14, paste0(sp_path, 'parcels_combine_2014.rds'))

res14 <- m14 |> 
  filter(parcel_type == 'RES') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(res14, dsn=task_gdb2, layer='residential_parcel_area_2014', append = F)

com14 <- m14 |> 
  filter(parcel_type == 'COM') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(com14, dsn=task_gdb2, layer='commercial_parcel_area_2014', append = F)

# 2022 layers
m22 <- st_read(dsn = delta_gdb, layer = 'parcels_combine_2022_fixed') |>
  mutate(
    parcel_type = case_when(
      MIN_LUCA %in% as.character(res.codes) ~ 'RES',
      MIN_LUCA %in% as.character(com.codes) ~ 'COM'    
    )
  ) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')
saveRDS(m22, paste0(sp_path, 'parcels_combine_2022.rds'))

res22 <- m22 |> 
  filter(parcel_type == 'RES') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(res22, dsn=task_gdb2, layer='residential_parcel_area_2022', append = F)

com22 <- m22 |> 
  filter(parcel_type == 'COM') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(com22, dsn=task_gdb2, layer='commercial_parcel_area_2022', append = F)

```

# 8 - Clip NLCD to res/com parcels

```{python}
import arcpy
from arcpy import env
arcpy.env.overwriteOutput = True

# Set environment settings
arcpy.env.workspace = r.task_gdb2

arcpy.analysis.PairwiseClip(
    in_features="nlcd2014_developed_area",
    clip_features="commercial_parcel_area_2014",
    out_feature_class="nlcd2014_developed_in_commercial_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)

arcpy.analysis.PairwiseClip(
    in_features="nlcd2014_developed_area",
    clip_features="residential_parcel_area_2014",
    out_feature_class="nlcd2014_developed_in_residential_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)

arcpy.analysis.PairwiseClip(
    in_features="nlcd2023_developed_area",
    clip_features="commercial_parcel_area_2022",
    out_feature_class="nlcd2023_developed_in_commercial_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)

arcpy.analysis.PairwiseClip(
    in_features="nlcd2023_developed_area",
    clip_features="residential_parcel_area_2022",
    out_feature_class="nlcd2023_developed_in_residential_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)
```

# 9 - Summarize RES/COM parcel areas to Tract/Muni/Community Subtype/MPO/State

```{r}

ct20 <- mapcdatakeys::tract_sf(2020)
msf <- mapcdatakeys::muni_sf(2020) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,cmsbt08), by = 'muni_id') %>%
  mutate(muni_acres = st_area(.) * 0.000247105)
munis <- munis |> left_join(msf |> st_drop_geometry() |> select(muni_id,muni_acres), by='muni_id')

m14 <- readRDS(paste0(sp_path, 'parcels_combine_2014.rds'))
m22 <- readRDS(paste0(sp_path, 'parcels_combine_2022.rds'))

# Tract, Muni, Community Subtype, MPO, Statewide summaries

par14_ct <- st_intersection(m14,ct20) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par14_ct, paste0(sp_path,'parcel2014_tract_inter.rds'))

par14_muni <- st_intersection(m14,msf) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par14_muni, paste0(sp_path,'parcel2014_muni_inter.rds'))


par22_ct <- st_intersection(m22,ct20) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par22_ct, paste0(sp_path,'parcel2022_tract_inter.rds'))

par22_muni <- st_intersection(m22,msf) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par22_muni, paste0(sp_path,'parcel2022_muni_inter.rds'))


 # Sub geography summaries
ct_res14_area <- par14_ct |> 
  filter(parcel_type=='RES') |> 
  group_by(ct20_id) |> 
  summarize(res_area_m2 = sum(area_m2))
saveRDS(ct_res14_area, paste0(sp_path, 'tract_residential_parcel_area_2014.rds'))

muni_res14_area <- par14_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(muni_id,muni_name) |> 
  summarize(res_area_m2 = sum(area_m2))
saveRDS(muni_res14_area, paste0(sp_path, 'muni_residential_parcel_area_2014.rds'))

`cmsubtyp_res14_area <- par14_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(cmsbt08) |> 
  summarize(res_area_m2 = sum(area_m2))
saveRDS(cmsubtyp_res14_area, paste0(sp_path, 'community_subtype_residential_parcel_area_2014.rds'))

sw_res14_area <- par14_muni |> 
  filter(parcel_type=='RES')  |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_res14_area <- par14_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(mpo) |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  rbind(sw_res14_area)
saveRDS(mpo_res14_area, paste0(sp_path,'mpo_and_statewide_residential_parcel_area_2014.rds'))

# Commercial
ct_com14_area <- par14_ct |> 
  filter(parcel_type=='COM') |> 
  group_by(ct20_id) |> 
  summarize(com_area_m2 = sum(area_m2))
saveRDS(ct_com14_area, paste0(sp_path,'tract_commercial_parcel_area_2014.rds'))

muni_com14_area <- par14_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(muni_id,muni_name) |> 
  summarize(com_area_m2 = sum(area_m2))
saveRDS(muni_com14_area, paste0(sp_path, 'muni_commercial_parcel_area_2014.rds'))

cmsubtyp_com14_area <- par14_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(cmsbt08) |> 
  summarize(com_area_m2 = sum(area_m2))
saveRDS(cmsubtyp_com14_area, paste0(sp_path, 'community_subtype_commercial_parcel_area_2014.rds'))

sw_com14_area <- par14_muni |> 
  filter(parcel_type=='COM') |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_com14_area <- par14_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(mpo) |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  rbind(sw_com14_area)
saveRDS(mpo_com14_area, paste0(sp_path,'mpo_and_statewide_commercial_parcel_area_2014.rds'))


# 2022 data

ct_res22_area <- par22_ct |> 
  filter(parcel_type=='RES') |> 
  group_by(ct20_id) |> 
  summarize(res_area_m2 = sum(area_m2))
saveRDS(ct_res22_area, paste0(sp_path, 'tract_residential_parcel_area_2022.rds'))

muni_res22_area <- par22_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(muni_id,muni_name) |> 
  summarize(res_area_m2 = sum(area_m2))
saveRDS(muni_res22_area, paste0(sp_path, 'muni_residential_parcel_area_2022.rds'))

cmsubtyp_res22_area <- par22_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(cmsbt08) |> 
  summarize(res_area_m2 = sum(area_m2))
saveRDS(cmsubtyp_res22_area, paste0(sp_path, 'community_subtype_residential_parcel_area_2022.rds'))

sw_res22_area <- par22_muni |> 
  filter(parcel_type=='RES')  |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_res22_area <- par22_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(mpo) |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  rbind(sw_res22_area)
saveRDS(mpo_res22_area, paste0(sp_path,'mpo_and_statewide_residential_parcel_area_2022.rds'))

# Commercial
ct_com22_area <- par22_ct |> 
  filter(parcel_type=='COM') |> 
  group_by(ct20_id) |> 
  summarize(com_area_m2 = sum(area_m2))
saveRDS(ct_com22_area, paste0(sp_path,'tract_commercial_parcel_area_2022.rds'))

muni_com22_area <- par22_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(muni_id,muni_name) |> 
  summarize(com_area_m2 = sum(area_m2))
saveRDS(muni_com22_area, paste0(sp_path, 'muni_commercial_parcel_area_2022.rds'))

cmsubtyp_com22_area <- par22_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(cmsbt08) |> 
  summarize(com_area_m2 = sum(area_m2))
saveRDS(cmsubtyp_com22_area, paste0(sp_path, 'community_subtype_commercial_parcel_area_2022.rds'))

sw_com22_area <- par22_muni |> 
  filter(parcel_type=='COM') |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_com22_area <- par22_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(mpo) |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  rbind(sw_com22_area)
saveRDS(mpo_com22_area, paste0(sp_path,'mpo_and_statewide_commercial_parcel_area_2022.rds'))


```

# 10a - Tract Denominators

```{r}

# Parcel denominators
res14.tract <- readRDS(paste0(sp_path, 'tract_residential_parcel_area_2014.rds')) |> 
  mutate(
    res_acres_2014 = res_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, res_acres_2014)

com14.tract <- readRDS(paste0(sp_path, 'tract_commercial_parcel_area_2014.rds')) |> 
  mutate(
    com_acres_2014 = com_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, com_acres_2014)

res23.tract <- readRDS(paste0(sp_path, 'tract_residential_parcel_area_2022.rds')) |> 
  mutate(
    res_acres_2023 = res_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, res_acres_2023)

com23.tract <- readRDS(paste0(sp_path, 'tract_commercial_parcel_area_2022.rds')) |> 
  mutate(
    com_acres_2023 = com_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, com_acres_2023)

# NLCD Developed Area denominators
nlcd14.tract <- readRDS(paste0(sp_path, 'nlcd2014_dev_area_tract.rds')) |> 
  mutate(
    dev_acres_2014 = as.numeric(developed_area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_acres_2014)

nlcd23.tract <- readRDS(paste0(sp_path, 'nlcd2023_dev_area_tract.rds')) |> 
  mutate(
    dev_acres_2023 = as.numeric(developed_area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_acres_2023)

# NLCD within parcel denominators
dev.res14.tract <- st_read(dsn=task_gdb2, layer='nlcd2014_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2014 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_res_acres_2014) |> 
  group_by(ct20_id) |> 
  summarize(dev_res_acres_2014 = sum(dev_res_acres_2014))
saveRDS(dev.res14.tract,paste0(sp_path,'nlcd2014_dev_res_tract.rds'))

dev.com14.tract <- st_read(dsn=task_gdb2, layer='nlcd2014_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2014 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_com_acres_2014) |> 
  group_by(ct20_id) |> 
  summarize(dev_com_acres_2014 = sum(dev_com_acres_2014))
saveRDS(dev.com14.tract,paste0(sp_path,'nlcd2014_dev_com_tract.rds'))

dev.res23.tract <- st_read(dsn=task_gdb2, layer='nlcd2023_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2023 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_res_acres_2023) |> 
  group_by(ct20_id) |> 
  summarize(dev_res_acres_2023 = sum(dev_res_acres_2023))
saveRDS(dev.res23.tract,paste0(sp_path,'nlcd2023_dev_res_tract.rds'))

dev.com23.tract <- st_read(dsn=task_gdb2, layer='nlcd2023_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2023 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_com_acres_2023) |> 
  group_by(ct20_id) |> 
  summarize(dev_com_acres_2023 = sum(dev_com_acres_2023))
saveRDS(dev.com23.tract,paste0(sp_path,'nlcd2023_dev_com_tract.rds'))

```

# 10b - Tract Metrics

```{r}
# Merge numerators and calculate metrics

met.ct <- left_join(
  ct20, acs14.tract, by='ct20_id') |> 
  left_join(acs23.tract, by='ct20_id') |> 
  left_join(emp.ct, by='ct20_id') |> 
  left_join(res14.tract, by='ct20_id') |> 
  left_join(res23.tract, by='ct20_id') |> 
  left_join(com14.tract, by='ct20_id') |> 
  left_join(com23.tract, by='ct20_id') |>
  left_join(nlcd14.tract, by='ct20_id') |>
  left_join(nlcd23.tract, by='ct20_id') |>
  left_join(dev.res14.tract, by='ct20_id') |>
  left_join(dev.com14.tract, by='ct20_id') |>
  left_join(dev.res23.tract, by='ct20_id') |>
  left_join(dev.com23.tract, by='ct20_id')

met.ct[is.na(met.ct)] <- 0

# Metrics by parcel acreage of res/com
met.res.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    pop_acre_2014 = total_pop_2014 / res_acres_2014,
    hh_acre_2014 = total_households_2014 / res_acres_2014,
    units_acre_2014 = total_units_2014 / res_acres_2014,
    units_1_acre_2014 = units_1_2014 / res_acres_2014,
    units_2_acre_2014 = units_2_2014 / res_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / res_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / res_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / res_acres_2014,
    pop_acre_2023 = total_pop_2023 / res_acres_2023,
    hh_acre_2023 = total_households_2023 / res_acres_2023,
    units_acre_2023 = total_units_2023 / res_acres_2023,
    units_1_acre_2023 = units_1_2023 / res_acres_2023,
    units_2_acre_2023 = units_2_2023 / res_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / res_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / res_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / res_acres_2023,
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(res_acres_2023),
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014
    )

st_write(met.res.ct, dsn=task_gdb2, layer = 'pop_hh_unit_res_acre_tract', append=F)
fwrite(met.res.ct |> st_drop_geometry(), paste0(deliv_path,'/pop_hh_unit_res_acre_tract.csv'))

met.com.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    emp_acre_2014 = emp2014 / com_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / com_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / com_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / com_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / com_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / com_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / com_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / com_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / com_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / com_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / com_acres_2014,
    emp_acre_2023 = emp2023 / com_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / com_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / com_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / com_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / com_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / com_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / com_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / com_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / com_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / com_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / com_acres_2023,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014
  )

st_write(met.com.ct, dsn=task_gdb2, layer = 'emp_com_acre_tract', append=F)
fwrite(met.com.ct |> st_drop_geometry(), paste0(deliv_path,'/emp_com_acre_tract.csv'))

# Metrics by NLCD developed land acreage
met.dev.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    pop_acre_2014 = total_pop_2014 / dev_acres_2014,
    hh_acre_2014 = total_households_2014 / dev_acres_2014,
    units_acre_2014 = total_units_2014 / dev_acres_2014,
    units_1_acre_2014 = units_1_2014 / dev_acres_2014,
    units_2_acre_2014 = units_2_2014 / dev_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / dev_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / dev_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / dev_acres_2014,
    pop_acre_2023 = total_pop_2023 / dev_acres_2023,
    hh_acre_2023 = total_households_2023 / dev_acres_2023,
    units_acre_2023 = total_units_2023 / dev_acres_2023,
    units_1_acre_2023 = units_1_2023 / dev_acres_2023,
    units_2_acre_2023 = units_2_2023 / dev_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / dev_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / dev_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / dev_acres_2023,
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_acres_2023),
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    emp_acre_2014 = emp2014 / dev_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / dev_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / dev_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / dev_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / dev_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / dev_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / dev_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / dev_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / dev_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / dev_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / dev_acres_2014,
    emp_acre_2023 = emp2023 / dev_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / dev_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / dev_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / dev_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / dev_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / dev_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / dev_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / dev_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / dev_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / dev_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / dev_acres_2023,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014
  )

st_write(met.dev.ct, dsn=task_gdb2, layer = 'pop_hh_units_emp_dev_acre_tract', append=F)
fwrite(met.dev.ct |> st_drop_geometry(), paste0(deliv_path,'/pop_hh_units_emp_dev_acre_tract.csv'))

# Metrics by NLCD developed land IN parcel res/com acreage

met.dev.res.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    res_acres_2014 = res_acres_2014,
    dev_acres_2014 = dev_acres_2014,
    dev_res_acres_2014 = dev_res_acres_2014,
    res_acres_2023 = res_acres_2023,
    dev_acres_2023 = dev_acres_2023,
    dev_res_acres_2023 = dev_res_acres_2023,
    chg_res_acres = res_acres_2023 - res_acres_2014,
    chg_dev_acres = dev_acres_2023 - dev_acres_2014,
    chg_dev_res_acres = dev_res_acres_2023 - dev_res_acres_2014,
    pct_chg_res_acres = 100*(res_acres_2023 - res_acres_2014)/res_acres_2014,
    pct_chg_dev_acres = 100*(dev_acres_2023 - dev_acres_2014)/dev_acres_2014,
    pct_chg_dev_res_acres = 100*(dev_res_acres_2023 - dev_res_acres_2014)/dev_res_acres_2014,
    pop_acre_2014 = total_pop_2014 / dev_res_acres_2014,
    hh_acre_2014 = total_households_2014 / dev_res_acres_2014,
    units_acre_2014 = total_units_2014 / dev_res_acres_2014,
    pop_acre_2023 = total_pop_2023 / dev_res_acres_2023,
    hh_acre_2023 = total_households_2023 / dev_res_acres_2023,
    units_acre_2023 = total_units_2023 / dev_res_acres_2023,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    units_1_acre_2014 = units_1_2014 / dev_res_acres_2014,
    units_2_acre_2014 = units_2_2014 / dev_res_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / dev_res_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / dev_res_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / dev_res_acres_2014,
    units_1_acre_2023 = units_1_2023 / dev_res_acres_2023,
    units_2_acre_2023 = units_2_2023 / dev_res_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / dev_res_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / dev_res_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / dev_res_acres_2023,
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_res_acres_2023),
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014
    )

st_write(met.dev.res.ct, dsn=task_gdb2, layer = 'pop_hh_unit_nlcd_dev_in_res_acre_tract', append = FALSE)
fwrite(met.dev.res.ct |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_unit_nlcd_dev_in_res_acre_tract.csv'))

met.dev.com.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    com_acres_2014 = com_acres_2014,
    dev_acres_2014 = dev_acres_2014,
    dev_com_acres_2014 = dev_com_acres_2014,
    com_acres_2023 = com_acres_2023,
    dev_acres_2023 = dev_acres_2023,
    dev_com_acres_2023 = dev_com_acres_2023,
    chg_com_acres = com_acres_2023 - com_acres_2014,
    chg_dev_acres = dev_acres_2023 - dev_acres_2014,
    chg_dev_com_acres = dev_com_acres_2023 - dev_com_acres_2014,
    pct_chg_com_acres = 100*(com_acres_2023 - com_acres_2014)/com_acres_2014,
    pct_chg_dev_acres = 100*(dev_acres_2023 - dev_acres_2014)/dev_acres_2014,
    pct_chg_dev_com_acres = 100*(dev_com_acres_2023 - dev_com_acres_2014)/dev_com_acres_2014,
    emp_acre_2014 = emp2014 / dev_com_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / dev_com_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / dev_com_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / dev_com_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / dev_com_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / dev_com_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / dev_com_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / dev_com_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / dev_com_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / dev_com_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / dev_com_acres_2014,
    emp_acre_2023 = emp2023 / dev_com_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / dev_com_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / dev_com_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / dev_com_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / dev_com_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / dev_com_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / dev_com_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / dev_com_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / dev_com_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / dev_com_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / dev_com_acres_2023,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014
  )

st_write(met.dev.com.ct, dsn=task_gdb2, layer = 'emp_nlcd_dev_in_com_acre_tract', append=F)
fwrite(met.dev.com.ct |> st_drop_geometry(), paste0(deliv_path, '/emp_nlcd_dev_in_com_acre_tract.csv'))

```

# 11a - Muni Denominators

```{r}

msf <- st_read(dsn=task_gdb2,layer='munis_no_coast') |> 
  rename(muni_id = TOWN_ID) |> 
  select(muni_id,SHAPE) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,muni_name,mpo,cmsbt08), by='muni_id') %>%
  mutate(muni_acres = as.numeric(st_area(.)) * 0.000247105)

munis <-  st_read(dsn=task_gdb2,layer='munis_no_coast') |> 
  rename(muni_id = TOWN_ID) |> 
  select(muni_id,SHAPE) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,muni_name,mpo,cmsbt08), by='muni_id') %>%
  mutate(muni_acres = as.numeric(st_area(.)) * 0.000247105) |> 
  setDT()

# Parcel denominators
res14.muni <- readRDS(paste0(sp_path, 'muni_residential_parcel_area_2014.rds')) |> 
  left_join(munis |> select(muni_id,mpo,cmsbt08,muni_acres), by = 'muni_id') |> 
  mutate(
    res_acres_2014 = res_area_m2 * 0.000247105,
    res_frac = res_acres_2014 / muni_acres
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, res_acres_2014, res_frac)

com14.muni <- readRDS(paste0(sp_path, 'muni_commercial_parcel_area_2014.rds')) |> 
  left_join(munis |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    com_acres_2014 = com_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, com_acres_2014)

res23.muni <- readRDS(paste0(sp_path, 'muni_residential_parcel_area_2022.rds')) |> 
  left_join(munis |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    res_acres_2023 = res_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, res_acres_2023)

com23.muni <- readRDS(paste0(sp_path, 'muni_commercial_parcel_area_2022.rds')) |> 
  left_join(munis |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    com_acres_2023 = com_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, com_acres_2023)

# NLCD Developed Area denominators
nlcd14.muni <- readRDS(paste0(sp_path, 'nlcd2014_dev_area_muni.rds')) |> 
  left_join(munis |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    dev_acres_2014 = as.numeric(developed_area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_acres_2014)

nlcd23.muni <- readRDS(paste0(sp_path, 'nlcd2023_dev_area_muni.rds')) |> 
  left_join(munis |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    dev_acres_2023 = as.numeric(developed_area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_acres_2023)

# NLCD within parcel denominators
dev.res14.muni <- st_read(dsn=task_gdb2, layer='nlcd2014_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(msf)) %>%
  st_intersection(msf) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2014 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_res_acres_2014) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_res_acres_2014 = sum(dev_res_acres_2014))
saveRDS(dev.res14.muni,paste0(sp_path,'nlcd2014_dev_res_muni.rds'))

dev.com14.muni <- st_read(dsn=task_gdb2, layer='nlcd2014_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(msf)) %>%
  st_intersection(msf) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2014 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_com_acres_2014) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_com_acres_2014 = sum(dev_com_acres_2014))
saveRDS(dev.com14.muni,paste0(sp_path,'nlcd2014_dev_com_muni.rds'))

dev.res23.muni <- st_read(dsn=task_gdb2, layer='nlcd2023_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(msf)) %>%
  st_intersection(msf) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2023 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_res_acres_2023) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_res_acres_2023 = sum(dev_res_acres_2023))
saveRDS(dev.res23.muni,paste0(sp_path,'nlcd2023_dev_res_muni.rds'))

dev.com23.muni <- st_read(dsn=task_gdb2, layer='nlcd2023_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(msf)) %>%
  st_intersection(msf) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2023 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_com_acres_2023) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_com_acres_2023 = sum(dev_com_acres_2023))
saveRDS(dev.com23.muni,paste0(sp_path,'nlcd2023_dev_com_muni.rds'))

```

# 11b - Muni Metrics

```{r}
# Muni Metrics

# Merge numerators and calculate metrics

met.muni <- left_join(
  msf, acs14.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |> 
  left_join(acs23.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |> 
  left_join(emp.muni, by=c('muni_id','muni_name')) |> 
  left_join(res14.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |> 
  left_join(res23.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |> 
  left_join(com14.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |> 
  left_join(com23.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |>
  left_join(nlcd14.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |>
  left_join(nlcd23.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |>
  left_join(dev.res14.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |>
  left_join(dev.com14.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |>
  left_join(dev.res23.muni, by=c('muni_id','muni_name','mpo','cmsbt08')) |>
  left_join(dev.com23.muni, by=c('muni_id','muni_name','mpo','cmsbt08'))

met.muni[is.na(met.muni)] <- 0

#
# Metrics by parcel acreage of res/com
met.res.muni <- met.muni |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(
    pop_acre_2014 = total_pop_2014 / res_acres_2014,
    hh_acre_2014 = total_households_2014 / res_acres_2014,
    units_acre_2014 = total_units_2014 / res_acres_2014,
    units_1_acre_2014 = units_1_2014 / res_acres_2014,
    units_2_acre_2014 = units_2_2014 / res_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / res_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / res_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / res_acres_2014,
    pop_acre_2023 = total_pop_2023 / res_acres_2023,
    hh_acre_2023 = total_households_2023 / res_acres_2023,
    units_acre_2023 = total_units_2023 / res_acres_2023,
    units_1_acre_2023 = units_1_2023 / res_acres_2023,
    units_2_acre_2023 = units_2_2023 / res_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / res_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / res_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / res_acres_2023,
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(res_acres_2023),
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_res_acres = res_acres_2023 - res_acres_2014,
    pct_chg_res_acres = round(100*(res_acres_2023 - res_acres_2014)/res_acres_2014,1)
    )

st_write(met.res.muni, dsn=task_gdb2, layer = 'pop_hh_unit_res_acre_muni', append = F)
fwrite(met.res.muni |> st_drop_geometry(), paste0(deliv_path,'/pop_hh_unit_res_acre_muni.csv'))

met.com.muni <- met.muni |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(
    emp_acre_2014 = emp2014 / com_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / com_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / com_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / com_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / com_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / com_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / com_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / com_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / com_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / com_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / com_acres_2014,
    emp_acre_2023 = emp2023 / com_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / com_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / com_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / com_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / com_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / com_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / com_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / com_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / com_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / com_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / com_acres_2023,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_com_acres = com_acres_2023 - com_acres_2014,
    pct_chg_com_acres = round(100*(com_acres_2023 - com_acres_2014)/com_acres_2014,1)
  )

st_write(met.com.muni, dsn=task_gdb2, layer = 'emp_com_acre_muni', append=F)
fwrite(met.com.muni |> st_drop_geometry(), paste0(deliv_path,'/emp_com_acre_muni.csv'))

# Metrics by NLCD developed land acreage
met.dev.muni <- met.muni |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(
    dev_acres_2014 = sum(dev_acres_2014),
    dev_acres_2023 = sum(dev_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_acres_2023),
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_acres_2023),
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_acres = dev_acres_2023 - dev_acres_2014,
    pct_chg_dev_acres = round(100*(dev_acres_2023 - dev_acres_2014)/dev_acres_2014,1)
  )

st_write(met.dev.muni, dsn=task_gdb2, layer = 'pop_hh_units_emp_dev_acre_muni', append =F)
fwrite(met.dev.muni |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_units_emp_dev_acre_muni.csv'))

# Metrics by NLCD developed land IN parcel res/com acreage

met.dev.res.muni <- met.muni |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(
    dev_res_acres_2014 = sum(dev_res_acres_2014),
    dev_res_acres_2023 = sum(dev_res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_res_acres_2023),
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_res_acres_2023),
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_dev_res_acres = dev_res_acres_2023 - dev_res_acres_2014,
    pct_chg_dev_res_acres = round(100*(dev_res_acres_2023 - dev_res_acres_2014)/dev_res_acres_2014,1)
    )

st_write(met.dev.res.muni, dsn=task_gdb2, layer = 'pop_hh_unit_nlcd_dev_in_res_acre_muni', append=F)
fwrite(met.dev.res.muni |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_unit_nlcd_dev_in_res_acre_muni.csv'))

met.dev.com.muni <- met.muni |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |>  
  summarize(
    dev_com_acres_2014 = sum(dev_com_acres_2014),
    dev_com_acres_2023 = sum(dev_com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_com_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_com_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_com_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_com_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_com_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_com_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_com_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_com_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_com_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_com_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_com_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_com_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_com_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_com_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_com_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_com_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_com_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_com_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_com_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_com_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_com_acres_2023),
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_com_acres = dev_com_acres_2023 - dev_com_acres_2014,
    pct_chg_dev_com_acres = round(100*(dev_com_acres_2023 - dev_com_acres_2014)/dev_com_acres_2014,1)
  )

st_write(met.dev.com.muni, dsn=task_gdb2, layer = 'emp_nlcd_dev_in_com_acre_muni', append = F)
fwrite(met.dev.com.muni |> st_drop_geometry(), paste0(deliv_path, '/emp_nlcd_dev_in_com_acre_muni.csv'))

```

# 12 - Community Subtype Metrics

```{r}
# Community Subtype

# Metrics by parcel acreage of res/com
# Metrics by parcel acreage of res/com

met.res.cmsbt08 <- met.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    res_acres_2014 = sum(res_acres_2014),
    res_acres_2023 = sum(res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(res_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(res_acres_2023)
    ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_res_acres = res_acres_2023 - res_acres_2014,
    pct_chg_res_acres = round(100*(chg_res_acres/sum(res_acres_2014)),1)
    )

st_write(met.res.cmsbt08, dsn=task_gdb2, layer = 'pop_hh_unit_res_acre_cmsbt08', append = F)
fwrite(met.res.cmsbt08 |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_unit_res_acre_cmsbt08.csv'))

met.com.cmsbt08 <- met.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    com_acres_2014 = sum(com_acres_2014),
    com_acres_2023 = sum(com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(com_acres_2014),
    emp_sect_1_acre_2014 = sum(emp_2014_sect_1) / sum(com_acres_2014),
    emp_sect_2_acre_2014 = sum(emp_2014_sect_2) / sum(com_acres_2014),
    emp_sect_3_acre_2014 = sum(emp_2014_sect_3) / sum(com_acres_2014),
    emp_sect_4_acre_2014 = sum(emp_2014_sect_4) / sum(com_acres_2014),
    emp_sect_5_acre_2014 = sum(emp_2014_sect_5) / sum(com_acres_2014),
    emp_sect_6_acre_2014 = sum(emp_2014_sect_6) / sum(com_acres_2014),
    emp_sect_7_acre_2014 = sum(emp_2014_sect_7) / sum(com_acres_2014),
    emp_sect_8_acre_2014 = sum(emp_2014_sect_8) / sum(com_acres_2014),
    emp_sect_9_acre_2014 = sum(emp_2014_sect_9) / sum(com_acres_2014),
    emp_sect_10_acre_2014 = sum(emp_2014_sect_10) / sum(com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(com_acres_2023),
    emp_sect_1_acre_2023 = sum(emp_2023_sect_1) / sum(com_acres_2023),
    emp_sect_2_acre_2023 = sum(emp_2023_sect_2) / sum(com_acres_2023),
    emp_sect_3_acre_2023 = sum(emp_2023_sect_3) / sum(com_acres_2023),
    emp_sect_4_acre_2023 = sum(emp_2023_sect_4) / sum(com_acres_2023),
    emp_sect_5_acre_2023 = sum(emp_2023_sect_5) / sum(com_acres_2023),
    emp_sect_6_acre_2023 = sum(emp_2023_sect_6) / sum(com_acres_2023),
    emp_sect_7_acre_2023 = sum(emp_2023_sect_7) / sum(com_acres_2023),
    emp_sect_8_acre_2023 = sum(emp_2023_sect_8) / sum(com_acres_2023),
    emp_sect_9_acre_2023 = sum(emp_2023_sect_9) / sum(com_acres_2023),
    emp_sect_10_acre_2023 = sum(emp_2023_sect_10) / sum(com_acres_2023),
    chg_emp_sect_1_acre = sum(emp_2023_sect_1 - emp_2014_sect_1),
    chg_emp_sect_2_acre = sum(emp_2023_sect_2 - emp_2014_sect_2),
    chg_emp_sect_3_acre = sum(emp_2023_sect_3 - emp_2014_sect_3),
    chg_emp_sect_4_acre = sum(emp_2023_sect_4 - emp_2014_sect_4),
    chg_emp_sect_5_acre = sum(emp_2023_sect_5 - emp_2014_sect_5),
    chg_emp_sect_6_acre = sum(emp_2023_sect_6 - emp_2014_sect_6),
    chg_emp_sect_7_acre = sum(emp_2023_sect_7 - emp_2014_sect_7),
    chg_emp_sect_8_acre = sum(emp_2023_sect_8 - emp_2014_sect_8),
    chg_emp_sect_9_acre = sum(emp_2023_sect_9 - emp_2014_sect_9),
    chg_emp_sect_10_acre = sum(emp_2023_sect_10 - emp_2014_sect_10)
  ) |> 
  mutate(
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_com_acres = com_acres_2023 - com_acres_2014,
    pct_chg_com_acres = round(100*(com_acres_2023 - com_acres_2014)/com_acres_2014,1)
  )

st_write(met.com.cmsbt08, dsn=task_gdb2, layer = 'emp_com_acre_cmsbt08', append=F)
fwrite(met.com.cmsbt08 |> st_drop_geometry(), paste0(deliv_path,'/emp_com_acre_cmsbt08.csv'))

# Metrics by NLCD developed land acreage
met.dev.cmsbt08 <- met.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    dev_acres_2014 = sum(dev_acres_2014),
    dev_acres_2023 = sum(dev_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_acres_2023)
  ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_acres = dev_acres_2023 - dev_acres_2014,
    pct_chg_dev_acres = round(100*(dev_acres_2023 - dev_acres_2014)/dev_acres_2014,1)
  )

st_write(met.dev.cmsbt08, dsn=task_gdb2, layer = 'pop_hh_units_emp_dev_acre_cmsbt08', append =F)
fwrite(met.dev.cmsbt08 |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_units_emp_dev_acre_cmsbt08.csv'))

# Metrics by NLCD developed land IN parcel res/com acreage

met.dev.res.cmsbt08 <- met.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    dev_res_acres_2014 = sum(dev_res_acres_2014),
    dev_res_acres_2023 = sum(dev_res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_res_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_res_acres_2023)
    ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_dev_res_acres = dev_res_acres_2023 - dev_res_acres_2014,
    pct_chg_dev_res_acres = round(100*(dev_res_acres_2023 - dev_res_acres_2014)/dev_res_acres_2014,1)
    )

st_write(met.dev.res.cmsbt08, dsn=task_gdb2, layer = 'pop_hh_unit_nlcd_dev_in_res_acre_cmsbt08', append=F)

fwrite(met.dev.res.cmsbt08 |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_unit_nlcd_dev_in_res_acre_cmsbt08.csv'))

met.dev.com.cmsbt08 <- met.muni |> 
  group_by(cmsbt08) |>  
  summarize(
    dev_com_acres_2014 = sum(dev_com_acres_2014),
    dev_com_acres_2023 = sum(dev_com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_com_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_com_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_com_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_com_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_com_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_com_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_com_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_com_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_com_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_com_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_com_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_com_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_com_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_com_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_com_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_com_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_com_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_com_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_com_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_com_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_com_acres_2023)
  )|> 
  mutate(
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_com_acres = dev_com_acres_2023 - dev_com_acres_2014,
    pct_chg_dev_com_acres = round(100*(dev_com_acres_2023 - dev_com_acres_2014)/dev_com_acres_2014,1)
  )

st_write(met.dev.com.cmsbt08, dsn=task_gdb2, layer = 'emp_nlcd_dev_in_com_acre_cmsbt08', append = F)
fwrite(met.dev.com.cmsbt08 |> st_drop_geometry(), paste0(deliv_path, '/emp_nlcd_dev_in_com_acre_cmsbt08.csv'))
```

# 13 - MPO and statewide

```{r}

# Metrics by parcel acreage of res/com
met.res.sw <- met.muni |> 
  summarize(
    res_acres_2014 = sum(res_acres_2014),
    res_acres_2023 = sum(res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(res_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(res_acres_2023)
    ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_res_acres = res_acres_2023 - res_acres_2014,
    pct_chg_res_acres = round(100*(chg_res_acres/sum(res_acres_2014)),1)
    ) |> 
  mutate(mpo='STATEWIDE')

met.res.mpo <- met.muni |> 
  group_by(mpo) |> 
  summarize(
    res_acres_2014 = sum(res_acres_2014),
    res_acres_2023 = sum(res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(res_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(res_acres_2023)
    ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_res_acres = res_acres_2023 - res_acres_2014,
    pct_chg_res_acres = round(100*(chg_res_acres/sum(res_acres_2014)),1)
    ) |> 
  rbind(met.res.sw)%>%
  st_make_valid(.)%>%
  st_cast('MULTIPOLYGON')

st_write(met.res.mpo, dsn=task_gdb2, layer = 'pop_hh_unit_res_acre_mpo', append = F)
fwrite(met.res.mpo |> st_drop_geometry(), paste0(deliv_path,'/pop_hh_unit_res_acre_mpo.csv'))

met.com.sw <- met.muni |> 
  summarize(
    com_acres_2014 = sum(com_acres_2014),
    com_acres_2023 = sum(com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(com_acres_2014),
    emp_sect_1_acre_2014 = sum(emp_2014_sect_1) / sum(com_acres_2014),
    emp_sect_2_acre_2014 = sum(emp_2014_sect_2) / sum(com_acres_2014),
    emp_sect_3_acre_2014 = sum(emp_2014_sect_3) / sum(com_acres_2014),
    emp_sect_4_acre_2014 = sum(emp_2014_sect_4) / sum(com_acres_2014),
    emp_sect_5_acre_2014 = sum(emp_2014_sect_5) / sum(com_acres_2014),
    emp_sect_6_acre_2014 = sum(emp_2014_sect_6) / sum(com_acres_2014),
    emp_sect_7_acre_2014 = sum(emp_2014_sect_7) / sum(com_acres_2014),
    emp_sect_8_acre_2014 = sum(emp_2014_sect_8) / sum(com_acres_2014),
    emp_sect_9_acre_2014 = sum(emp_2014_sect_9) / sum(com_acres_2014),
    emp_sect_10_acre_2014 = sum(emp_2014_sect_10) / sum(com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(com_acres_2023),
    emp_sect_1_acre_2023 = sum(emp_2023_sect_1) / sum(com_acres_2023),
    emp_sect_2_acre_2023 = sum(emp_2023_sect_2) / sum(com_acres_2023),
    emp_sect_3_acre_2023 = sum(emp_2023_sect_3) / sum(com_acres_2023),
    emp_sect_4_acre_2023 = sum(emp_2023_sect_4) / sum(com_acres_2023),
    emp_sect_5_acre_2023 = sum(emp_2023_sect_5) / sum(com_acres_2023),
    emp_sect_6_acre_2023 = sum(emp_2023_sect_6) / sum(com_acres_2023),
    emp_sect_7_acre_2023 = sum(emp_2023_sect_7) / sum(com_acres_2023),
    emp_sect_8_acre_2023 = sum(emp_2023_sect_8) / sum(com_acres_2023),
    emp_sect_9_acre_2023 = sum(emp_2023_sect_9) / sum(com_acres_2023),
    emp_sect_10_acre_2023 = sum(emp_2023_sect_10) / sum(com_acres_2023),
    chg_emp_sect_1_acre = sum(emp_2023_sect_1 - emp_2014_sect_1),
    chg_emp_sect_2_acre = sum(emp_2023_sect_2 - emp_2014_sect_2),
    chg_emp_sect_3_acre = sum(emp_2023_sect_3 - emp_2014_sect_3),
    chg_emp_sect_4_acre = sum(emp_2023_sect_4 - emp_2014_sect_4),
    chg_emp_sect_5_acre = sum(emp_2023_sect_5 - emp_2014_sect_5),
    chg_emp_sect_6_acre = sum(emp_2023_sect_6 - emp_2014_sect_6),
    chg_emp_sect_7_acre = sum(emp_2023_sect_7 - emp_2014_sect_7),
    chg_emp_sect_8_acre = sum(emp_2023_sect_8 - emp_2014_sect_8),
    chg_emp_sect_9_acre = sum(emp_2023_sect_9 - emp_2014_sect_9),
    chg_emp_sect_10_acre = sum(emp_2023_sect_10 - emp_2014_sect_10)
  ) |> 
  mutate(
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_com_acres = com_acres_2023 - com_acres_2014,
    pct_chg_com_acres = round(100*(com_acres_2023 - com_acres_2014)/com_acres_2014,1)
  ) |> 
  mutate(mpo='STATEWIDE')

met.com.mpo <- met.muni |> 
  group_by(mpo) |> 
  summarize(
    com_acres_2014 = sum(com_acres_2014),
    com_acres_2023 = sum(com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(com_acres_2014),
    emp_sect_1_acre_2014 = sum(emp_2014_sect_1) / sum(com_acres_2014),
    emp_sect_2_acre_2014 = sum(emp_2014_sect_2) / sum(com_acres_2014),
    emp_sect_3_acre_2014 = sum(emp_2014_sect_3) / sum(com_acres_2014),
    emp_sect_4_acre_2014 = sum(emp_2014_sect_4) / sum(com_acres_2014),
    emp_sect_5_acre_2014 = sum(emp_2014_sect_5) / sum(com_acres_2014),
    emp_sect_6_acre_2014 = sum(emp_2014_sect_6) / sum(com_acres_2014),
    emp_sect_7_acre_2014 = sum(emp_2014_sect_7) / sum(com_acres_2014),
    emp_sect_8_acre_2014 = sum(emp_2014_sect_8) / sum(com_acres_2014),
    emp_sect_9_acre_2014 = sum(emp_2014_sect_9) / sum(com_acres_2014),
    emp_sect_10_acre_2014 = sum(emp_2014_sect_10) / sum(com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(com_acres_2023),
    emp_sect_1_acre_2023 = sum(emp_2023_sect_1) / sum(com_acres_2023),
    emp_sect_2_acre_2023 = sum(emp_2023_sect_2) / sum(com_acres_2023),
    emp_sect_3_acre_2023 = sum(emp_2023_sect_3) / sum(com_acres_2023),
    emp_sect_4_acre_2023 = sum(emp_2023_sect_4) / sum(com_acres_2023),
    emp_sect_5_acre_2023 = sum(emp_2023_sect_5) / sum(com_acres_2023),
    emp_sect_6_acre_2023 = sum(emp_2023_sect_6) / sum(com_acres_2023),
    emp_sect_7_acre_2023 = sum(emp_2023_sect_7) / sum(com_acres_2023),
    emp_sect_8_acre_2023 = sum(emp_2023_sect_8) / sum(com_acres_2023),
    emp_sect_9_acre_2023 = sum(emp_2023_sect_9) / sum(com_acres_2023),
    emp_sect_10_acre_2023 = sum(emp_2023_sect_10) / sum(com_acres_2023),
    chg_emp_sect_1_acre = sum(emp_2023_sect_1 - emp_2014_sect_1),
    chg_emp_sect_2_acre = sum(emp_2023_sect_2 - emp_2014_sect_2),
    chg_emp_sect_3_acre = sum(emp_2023_sect_3 - emp_2014_sect_3),
    chg_emp_sect_4_acre = sum(emp_2023_sect_4 - emp_2014_sect_4),
    chg_emp_sect_5_acre = sum(emp_2023_sect_5 - emp_2014_sect_5),
    chg_emp_sect_6_acre = sum(emp_2023_sect_6 - emp_2014_sect_6),
    chg_emp_sect_7_acre = sum(emp_2023_sect_7 - emp_2014_sect_7),
    chg_emp_sect_8_acre = sum(emp_2023_sect_8 - emp_2014_sect_8),
    chg_emp_sect_9_acre = sum(emp_2023_sect_9 - emp_2014_sect_9),
    chg_emp_sect_10_acre = sum(emp_2023_sect_10 - emp_2014_sect_10)
  ) |> 
  mutate(
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_com_acres = com_acres_2023 - com_acres_2014,
    pct_chg_com_acres = round(100*(com_acres_2023 - com_acres_2014)/com_acres_2014,1)
  ) |> 
  rbind(met.com.sw)%>%
  st_make_valid(.)%>%
  st_cast('MULTIPOLYGON')

st_write(met.com.mpo, dsn=task_gdb2, layer = 'emp_com_acre_mpo', append=F)
fwrite(met.com.mpo |> st_drop_geometry(), paste0(deliv_path,'/emp_com_acre_mpo.csv'))


# Metrics by NLCD developed land acreage
met.dev.sw <- met.muni |>
  summarize(
    dev_acres_2014 = sum(dev_acres_2014),
    dev_acres_2023 = sum(dev_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_acres_2023)
  ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_acres = dev_acres_2023 - dev_acres_2014,
    pct_chg_dev_acres = round(100*(dev_acres_2023 - dev_acres_2014)/dev_acres_2014,1)
  ) |> 
  mutate(mpo='STATEWIDE')

met.dev.mpo <- met.muni |> 
  group_by(mpo) |> 
  summarize(
    dev_acres_2014 = sum(dev_acres_2014),
    dev_acres_2023 = sum(dev_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_acres_2023)
  ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_acres = dev_acres_2023 - dev_acres_2014,
    pct_chg_dev_acres = round(100*(dev_acres_2023 - dev_acres_2014)/dev_acres_2014,1)
  ) %>%
  st_make_valid(.)%>%
  st_cast('MULTIPOLYGON') |> 
  rbind(met.dev.sw)

st_write(met.dev.mpo, dsn=task_gdb2, layer = 'pop_hh_units_emp_dev_acre_mpo', append =F)
fwrite(met.dev.mpo |> st_drop_geometry(), paste0(deliv_path,'/pop_hh_units_emp_dev_acre_mpo.csv'))

# Metrics by NLCD developed land IN parcel res/com acreage

met.dev.res.sw <- met.muni |> 
  summarize(
    dev_res_acres_2014 = sum(dev_res_acres_2014),
    dev_res_acres_2023 = sum(dev_res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_res_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_res_acres_2023)
    ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_dev_res_acres = dev_res_acres_2023 - dev_res_acres_2014,
    pct_chg_dev_res_acres = round(100*(dev_res_acres_2023 - dev_res_acres_2014)/dev_res_acres_2014,1)
    ) |> 
  mutate(mpo='STATEWIDE')

met.dev.res.mpo <- met.muni |> 
  group_by(mpo) |> 
  summarize(
    dev_res_acres_2014 = sum(dev_res_acres_2014),
    dev_res_acres_2023 = sum(dev_res_acres_2023),
    pop_acre_2014 = sum(total_pop_2014) / sum(dev_res_acres_2014),
    hh_acre_2014 = sum(total_households_2014) / sum(dev_res_acres_2014),
    units_acre_2014 = sum(total_units_2014) / sum(dev_res_acres_2014),
    units_1_acre_2014 = sum(units_1_2014) / sum(dev_res_acres_2014),
    units_2_acre_2014 = sum(units_2_2014) / sum(dev_res_acres_2014),
    units_3_4_acre_2014 = sum(units_3_4_2014) / sum(dev_res_acres_2014),
    units_5_9_acre_2014 = sum(units_5_9_2014) / sum(dev_res_acres_2014),
    units_10p_acre_2014 = sum(units_10p_2014) / sum(dev_res_acres_2014),
    pop_acre_2023 = sum(total_pop_2023) / sum(dev_res_acres_2023),
    hh_acre_2023 = sum(total_households_2023) / sum(dev_res_acres_2023),
    units_acre_2023 = sum(total_units_2023) / sum(dev_res_acres_2023),
    units_1_acre_2023 = sum(units_1_2023) / sum(dev_res_acres_2023),
    units_2_acre_2023 = sum(units_2_2023) / sum(dev_res_acres_2023),
    units_3_4_acre_2023 = sum(units_3_4_2023) / sum(dev_res_acres_2023),
    units_5_9_acre_2023 = sum(units_5_9_2023) / sum(dev_res_acres_2023),
    units_10p_acre_2023 = sum(units_10p_2023) / sum(dev_res_acres_2023),
    units_MF_acre_2014 = sum(units_2_2014 + units_3_4_2014 + units_5_9_2014 + units_10p_2014) / sum(dev_res_acres_2014),
    units_MF_acre_2023 = sum(units_2_2023 + units_3_4_2023 + units_5_9_2023 + units_10p_2023) / sum(dev_res_acres_2023)
    ) |> 
  mutate(
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    pct_chg_pop_acre = 100*(pop_acre_2023 - pop_acre_2014)/pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    pct_chg_hh_acre = 100*(hh_acre_2023 - hh_acre_2014)/hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    pct_chg_units_acre = 100*(units_acre_2023 - units_acre_2014)/units_acre_2014,
    chg_units_SF_acre = units_1_acre_2023 - units_1_acre_2014,
    pct_chg_units_SF_acre = 100*(units_1_acre_2023 - units_1_acre_2014)/units_1_acre_2014,
    chg_units_MF_acre = units_MF_acre_2023 - units_MF_acre_2014,
    pct_chg_units_MF_acre = 100*(units_MF_acre_2023 - units_MF_acre_2014)/units_MF_acre_2014,
    chg_dev_res_acres = dev_res_acres_2023 - dev_res_acres_2014,
    pct_chg_dev_res_acres = round(100*(dev_res_acres_2023 - dev_res_acres_2014)/dev_res_acres_2014,1)
    ) |> 
  rbind(met.dev.res.sw) %>%
  st_make_valid(.) %>%
  st_cast(.,'MULTIPOLYGON')

st_write(met.dev.res.mpo, dsn=task_gdb2, layer = 'pop_hh_unit_nlcd_dev_in_res_acre_mpo', append=F)
fwrite(met.dev.res.mpo |> st_drop_geometry(), paste0(deliv_path, '/pop_hh_unit_nlcd_dev_in_res_acre_mpo.csv'))

met.dev.com.sw <- met.muni |>
  summarize(
    dev_com_acres_2014 = sum(dev_com_acres_2014),
    dev_com_acres_2023 = sum(dev_com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_com_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_com_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_com_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_com_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_com_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_com_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_com_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_com_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_com_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_com_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_com_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_com_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_com_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_com_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_com_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_com_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_com_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_com_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_com_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_com_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_com_acres_2023)
  )|> 
  mutate(
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_com_acres = dev_com_acres_2023 - dev_com_acres_2014,
    pct_chg_dev_com_acres = round(100*(dev_com_acres_2023 - dev_com_acres_2014)/dev_com_acres_2014,1)
  ) |> 
  mutate(mpo='STATEWIDE')

met.dev.com.mpo <- met.muni |> 
  group_by(mpo) |>  
  summarize(
    dev_com_acres_2014 = sum(dev_com_acres_2014),
    dev_com_acres_2023 = sum(dev_com_acres_2023),
    emp_acre_2014 = sum(emp2014) / sum(dev_com_acres_2014),
    emp_2014_sect_1_acre = sum(emp_2014_sect_1) / sum(dev_com_acres_2014),
    emp_2014_sect_2_acre = sum(emp_2014_sect_2) / sum(dev_com_acres_2014),
    emp_2014_sect_3_acre = sum(emp_2014_sect_3) / sum(dev_com_acres_2014),
    emp_2014_sect_4_acre = sum(emp_2014_sect_4) / sum(dev_com_acres_2014),
    emp_2014_sect_5_acre = sum(emp_2014_sect_5) / sum(dev_com_acres_2014),
    emp_2014_sect_6_acre = sum(emp_2014_sect_6) / sum(dev_com_acres_2014),
    emp_2014_sect_7_acre = sum(emp_2014_sect_7) / sum(dev_com_acres_2014),
    emp_2014_sect_8_acre = sum(emp_2014_sect_8) / sum(dev_com_acres_2014),
    emp_2014_sect_9_acre = sum(emp_2014_sect_9) / sum(dev_com_acres_2014),
    emp_2014_sect_10_acre = sum(emp_2014_sect_10) / sum(dev_com_acres_2014),
    emp_acre_2023 = sum(emp2023) / sum(dev_com_acres_2023),
    emp_2023_sect_1_acre = sum(emp_2023_sect_1) / sum(dev_com_acres_2023),
    emp_2023_sect_2_acre = sum(emp_2023_sect_2) / sum(dev_com_acres_2023),
    emp_2023_sect_3_acre = sum(emp_2023_sect_3) / sum(dev_com_acres_2023),
    emp_2023_sect_4_acre = sum(emp_2023_sect_4) / sum(dev_com_acres_2023),
    emp_2023_sect_5_acre = sum(emp_2023_sect_5) / sum(dev_com_acres_2023),
    emp_2023_sect_6_acre = sum(emp_2023_sect_6) / sum(dev_com_acres_2023),
    emp_2023_sect_7_acre = sum(emp_2023_sect_7) / sum(dev_com_acres_2023),
    emp_2023_sect_8_acre = sum(emp_2023_sect_8) / sum(dev_com_acres_2023),
    emp_2023_sect_9_acre = sum(emp_2023_sect_9) / sum(dev_com_acres_2023),
    emp_2023_sect_10_acre = sum(emp_2023_sect_10) / sum(dev_com_acres_2023)
  )|> 
  mutate(
    chg_emp_sect_1_acre = emp_2023_sect_1_acre - emp_2014_sect_1_acre,
    chg_emp_sect_2_acre = emp_2023_sect_2_acre - emp_2014_sect_2_acre,
    chg_emp_sect_3_acre = emp_2023_sect_3_acre - emp_2014_sect_3_acre,
    chg_emp_sect_4_acre = emp_2023_sect_4_acre - emp_2014_sect_4_acre,
    chg_emp_sect_5_acre = emp_2023_sect_5_acre - emp_2014_sect_5_acre,
    chg_emp_sect_6_acre = emp_2023_sect_6_acre - emp_2014_sect_6_acre,
    chg_emp_sect_7_acre = emp_2023_sect_7_acre - emp_2014_sect_7_acre,
    chg_emp_sect_8_acre = emp_2023_sect_8_acre - emp_2014_sect_8_acre,
    chg_emp_sect_9_acre = emp_2023_sect_9_acre - emp_2014_sect_9_acre,
    chg_emp_sect_10_acre = emp_2023_sect_10_acre - emp_2014_sect_10_acre,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    pct_chg_emp_acre = 100*(emp_acre_2023 - emp_acre_2014)/emp_acre_2014,
    chg_dev_com_acres = dev_com_acres_2023 - dev_com_acres_2014,
    pct_chg_dev_com_acres = round(100*(dev_com_acres_2023 - dev_com_acres_2014)/dev_com_acres_2014,1)
  ) |> 
  rbind(met.dev.com.sw) %>%
  st_make_valid(.) %>%
  st_cast(.,'MULTIPOLYGON')

st_write(met.dev.com.mpo, dsn=task_gdb2, layer = 'emp_nlcd_dev_in_com_acre_mpo', append = F)
fwrite(met.dev.com.mpo |> st_drop_geometry(), paste0(deliv_path, '/emp_nlcd_dev_in_com_acre_mpo.csv'))

```

# 12 - Parcel LUC Changes

```{r}

p14 <- m14 |> 
  rename(MIN_LUCA_2014 = MIN_LUCA,
         MAX_LUCA_2014 = MAX_LUCA,
         LUCA_DES_2014 = LUCA_DES_L,
         LUCA_SYMB_2014 = LUCA_SYMB,
         parcel_area_2014 = Shape_Area
         ) |> 
  select(LOC_ID,MIN_LUCA_2014,MAX_LUCA_2014,LUCA_DES_2014,LUCA_SYMB_2014,parcel_area_2014) |> 
  setDT()
p22 <- m22 |> 
  rename(MIN_LUCA_2022 = MIN_LUCA,
         MAX_LUCA_2022 = MAX_LUCA,
         LUCA_DES_2022 = LUCA_DES_L,
         LUCA_SYMB_2022 = LUCA_SYMB,
         parcel_area_2022 = Shape_Area
         ) |> 
  select(LOC_ID,MIN_LUCA_2022,MAX_LUCA_2022,LUCA_DES_2022,LUCA_SYMB_2022,parcel_area_2022) |> 
  setDT()
pchg <- left_join(p14, p22, by='LOC_ID')

saveRDS(pchg,paste0(sp_path,'parcel_luc_change_2014_2022_sameLOCID.rds'))

```
