---
title: "Land Use Efficiency Metrics"
format: html
editor: visual
---

# 1 - Setup: paths and parameters

```{r, setup, include=FALSE}
options(scipen=999)

pacman::p_load(
  sf,
  data.table,
  tidycensus,
  tidyverse,
  openxlsx,
  tidygeocoder,
  terra,
  reticulate
)

# devtools::install_github('MAPC/mapcdatakeys')
library(mapcdatakeys)

reticulate::use_python("C:/Program Files/ArcGIS/Pro/bin/Python/envs/arcgispro-py3/python.exe")

# Paths
if (grepl('VM',Sys.info()[['nodename']])){
  base <- '//data-sync/Public/DataServices'
} else {
  base <- 'S:/Network Shares/K Drive/DataServices'
}


data_path <- file.path(base,'Projects/Current_Projects/Environment/EEA_MLCA/Data')
task_path = paste0(data_path,'/Task_3/')
sp_path <- paste0(task_path,'Spatial/')
task_gdb = paste0(sp_path, 'LandUse_Efficiency_Metrics.gdb')
task_gpkg = paste0(sp_path, 'LandUse_Efficiency_Metrics.gpkg')

parcel_root <- paste0(base,'/Datasets/Parcel_DB/Data/parcel_timeseries')
assess_path <- file.path(parcel_root,'_Assess_CSVs')
poly_gdb <- paste0(parcel_root,"/_Merged_polys/All_parcel_polys.gdb")
comp_gdb <- paste0(parcel_root,"/_Merged_polys/Assess_joined_parcel_polys.gdb")
delta_gdb <- paste0(parcel_root,"/_Merged_polys/Delta_parcel_polys.gdb")
year_gpkg <- paste0(parcel_root,"/_Merged_polys/parcels_by_year.gpkg")
muni_gpkg <- paste0(parcel_root,"/_Merged_polys/parcels_by_muni_timeseries.gpkg")


munis <- mapcdatakeys::all_muni_data_keys |> 
  select(muni_id,muni_name,cosub_5y14,cosub_5y23,cmsbt08,mpo) |> 
  mutate(cosub_5y14 = as.character(cosub_5y14),
         cosub_5y23 = as.character(cosub_5y23))

blk_2020 <- geog_xw_2020 |>
  select(bl20_id, muni_id) |> 
  left_join(munis, by = 'muni_id') |>
  select(bl20_id, muni_id, muni_name, cmsbt08, mpo) |>
  setDT()

knitr::opts_knit$set(root.dir = data_path)
```

# 2 - ACS data 2014

```{r}


# NEED TO SWITCH TO 2020 TRACT + MUNI GEOGRAPHIES for the 2014 data
# AND SWITCH BACK THE 2023 DATA to 2020 tracts as well
ct_xw <- fread(paste0(data_path, '/nhgis_bg2010_tr2020_25.csv')) |>
  mutate(bg2010ge = as.character(bg2010ge))


vars14 <- load_variables('acs5', year = 2014) |>
  setDT()

uv14 <- vars14[concept == 'UNITS IN STRUCTURE', name]
hv14 <- vars14[concept == 'HOUSEHOLD TYPE BY TENURE' &
                 label == 'Estimate!!Total', name]
pv14 <- vars14[concept == 'TOTAL POPULATION', name]

vs14 <- c(pv14, hv14, uv14)

# Tract-level
acs14.tract <- get_acs(
  year = 2014,
  state = 'MA',
  survey = 'acs5',
  variables = vs14,
  geography = 'block group',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'bg2010ge',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    bg2010ge,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  left_join(ct_xw, by = 'bg2010ge') |>
  mutate(
    total_pop = (total_pop * wt_pop),
    total_households = (total_households * wt_hh),
    total_units = (total_units * wt_hu),
    units_1 = (units_1 * wt_hu),
    units_2 = (units_2 * wt_hu),
    units_3_4 = (units_3_4 * wt_hu),
    units_5_9 = (units_5_9 * wt_hu),
    units_10p = (units_10p * wt_hu),
  ) |>
  rename(ct20_id = tr2020ge) |>
  group_by(ct20_id) |>
  summarize(
    total_pop = round(sum(total_pop)),
    total_households = round(sum(total_households)),
    total_units = round(sum(total_units)),
    units_1 = round(sum(units_1)),
    units_2 = round(sum(units_2)),
    units_3_4 = round(sum(units_3_4)),
    units_5_9 = round(sum(units_5_9)),
    units_10p = round(sum(units_10p))
  ) |>
  setorder(ct20_id) |>
  mutate(ct20_id = as.character(ct20_id)) |> 
  setDT()
setnames(acs14.tract, c('ct20_id',paste0(names(acs14.tract)[2:9],'_2014')))

# Muni-level
acs14.muni <- get_acs(
  year = 2014,
  state = 'MA',
  survey = 'acs5',
  variables = vs14,
  geography = 'county subdivision',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'cosub_5y14',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    cosub_5y14,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  left_join(munis, by = 'cosub_5y14') |>
  relocate(c(muni_id, muni_name, cmsbt08, mpo)) |>
  select(-c(cosub_5y14, cosub_5y23)) |>
  filter(!is.na(muni_id)) |>
  setorder(muni_id) |>
  setDT()

acs14.cmsbt <- acs14.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    total_pop_2014 = round(sum(total_pop)),
    total_households_2014 = round(sum(total_households)),
    total_units_2014 = round(sum(total_units)),
    units_1_2014 = round(sum(units_1)),
    units_2_2014 = round(sum(units_2)),
    units_3_4_2014 = round(sum(units_3_4)),
    units_5_9_2014 = round(sum(units_5_9)),
    units_10p_2014 = round(sum(units_10p))
  ) 
  
acs14.sw <- acs14.muni |> 
  summarize(
    total_pop_2014 = round(sum(total_pop)),
    total_households_2014 = round(sum(total_households)),
    total_units_2014 = round(sum(total_units)),
    units_1_2014 = round(sum(units_1)),
    units_2_2014 = round(sum(units_2)),
    units_3_4_2014 = round(sum(units_3_4)),
    units_5_9_2014 = round(sum(units_5_9)),
    units_10p_2014 = round(sum(units_10p))
  ) |> 
  mutate(mpo='STATEWIDE')

acs14.mpo <- acs14.muni |> 
  group_by(mpo) |> 
  summarize(
    total_pop_2014 = round(sum(total_pop)),
    total_households_2014 = round(sum(total_households)),
    total_units_2014 = round(sum(total_units)),
    units_1_2014 = round(sum(units_1)),
    units_2_2014 = round(sum(units_2)),
    units_3_4_2014 = round(sum(units_3_4)),
    units_5_9_2014 = round(sum(units_5_9)),
    units_10p_2014 = round(sum(units_10p))
  ) |> 
  rbind(acs14.sw)

setnames(acs14.muni, c(names(acs14.muni)[1:4],paste0(names(acs14.muni)[5:12],'_2014')))

```

# 3 - ACS data 2023

```{r}

# ct_xw <- fread(paste0(data_path, '/nhgis_bg2020_tr2010_25.csv')) |>
#   mutate(bg2020ge = as.character(bg2020ge))

vars23 <- load_variables('acs5', year = 2023) |>
  setDT()

uv23 <- vars23[concept == 'Units in Structure', name]
hv23 <- vars23[concept == 'Households by Type' &
                 label == 'Estimate!!Total:', name]
pv23 <- vars23[concept == 'Total Population', name]

vs23 <- c(pv23, hv23, uv23)

# Tract-level
acs23.tract <- get_acs(
  year = 2023,
  state = 'MA',
  survey = 'acs5',
  variables = vs23,
  geography = 'tract',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'ct20_id',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    ct20_id,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  setorder(ct20_id) |>
  mutate(ct20_id = as.character(ct20_id)) |> 
  setDT()
setnames(acs23.tract, c('ct20_id',paste0(names(acs23.tract)[2:9],'_2023')))

# Muni-level
acs23.muni <- get_acs(
  year = 2023,
  state = 'MA',
  survey = 'acs5',
  variables = vs23,
  geography = 'county subdivision',
  cache_table = TRUE
) |>
  select(-moe, -NAME) |>
  pivot_wider(id_cols = 'GEOID',
              names_from = 'variable',
              values_from = 'estimate') |>
  setnames(
    c(
      'cosub_5y23',
      'total_pop',
      'total_households',
      'total_units',
      'sf_detached',
      'sf_attached',
      'units_2',
      'units_3_4',
      'units_5_9',
      'units1019',
      'units2049',
      'units50p',
      'mobilehome',
      'RV_boat'
    )
  ) |>
  mutate(
    units_1 = sf_attached + sf_detached + mobilehome,
    units_10p = units1019 + units2049 + units50p
  ) |>
  select(
    cosub_5y23,
    total_pop,
    total_households,
    total_units,
    units_1,
    units_2,
    units_3_4,
    units_5_9,
    units_10p
  ) |>
  left_join(munis, by = 'cosub_5y23') |>
  relocate(c(muni_id, muni_name, cmsbt08, mpo)) |>
  select(-c(cosub_5y14, cosub_5y23)) |>
  filter(!is.na(muni_id)) |>
  setorder(muni_id) |>
  setDT()

acs23.cmsbt <- acs23.muni |> 
  group_by(cmsbt08) |> 
  summarize(
    total_pop_2023 = round(sum(total_pop)),
    total_households_2023 = round(sum(total_households)),
    total_units_2023 = round(sum(total_units)),
    units_1_2023 = round(sum(units_1)),
    units_2_2023 = round(sum(units_2)),
    units_3_4_2023 = round(sum(units_3_4)),
    units_5_9_2023 = round(sum(units_5_9)),
    units_10p_2023 = round(sum(units_10p))
  ) 
  
acs23.sw <- acs23.muni |> 
  summarize(
    total_pop_2023 = round(sum(total_pop)),
    total_households_2023 = round(sum(total_households)),
    total_units_2023 = round(sum(total_units)),
    units_1_2023 = round(sum(units_1)),
    units_2_2023 = round(sum(units_2)),
    units_3_4_2023 = round(sum(units_3_4)),
    units_5_9_2023 = round(sum(units_5_9)),
    units_10p_2023 = round(sum(units_10p))
  ) |> 
  mutate(mpo='STATEWIDE')

acs23.mpo <- acs23.muni |> 
  group_by(mpo) |> 
  summarize(
    total_pop_2023 = round(sum(total_pop)),
    total_households_2023 = round(sum(total_households)),
    total_units_2023 = round(sum(total_units)),
    units_1_2023 = round(sum(units_1)),
    units_2_2023 = round(sum(units_2)),
    units_3_4_2023 = round(sum(units_3_4)),
    units_5_9_2023 = round(sum(units_5_9)),
    units_10p_2023 = round(sum(units_10p))
  ) |> 
  rbind(acs23.sw)

setnames(acs23.muni, c(names(acs23.muni)[1:4],paste0(names(acs23.muni)[5:12],'_2023')))

```

# 4 - LEHD data 2014 + 2023

```{r}

emp14 <- lehdr::grab_lodes(state = 'ma',
                           year = 2014,
                           lodes_type = 'wac') |>
  mutate(
    emp_2014_sect_1 = CNS04,
    emp_2014_sect_2 = CNS15 + CNS16,
    emp_2014_sect_3 = CNS10 + CNS11,
    emp_2014_sect_4 = CNS20,
    emp_2014_sect_5 = CNS09,
    emp_2014_sect_6 = CNS07 + CNS17 + CNS18,
    emp_2014_sect_7 = CNS05,
    emp_2014_sect_8 = CNS01 + CNS02 + CNS19,
    emp_2014_sect_9 = CNS12 + CNS13 + CNS14,
    emp_2014_sect_10 = CNS03 + CNS06 + CNS08
  ) |>
  select(
    w_geocode,
    C000,
    emp_2014_sect_1,
    emp_2014_sect_2,
    emp_2014_sect_3,
    emp_2014_sect_4,
    emp_2014_sect_5,
    emp_2014_sect_6,
    emp_2014_sect_7,
    emp_2014_sect_8,
    emp_2014_sect_9,
    emp_2014_sect_10
  ) |>
  rename(bl20_id = w_geocode, emp_2014 = C000) |>
  mutate(bl20_id = as.character(bl20_id)) |>
  setDT()

emp23 <- lehdr::grab_lodes(state = 'ma',
                           year = 2023,
                           lodes_type = 'wac') |>
  mutate(
    emp_2023_sect_1 = CNS04,
    emp_2023_sect_2 = CNS15 + CNS16,
    emp_2023_sect_3 = CNS10 + CNS11,
    emp_2023_sect_4 = CNS20,
    emp_2023_sect_5 = CNS09,
    emp_2023_sect_6 = CNS07 + CNS17 + CNS18,
    emp_2023_sect_7 = CNS05,
    emp_2023_sect_8 = CNS01 + CNS02 + CNS19,
    emp_2023_sect_9 = CNS12 + CNS13 + CNS14,
    emp_2023_sect_10 = CNS03 + CNS06 + CNS08
  ) |>
  select(
    w_geocode,
    C000,
    emp_2023_sect_1,
    emp_2023_sect_2,
    emp_2023_sect_3,
    emp_2023_sect_4,
    emp_2023_sect_5,
    emp_2023_sect_6,
    emp_2023_sect_7,
    emp_2023_sect_8,
    emp_2023_sect_9,
    emp_2023_sect_10
  ) |>
  rename(bl20_id = w_geocode, emp_2023 = C000) |>
  mutate(bl20_id = as.character(bl20_id)) |> 
  setDT()


emp.blk <- blk_2020 |> 
  mutate(bl20_id = as.character(bl20_id)) |> 
  left_join(emp14, by = 'bl20_id') |> 
  left_join(emp23, by='bl20_id') |> 
  setDT()

emp.ct <- emp.blk |> 
  mutate(ct20_id = substr(bl20_id,1,11)) |> 
  group_by(ct20_id) |> 
  summarize(
    emp2014 = sum(emp_2014),
    emp2023 = sum(emp_2023),
    emp_2014_sect_1 = sum(emp_2014_sect_1),
    emp_2014_sect_2 = sum(emp_2014_sect_2),
    emp_2014_sect_3 = sum(emp_2014_sect_3),
    emp_2014_sect_4 = sum(emp_2014_sect_4),
    emp_2014_sect_5 = sum(emp_2014_sect_5),
    emp_2014_sect_6 = sum(emp_2014_sect_6),
    emp_2014_sect_7 = sum(emp_2014_sect_7),
    emp_2014_sect_8 = sum(emp_2014_sect_8),
    emp_2014_sect_9 = sum(emp_2014_sect_9),
    emp_2014_sect_10 = sum(emp_2014_sect_10),
    emp_2023_sect_1 = sum(emp_2023_sect_1),
    emp_2023_sect_2 = sum(emp_2023_sect_2),
    emp_2023_sect_3 = sum(emp_2023_sect_3),
    emp_2023_sect_4 = sum(emp_2023_sect_4),
    emp_2023_sect_5 = sum(emp_2023_sect_5),
    emp_2023_sect_6 = sum(emp_2023_sect_6),
    emp_2023_sect_7 = sum(emp_2023_sect_7),
    emp_2023_sect_8 = sum(emp_2023_sect_8),
    emp_2023_sect_9 = sum(emp_2023_sect_9),
    emp_2023_sect_10 = sum(emp_2023_sect_10)
  ) |> 
  setDT()

emp.muni <- emp.blk |> 
  group_by(muni_id,muni_name) |> 
  summarize(
    emp2014 = sum(emp_2014),
    emp2023 = sum(emp_2023),
    emp_2014_sect_1 = sum(emp_2014_sect_1),
    emp_2014_sect_2 = sum(emp_2014_sect_2),
    emp_2014_sect_3 = sum(emp_2014_sect_3),
    emp_2014_sect_4 = sum(emp_2014_sect_4),
    emp_2014_sect_5 = sum(emp_2014_sect_5),
    emp_2014_sect_6 = sum(emp_2014_sect_6),
    emp_2014_sect_7 = sum(emp_2014_sect_7),
    emp_2014_sect_8 = sum(emp_2014_sect_8),
    emp_2014_sect_9 = sum(emp_2014_sect_9),
    emp_2014_sect_10 = sum(emp_2014_sect_10),
    emp_2023_sect_1 = sum(emp_2023_sect_1),
    emp_2023_sect_2 = sum(emp_2023_sect_2),
    emp_2023_sect_3 = sum(emp_2023_sect_3),
    emp_2023_sect_4 = sum(emp_2023_sect_4),
    emp_2023_sect_5 = sum(emp_2023_sect_5),
    emp_2023_sect_6 = sum(emp_2023_sect_6),
    emp_2023_sect_7 = sum(emp_2023_sect_7),
    emp_2023_sect_8 = sum(emp_2023_sect_8),
    emp_2023_sect_9 = sum(emp_2023_sect_9),
    emp_2023_sect_10 = sum(emp_2023_sect_10)
  ) |> 
  setDT()

emp.cmsbt <- emp.blk |> 
  group_by(cmsbt08) |> 
  summarize(
    emp2014 = sum(emp_2014),
    emp2023 = sum(emp_2023),
    emp_2014_sect_1 = sum(emp_2014_sect_1),
    emp_2014_sect_2 = sum(emp_2014_sect_2),
    emp_2014_sect_3 = sum(emp_2014_sect_3),
    emp_2014_sect_4 = sum(emp_2014_sect_4),
    emp_2014_sect_5 = sum(emp_2014_sect_5),
    emp_2014_sect_6 = sum(emp_2014_sect_6),
    emp_2014_sect_7 = sum(emp_2014_sect_7),
    emp_2014_sect_8 = sum(emp_2014_sect_8),
    emp_2014_sect_9 = sum(emp_2014_sect_9),
    emp_2014_sect_10 = sum(emp_2014_sect_10),
    emp_2023_sect_1 = sum(emp_2023_sect_1),
    emp_2023_sect_2 = sum(emp_2023_sect_2),
    emp_2023_sect_3 = sum(emp_2023_sect_3),
    emp_2023_sect_4 = sum(emp_2023_sect_4),
    emp_2023_sect_5 = sum(emp_2023_sect_5),
    emp_2023_sect_6 = sum(emp_2023_sect_6),
    emp_2023_sect_7 = sum(emp_2023_sect_7),
    emp_2023_sect_8 = sum(emp_2023_sect_8),
    emp_2023_sect_9 = sum(emp_2023_sect_9),
    emp_2023_sect_10 = sum(emp_2023_sect_10)
  ) |> 
  setDT()

emp.sw <- emp.blk |> 
  summarize(
    emp2014 = sum(emp_2014),
    emp2023 = sum(emp_2023),
    emp_2014_sect_1 = sum(emp_2014_sect_1),
    emp_2014_sect_2 = sum(emp_2014_sect_2),
    emp_2014_sect_3 = sum(emp_2014_sect_3),
    emp_2014_sect_4 = sum(emp_2014_sect_4),
    emp_2014_sect_5 = sum(emp_2014_sect_5),
    emp_2014_sect_6 = sum(emp_2014_sect_6),
    emp_2014_sect_7 = sum(emp_2014_sect_7),
    emp_2014_sect_8 = sum(emp_2014_sect_8),
    emp_2014_sect_9 = sum(emp_2014_sect_9),
    emp_2014_sect_10 = sum(emp_2014_sect_10),
    emp_2023_sect_1 = sum(emp_2023_sect_1),
    emp_2023_sect_2 = sum(emp_2023_sect_2),
    emp_2023_sect_3 = sum(emp_2023_sect_3),
    emp_2023_sect_4 = sum(emp_2023_sect_4),
    emp_2023_sect_5 = sum(emp_2023_sect_5),
    emp_2023_sect_6 = sum(emp_2023_sect_6),
    emp_2023_sect_7 = sum(emp_2023_sect_7),
    emp_2023_sect_8 = sum(emp_2023_sect_8),
    emp_2023_sect_9 = sum(emp_2023_sect_9),
    emp_2023_sect_10 = sum(emp_2023_sect_10)
  ) |> 
  mutate(mpo = 'STATEWIDE') |> 
  setDT()

emp.mpo <- emp.blk |> 
  group_by(mpo) |> 
  summarize(
    emp2014 = sum(emp_2014),
    emp2023 = sum(emp_2023),
    emp_2014_sect_1 = sum(emp_2014_sect_1),
    emp_2014_sect_2 = sum(emp_2014_sect_2),
    emp_2014_sect_3 = sum(emp_2014_sect_3),
    emp_2014_sect_4 = sum(emp_2014_sect_4),
    emp_2014_sect_5 = sum(emp_2014_sect_5),
    emp_2014_sect_6 = sum(emp_2014_sect_6),
    emp_2014_sect_7 = sum(emp_2014_sect_7),
    emp_2014_sect_8 = sum(emp_2014_sect_8),
    emp_2014_sect_9 = sum(emp_2014_sect_9),
    emp_2014_sect_10 = sum(emp_2014_sect_10),
    emp_2023_sect_1 = sum(emp_2023_sect_1),
    emp_2023_sect_2 = sum(emp_2023_sect_2),
    emp_2023_sect_3 = sum(emp_2023_sect_3),
    emp_2023_sect_4 = sum(emp_2023_sect_4),
    emp_2023_sect_5 = sum(emp_2023_sect_5),
    emp_2023_sect_6 = sum(emp_2023_sect_6),
    emp_2023_sect_7 = sum(emp_2023_sect_7),
    emp_2023_sect_8 = sum(emp_2023_sect_8),
    emp_2023_sect_9 = sum(emp_2023_sect_9),
    emp_2023_sect_10 = sum(emp_2023_sect_10)
  ) |> 
  rbind(emp.sw) |> 
  setDT()




```

# 5 - NLCD Developed Area calculation

```{r}
msf <- muni_sf(2020) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,mpo,cmsbt08), by = 'muni_id')
ct20 <- tract_sf(2020)

dev2014 <- st_read(dsn = task_gdb, layer='nlcd2014_developed_area') |> 
  st_transform(crs = st_crs(msf))

dev14_ct <- st_intersection(dev2014,ct20) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev14_ct, dsn = task_gpkg, layer = 'nlcd2014_developed_tract_inter', append=F)

dev14_ct <- dev14_ct |> 
  group_by(ct20_id) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev14_ct, paste0(sp_path, 'nlcd2014_dev_area_tract.rds'))

dev14_ms <- st_intersection(dev2014,msf) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev14_ms, dsn = task_gpkg, layer = 'nlcd2014_developed_muni_inter', append=F)

dev14_muni <- dev14_ms |> 
  group_by(muni_id, muni_name) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev14_muni, paste0(sp_path, 'nlcd2014_dev_area_muni.rds'))

dev14_cmsbt <- dev14_ms |> 
  group_by(cmsbt08) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev14_cmsbt, paste0(sp_path, 'nlcd2014_dev_area_commsubtype.rds'))

dev14_sw <- dev14_ms |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  mutate(mpo='STATEWIDE')

dev14_mpo <- dev14_ms |> 
  group_by(mpo) |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  rbind(dev14_sw)
saveRDS(dev14_mpo, paste0(sp_path, 'nlcd2014_dev_area_mpo.rds'))

# 2023 NLCD

dev2023 <- st_read(dsn = task_gdb, layer='nlcd2023_developed_area') |> 
    st_transform(crs = st_crs(msf))

dev23_ct <- st_intersection(dev2023,ct20) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev23_ct, dsn = task_gpkg, layer = 'nlcd2023_developed_tract_inter', append=F)

dev23_ct <- dev23_ct |> 
  group_by(ct20_id) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev23_ct, paste0(sp_path, 'nlcd2023_dev_area_tract.rds'))

dev23_ms <- st_intersection(dev2023,msf) %>% 
  mutate(area_m2 = st_area(.))
st_write(dev23_ms, dsn = task_gpkg, layer = 'nlcd2023_developed_muni_inter', append=F)

dev23_muni <- dev23_ms |> 
  group_by(muni_id, muni_name) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev23_muni, paste0(sp_path, 'nlcd2023_dev_area_muni.rds'))

dev23_cmsbt <- dev23_ms |> 
  group_by(cmsbt08) |> 
  summarize(developed_area_m2 = sum(area_m2))
saveRDS(dev23_cmsbt, paste0(sp_path, 'nlcd2023_dev_area_commsubtype.rds'))

dev23_sw <- dev23_ms |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  mutate(mpo='STATEWIDE')

dev23_mpo <- dev23_ms |> 
  group_by(mpo) |> 
  summarize(developed_area_m2 = sum(area_m2)) |> 
  rbind(dev23_sw)
saveRDS(dev23_mpo, paste0(sp_path, 'nlcd2023_dev_area_mpo.rds'))


```

# 6 - Parcels land use codes and area

## 6a - Parcels combined for \<=2014

```{r}
msf <- mapcdatakeys::muni_sf(2020) |> 
  select(muni_name,geometry)

m14 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2014')

na14 <- m14 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2014) |> 
  select(-muni_name)

m14 <- m14 |> 
  filter(!is.na(CITY)) |> 
  rbind(na14) |> 
  mutate(SUMYEAR=2014)

city2014 <- unique(m14$CITY)

m13 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2013') 
na13 <- m13 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2013) |> 
  select(-muni_name)

m13 <- m13 |> 
  filter(!is.na(CITY)) |> 
  rbind(na13)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m13)
city2014 <- unique(m14$CITY)

m12 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2012') 
na12 <- m12 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2012) |> 
  select(-muni_name)

m12 <- m12 |> 
  filter(!is.na(CITY)) |> 
  rbind(na12)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m12)
city2014 <- unique(m14$CITY)

m11 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2011') 
na11 <- m11 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2011) |> 
  select(-muni_name)

m11 <- m11 |> 
  filter(!is.na(CITY)) |> 
  rbind(na11)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m11)
city2014 <- unique(m14$CITY)


m10 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2010') 
na10 <- m10 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2010) |> 
  select(-muni_name)

m10 <- m10 |> 
  filter(!is.na(CITY)) |> 
  rbind(na11)|> 
  filter(!CITY %in% city2014) |> 
  mutate(SUMYEAR = 2014)

m14 <- rbind(m14,m10)

# Boston and Nantucket have minimum years > 2014
m16 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2016')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na16 <- m16 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2016) |> 
  select(-muni_name)
m16 <- m16 |> 
  filter(!is.na(CITY)) |> 
  rbind(na16) |> 
  filter(CITY == 'Nantucket') |> 
  mutate(SUMYEAR=2014) |> 
  rename(SHAPE=geom)

m20 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2020')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na20 <- m20 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2020) |> 
  select(-muni_name)
m20 <- m20 |> 
  filter(!is.na(CITY)) |> 
  rbind(na20) |> 
  filter(CITY == 'Boston') |> 
  mutate(SUMYEAR=2014) |> 
  rename(SHAPE=geom)

m14 <- rbind(m14,m20,m16) %>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

st_write(m14, dsn = delta_gdb, layer = 'parcels_combine_2014', append=F)


```

## 6b - Parcels combined for \>=2022

```{r}

msf <- mapcdatakeys::muni_sf(2020) |> 
  select(muni_name,geometry)

#2022
m22 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2022')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na22 <- m22 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2022) |> 
  select(-muni_name)

m22 <- m22 |> 
  filter(!is.na(CITY)) |> 
  rbind(na22) |> 
  mutate(SUMYEAR=2022)

city2022 <- unique(m22$CITY)

#2023
m23 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2023')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na23 <- m23 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2023) |> 
  select(-muni_name)

m23 <- m23 |> 
  filter(!is.na(CITY)) |> 
  rbind(na23) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m23)
city2022 <- unique(m22$CITY)

#2024
m24 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2024')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na24 <- m24 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2024) |> 
  select(-muni_name)

m24 <- m24 |> 
  filter(!is.na(CITY)) |> 
  rbind(na24) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m24)
city2022 <- unique(m22$CITY)

#2025
m25 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2025')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na25 <- m25 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2025) |> 
  select(-muni_name)

m25 <- m25 |> 
  filter(!is.na(CITY)) |> 
  rbind(na25) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m25)
city2022 <- unique(m22$CITY)

#2026
m26 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2026')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

na26 <- m26 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2026) |> 
  select(-muni_name)

m26 <- m26 |> 
  filter(!is.na(CITY)) |> 
  rbind(na26) |> 
  filter(!CITY %in% city2022) |> 
  mutate(SUMYEAR=2022)

m22 <- rbind(m22,m26)
city2022 <- unique(m22$CITY)

# Two towns with pre-2022 data only
m19 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2019')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na19 <- m19 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2019) |> 
  select(-muni_name)
m19 <- m19 |> 
  filter(!is.na(CITY)) |> 
  rbind(na19) |> 
  filter(CITY == 'Pelham') |> 
  mutate(SUMYEAR=2022)

m18 <- st_read(dsn = year_gpkg, layer = 'muni_parcels_2018')%>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")
na18 <- m18 |> filter(is.na(CITY)) |> 
  st_join(msf, join=st_within, left = T, largest=T) |> 
  mutate(CITY = muni_name,
         YEAR = 2018) |> 
  select(-muni_name)
m18 <- m18 |> 
  filter(!is.na(CITY)) |> 
  rbind(na18) |> 
  filter(CITY == 'Chatham') |> 
  mutate(SUMYEAR=2022)

# Final compilation
m22 <- rbind(m22,m19,m18)
m22 <- m22 %>% 
  st_make_valid(.) %>%
  st_cast(., "MULTIPOLYGON")

st_write(m22, dsn = delta_gdb, layer = 'parcels_combine_2022', append=F)

```

# 7 - Calculate residential and commercial parcel area

```{r}

ct20 <- mapcdatakeys::tract_sf(2020)
msf <- mapcdatakeys::muni_sf(2020) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,cmsbt08), by = 'muni_id')

res.codes <- c('010','012','013','014','016','017','018','019','061','071','081','091',10,12,13,14,16,17,18,19,61,71,81,91,945,959,970,973,100:131,133:139,150:199)
com.codes <- c('013','014','023','030','031','032','034','035','036','037','038','039','040','041','042','043','045','046','047','048','049','063','064','073','074','083','093','094',13,14,23,30,31,32,34,35,36,37,38,39,40,41,42,43,45,46,47,48,49,63,64,73,74,83,93,94,140,220,300:393,400:442,917,927,940:943,955:957,960:969)

# 2014 layers
m14 <- st_read(dsn = delta_gdb, layer = 'parcels_combine_2014') |> 
  mutate(
    parcel_type = case_when(
      MIN_LUCA %in% as.character(res.codes) ~ 'RES',
      MIN_LUCA %in% as.character(com.codes) ~ 'COM'    
    )
  )
saveRDS(m14, paste0(sp_path, 'parcels_combine_2014.rds'))

res14 <- m14 |> 
  filter(parcel_type == 'RES') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(res14, dsn=task_gpkg, layer='residential_parcel_area_2014', append = F)

com14 <- m14 |> 
  filter(parcel_type == 'COM') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(com14, dsn=task_gpkg, layer='commercial_parcel_area_2014', append = F)

# 2022 layers
m22 <- st_read(dsn = delta_gdb, layer = 'parcels_combine_2022') |> 
  mutate(
    parcel_type = case_when(
      MIN_LUCA %in% as.character(res.codes) ~ 'RES',
      MIN_LUCA %in% as.character(com.codes) ~ 'COM'    
    )
  ) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')
saveRDS(m22, paste0(sp_path, 'parcels_combine_2022.rds'))

res22 <- m22 |> 
  filter(parcel_type == 'RES') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(res22, dsn=task_gpkg, layer='residential_parcel_area_2022', append = F)

com22 <- m22 |> 
  filter(parcel_type == 'COM') |> 
  select(LOC_ID,MIN_LUCA,MAX_LUCA,LUCA_DES_L,LUCA_SYMB,Shape_Area) |> 
  rename(parcel_area_m2 = Shape_Area)
st_write(com22, dsn=task_gpkg, layer='commercial_parcel_area_2022', append = F)

```

# 8 - Clip NLCD to res/com parcels

```{python}
import arcpy
from arcpy import env
arcpy.env.overwriteOutput = True

# Set environment settings
arcpy.env.workspace = r.task_gdb

arcpy.analysis.PairwiseClip(
    in_features="nlcd2014_developed_area",
    clip_features="commercial_parcel_area_2014",
    out_feature_class="nlcd2014_developed_in_commercial_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)

arcpy.analysis.PairwiseClip(
    in_features="nlcd2014_developed_area",
    clip_features="residential_parcel_area_2014",
    out_feature_class="nlcd2014_developed_in_residential_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)

arcpy.analysis.PairwiseClip(
    in_features="nlcd2023_developed_area",
    clip_features="commercial_parcel_area_2022",
    out_feature_class="nlcd2023_developed_in_commercial_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)

arcpy.analysis.PairwiseClip(
    in_features="nlcd2023_developed_area",
    clip_features="residential_parcel_area_2022",
    out_feature_class="nlcd2023_developed_in_residential_parcel",
    cluster_tolerance=None,
    precision="MAX_PRECISION"
)
```

# 9 - Summarize RES/COM parcel areas to Tract/Muni/Community Subtype/MPO/State
```{r}

ct20 <- mapcdatakeys::tract_sf(2020)
msf <- mapcdatakeys::muni_sf(2020) |> 
  left_join(mapcdatakeys::all_muni_data_keys |> select(muni_id,cmsbt08), by = 'muni_id')

m14 <- readRDS(paste0(sp_path, 'parcels_combine_2014.rds'))
m22 <- readRDS(paste0(sp_path, 'parcels_combine_2022.rds'))

# Tract, Muni, Community Subtype, MPO, Statewide summaries

par14_ct <- st_intersection(m14,ct20) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par14_ct, paste0(sp_path,'parcel2014_tract_inter.rds'))

par14_muni <- st_intersection(m14,msf) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par14_muni, paste0(sp_path,'parcel2014_muni_inter.rds'))

par22_ct <- st_intersection(m22,ct20) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par22_ct, paste0(sp_path,'parcel2022_tract_inter.rds'))

par22_muni <- st_intersection(m22,msf) %>%
  st_make_valid(.) %>%
  st_cast(., 'MULTIPOLYGON')%>%
  mutate(area_m2 = st_area(.))
saveRDS(par22_muni, paste0(sp_path,'parcel2022_muni_inter.rds'))


 # Sub geography summaries
ct_res14_area <- par14_ct |> 
  filter(parcel_type=='RES') |> 
  group_by(ct20_id) |> 
  summarize(res_area_m2 = sum(area_m2))
st_write(ct_res14_area, dsn=task_gpkg, append=F, layer='tract_residential_parcel_area_2014')

muni_res14_area <- par14_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(muni_id,muni_name) |> 
  summarize(res_area_m2 = sum(area_m2))
st_write(muni_res14_area, dsn=task_gpkg, append=F, layer='muni_residential_parcel_area_2014')

cmsubtyp_res14_area <- par14_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(cmsbt08) |> 
  summarize(res_area_m2 = sum(area_m2))
st_write(cmsubtyp_res14_area, dsn=task_gpkg, append=F, layer='community_subtype_residential_parcel_area_2014')

sw_res14_area <- par14_muni |> 
  filter(parcel_type=='RES')  |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_res14_area <- par14_muni |> 
  filter(parcel_type=='RES') |> 
  group_by(mpo) |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  rbind(sw_res14_area)
st_write(mpo_res14_area, dsn=task_gpkg, append=F, layer='mpo_and_statewide_residential_parcel_area_2014')

# Commercial
ct_com14_area <- par14_ct |> 
  filter(parcel_type=='COM') |> 
  group_by(ct20_id) |> 
  summarize(com_area_m2 = sum(area_m2))
st_write(ct_com14_area, dsn=task_gpkg, append=F, layer='tract_commercial_parcel_area_2014')

muni_com14_area <- par14_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(muni_id,muni_name) |> 
  summarize(com_area_m2 = sum(area_m2))
st_write(muni_com14_area, dsn=task_gpkg, append=F, layer='muni_commercial_parcel_area_2014')

cmsubtyp_com14_area <- par14_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(cmsbt08) |> 
  summarize(com_area_m2 = sum(area_m2))
st_write(cmsubtyp_com14_area, dsn=task_gpkg, append=F, layer='community_subtype_commercial_parcel_area_2014')

sw_com14_area <- par14_muni |> 
  filter(parcel_type=='COM') |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_com14_area <- par14_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(mpo) |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  rbind(sw_com14_area)
st_write(mpo_com14_area, dsn=task_gpkg, append=F, layer='mpo_and_statewide_commercial_parcel_area_2014')


# 2022 data

ct_res22_area <- par22_ct |> 
  filter(parcel_type=='RES')  |> 
  group_by(ct20_id) |> 
  summarize(res_area_m2 = sum(area_m2))
st_write(ct_res22_area, dsn=task_gpkg, append=F, layer='tract_residential_parcel_area_2022')

muni_res22_area <- par22_muni |> 
  filter(parcel_type=='RES')  |> 
  group_by(muni_id,muni_name) |> 
  summarize(res_area_m2 = sum(area_m2))
st_write(muni_res22_area, dsn=task_gpkg, append=F, layer='muni_residential_parcel_area_2022')

cmsubtyp_res22_area <- par22_muni |> 
  filter(parcel_type=='RES')  |> 
  group_by(cmsbt08) |> 
  summarize(res_area_m2 = sum(area_m2))
st_write(cmsubtyp_res22_area, dsn=task_gpkg, append=F, layer='community_subtype_residential_parcel_area_2022')

sw_res22_area <- par22_muni |> 
  filter(parcel_type=='RES')  |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_res22_area <- par22_muni |> 
  filter(parcel_type=='RES')  |> 
  group_by(mpo) |> 
  summarize(res_area_m2 = sum(area_m2)) |> 
  rbind(sw_res22_area)
st_write(mpo_res22_area, dsn=task_gpkg, append=F, layer='mpo_and_statewide_residential_parcel_area_2022')

# Commercial
ct_com22_area <- par22_ct |> 
  filter(parcel_type=='COM')  |> 
  group_by(ct20_id) |> 
  summarize(com_area_m2 = sum(area_m2))
st_write(ct_com22_area, dsn=task_gpkg, append=F, layer='tract_commercial_parcel_area_2022')

muni_com22_area <- par22_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(muni_id,muni_name) |> 
  summarize(com_area_m2 = sum(area_m2))
st_write(muni_com22_area, dsn=task_gpkg, append=F, layer='muni_commercial_parcel_area_2022')

cmsubtyp_com22_area <- par22_muni |> 
  filter(parcel_type=='COM')  |> 
  group_by(cmsbt08) |> 
  summarize(com_area_m2 = sum(area_m2))
st_write(cmsubtyp_com22_area, dsn=task_gpkg, append=F, layer='community_subtype_commercial_parcel_area_2022')

sw_com22_area <- par22_muni |> 
  filter(parcel_type=='COM')  |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  mutate(mpo = 'STATEWIDE')

mpo_com22_area <- par22_muni |> 
  filter(parcel_type=='COM') |> 
  group_by(mpo) |> 
  summarize(com_area_m2 = sum(area_m2)) |> 
  rbind(sw_com22_area)
st_write(mpo_com22_area, dsn=task_gpkg, append=F, layer='mpo_and_statewide_commercial_parcel_area_2022')


```
# 10 - Tract Metrics

```{r}

# Parcel denominators
res14.tract <- st_read(dsn=task_gpkg, layer='tract_residential_parcel_area_2014') |> 
  mutate(
    res_acres_2014 = res_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, res_acres_2014)

com14.tract <- st_read(dsn=task_gpkg, layer='tract_commercial_parcel_area_2014') |> 
  mutate(
    com_acres_2014 = com_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, com_acres_2014)

res23.tract <- st_read(dsn=task_gpkg, layer='tract_residential_parcel_area_2022') |> 
  mutate(
    res_acres_2023 = res_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, res_acres_2023)

com23.tract <- st_read(dsn=task_gpkg, layer='tract_commercial_parcel_area_2022') |> 
  mutate(
    com_acres_2023 = com_area_m2 * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, com_acres_2023)

# NLCD Developed Area denominators
nlcd14.tract <- readRDS(paste0(sp_path, 'nlcd2014_dev_area_tract.rds')) |> 
  mutate(
    dev_acres_2014 = as.numeric(developed_area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_acres_2014)

nlcd23.tract <- readRDS(paste0(sp_path, 'nlcd2023_dev_area_tract.rds')) |> 
  mutate(
    dev_acres_2023 = as.numeric(developed_area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_acres_2023)

# NLCD within parcel denominators
dev.res14.tract <- st_read(dsn=task_gdb, layer='nlcd2014_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2014 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_res_acres_2014) |> 
  group_by(ct20_id) |> 
  summarize(dev_res_acres_2014 = sum(dev_res_acres_2014))
saveRDS(dev.res14.tract,paste0(sp_path,'nlcd2014_dev_res_tract.rds'))

dev.com14.tract <- st_read(dsn=task_gdb, layer='nlcd2014_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2014 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_com_acres_2014) |> 
  group_by(ct20_id) |> 
  summarize(dev_com_acres_2014 = sum(dev_com_acres_2014))
saveRDS(dev.com14.tract,paste0(sp_path,'nlcd2014_dev_com_tract.rds'))

dev.res23.tract <- st_read(dsn=task_gdb, layer='nlcd2023_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2023 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_res_acres_2023) |> 
  group_by(ct20_id) |> 
  summarize(dev_res_acres_2023 = sum(dev_res_acres_2023))
saveRDS(dev.res23.tract,paste0(sp_path,'nlcd2023_dev_res_tract.rds'))

dev.com23.tract <- st_read(dsn=task_gdb, layer='nlcd2023_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2023 = as.numeric(area_m2) * 0.000247105,
    ct20_id = as.character(ct20_id)
    ) |>
  st_drop_geometry() |> 
  select(ct20_id, dev_com_acres_2023) |> 
  group_by(ct20_id) |> 
  summarize(dev_com_acres_2023 = sum(dev_com_acres_2023))
saveRDS(dev.com23.tract,paste0(sp_path,'nlcd2023_dev_com_tract.rds'))

# Merge numerators and calculate metrics

met.ct <- left_join(
  ct20, acs14.tract, by='ct20_id') |> 
  left_join(acs23.tract, by='ct20_id') |> 
  left_join(emp.ct, by='ct20_id') |> 
  left_join(res14.tract, by='ct20_id') |> 
  left_join(res23.tract, by='ct20_id') |> 
  left_join(com14.tract, by='ct20_id') |> 
  left_join(com23.tract, by='ct20_id') |>
  left_join(nlcd14.tract, by='ct20_id') |>
  left_join(nlcd23.tract, by='ct20_id') |>
  left_join(dev.res14.tract, by='ct20_id') |>
  left_join(dev.com14.tract, by='ct20_id') |>
  left_join(dev.res23.tract, by='ct20_id') |>
  left_join(dev.com23.tract, by='ct20_id')

met.ct[is.na(met.ct)] <- 0

# Metrics by parcel acreage of res/com
met.res.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    pop_acre_2014 = total_pop_2014 / res_acres_2014,
    hh_acre_2014 = total_households_2014 / res_acres_2014,
    units_acre_2014 = total_units_2014 / res_acres_2014,
    units_1_acre_2014 = units_1_2014 / res_acres_2014,
    units_2_acre_2014 = units_2_2014 / res_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / res_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / res_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / res_acres_2014,
    pop_acre_2023 = total_pop_2023 / res_acres_2023,
    hh_acre_2023 = total_households_2023 / res_acres_2023,
    units_acre_2023 = total_units_2023 / res_acres_2023,
    units_1_acre_2023 = units_1_2023 / res_acres_2023,
    units_2_acre_2023 = units_2_2023 / res_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / res_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / res_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / res_acres_2023,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_1_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014
    )

st_write(met.res.ct, dsn=task_gpkg, layer = 'pop_hh_unit_res_acre_tract')

met.com.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    emp_acre_2014 = emp2014 / com_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / com_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / com_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / com_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / com_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / com_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / com_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / com_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / com_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / com_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / com_acres_2014,
    emp_acre_2023 = emp2023 / com_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / com_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / com_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / com_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / com_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / com_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / com_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / com_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / com_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / com_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / com_acres_2023,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10
  )

st_write(met.com.ct, dsn=task_gpkg, layer = 'emp_com_acre_tract')


# Metrics by NLCD developed land acreage
met.dev.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    pop_acre_2014 = total_pop_2014 / dev_acres_2014,
    hh_acre_2014 = total_households_2014 / dev_acres_2014,
    units_acre_2014 = total_units_2014 / dev_acres_2014,
    units_1_acre_2014 = units_1_2014 / dev_acres_2014,
    units_2_acre_2014 = units_2_2014 / dev_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / dev_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / dev_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / dev_acres_2014,
    pop_acre_2023 = total_pop_2023 / dev_acres_2023,
    hh_acre_2023 = total_households_2023 / dev_acres_2023,
    units_acre_2023 = total_units_2023 / dev_acres_2023,
    units_1_acre_2023 = units_1_2023 / dev_acres_2023,
    units_2_acre_2023 = units_2_2023 / dev_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / dev_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / dev_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / dev_acres_2023,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_1_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014,
    emp_acre_2014 = emp2014 / dev_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / dev_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / dev_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / dev_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / dev_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / dev_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / dev_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / dev_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / dev_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / dev_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / dev_acres_2014,
    emp_acre_2023 = emp2023 / dev_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / dev_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / dev_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / dev_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / dev_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / dev_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / dev_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / dev_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / dev_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / dev_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / dev_acres_2023,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10
  )

st_write(met.dev.ct, dsn=task_gpkg, layer = 'pop_hh_units_emp_dev_acre_tract')

# Metrics by NLCD developed land IN parcel res/com acreage

met.dev.res.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    pop_acre_2014 = total_pop_2014 / res_acres_2014,
    hh_acre_2014 = total_households_2014 / res_acres_2014,
    units_acre_2014 = total_units_2014 / res_acres_2014,
    units_1_acre_2014 = units_1_2014 / res_acres_2014,
    units_2_acre_2014 = units_2_2014 / res_acres_2014,
    units_3_4_acre_2014 = units_3_4_2014 / res_acres_2014,
    units_5_9_acre_2014 = units_5_9_2014 / res_acres_2014,
    units_10p_acre_2014 = units_10p_2014 / res_acres_2014,
    pop_acre_2023 = total_pop_2023 / res_acres_2023,
    hh_acre_2023 = total_households_2023 / res_acres_2023,
    units_acre_2023 = total_units_2023 / res_acres_2023,
    units_1_acre_2023 = units_1_2023 / res_acres_2023,
    units_2_acre_2023 = units_2_2023 / res_acres_2023,
    units_3_4_acre_2023 = units_3_4_2023 / res_acres_2023,
    units_5_9_acre_2023 = units_5_9_2023 / res_acres_2023,
    units_10p_acre_2023 = units_10p_2023 / res_acres_2023,
    chg_pop_acre = pop_acre_2023 - pop_acre_2014,
    chg_hh_acre = hh_acre_2023 - hh_acre_2014,
    chg_units_acre = units_acre_2023 - units_acre_2014,
    chg_units_1_acre = units_1_acre_2023 - units_1_acre_2014,
    chg_units_2_acre_2023 = units_2_acre_2023 - units_2_acre_2014,
    chg_units_3_4_acre = units_3_4_acre_2023 - units_3_4_acre_2014,
    chg_units_5_9_acre = units_5_9_acre_2023 - units_5_9_acre_2014,
    chg_units_10p_acre = units_10p_acre_2023 - units_10p_acre_2014
    )

st_write(met.dev.res.ct, dsn=task_gpkg, layer = 'pop_hh_unit_nlcd_dev_in_res_acre_tract')

met.dev.com.ct <- met.ct |> 
  group_by(ct20_id) |> 
  summarize(
    emp_acre_2014 = emp2014 / dev_com_acres_2014,
    emp_sect_1_acre_2014 = emp_2014_sect_1 / dev_com_acres_2014,
    emp_sect_2_acre_2014 = emp_2014_sect_2 / dev_com_acres_2014,
    emp_sect_3_acre_2014 = emp_2014_sect_3 / dev_com_acres_2014,
    emp_sect_4_acre_2014 = emp_2014_sect_4 / dev_com_acres_2014,
    emp_sect_5_acre_2014 = emp_2014_sect_5 / dev_com_acres_2014,
    emp_sect_6_acre_2014 = emp_2014_sect_6 / dev_com_acres_2014,
    emp_sect_7_acre_2014 = emp_2014_sect_7 / dev_com_acres_2014,
    emp_sect_8_acre_2014 = emp_2014_sect_8 / dev_com_acres_2014,
    emp_sect_9_acre_2014 = emp_2014_sect_9 / dev_com_acres_2014,
    emp_sect_10_acre_2014 = emp_2014_sect_10 / dev_com_acres_2014,
    emp_acre_2023 = emp2023 / dev_com_acres_2023,
    emp_sect_1_acre_2023 = emp_2023_sect_1 / dev_com_acres_2023,
    emp_sect_2_acre_2023 = emp_2023_sect_2 / dev_com_acres_2023,
    emp_sect_3_acre_2023 = emp_2023_sect_3 / dev_com_acres_2023,
    emp_sect_4_acre_2023 = emp_2023_sect_4 / dev_com_acres_2023,
    emp_sect_5_acre_2023 = emp_2023_sect_5 / dev_com_acres_2023,
    emp_sect_6_acre_2023 = emp_2023_sect_6 / dev_com_acres_2023,
    emp_sect_7_acre_2023 = emp_2023_sect_7 / dev_com_acres_2023,
    emp_sect_8_acre_2023 = emp_2023_sect_8 / dev_com_acres_2023,
    emp_sect_9_acre_2023 = emp_2023_sect_9 / dev_com_acres_2023,
    emp_sect_10_acre_2023 = emp_2023_sect_10 / dev_com_acres_2023,
    chg_emp_acre = emp_acre_2023 - emp_acre_2014,
    chg_emp_sect_1_acre = emp_2023_sect_1 - emp_2014_sect_1,
    chg_emp_sect_2_acre = emp_2023_sect_2 - emp_2014_sect_2,
    chg_emp_sect_3_acre = emp_2023_sect_3 - emp_2014_sect_3,
    chg_emp_sect_4_acre = emp_2023_sect_4 - emp_2014_sect_4,
    chg_emp_sect_5_acre = emp_2023_sect_5 - emp_2014_sect_5,
    chg_emp_sect_6_acre = emp_2023_sect_6 - emp_2014_sect_6,
    chg_emp_sect_7_acre = emp_2023_sect_7 - emp_2014_sect_7,
    chg_emp_sect_8_acre = emp_2023_sect_8 - emp_2014_sect_8,
    chg_emp_sect_9_acre = emp_2023_sect_9 - emp_2014_sect_9,
    chg_emp_sect_10_acre = emp_2023_sect_10 - emp_2014_sect_10
  )

st_write(met.dev.com.ct, dsn=task_gpkg, layer = 'emp_nlcd_dev_in_com_acre_tract')


```
# 11 - Muni+ Metrics

```{r}


# Parcel denominators
res14.muni <- st_read(dsn=task_gpkg, layer='muni_residential_parcel_area_2014') |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    res_acres_2014 = res_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, res_acres_2014)

com14.muni <- st_read(dsn=task_gpkg, layer='muni_commercial_parcel_area_2014') |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    com_acres_2014 = com_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, com_acres_2014)

res23.muni <- st_read(dsn=task_gpkg, layer='muni_residential_parcel_area_2022') |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    res_acres_2023 = res_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, res_acres_2023)

com23.muni <- st_read(dsn=task_gpkg, layer='muni_commercial_parcel_area_2022') |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    com_acres_2023 = com_area_m2 * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, com_acres_2023)

# NLCD Developed Area denominators
nlcd14.muni <- readRDS(paste0(sp_path, 'nlcd2014_dev_area_muni.rds')) |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    dev_acres_2014 = as.numeric(developed_area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_acres_2014)

nlcd23.muni <- readRDS(paste0(sp_path, 'nlcd2023_dev_area_muni.rds')) |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  mutate(
    dev_acres_2023 = as.numeric(developed_area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_acres_2023)

# NLCD within parcel denominators
dev.res14.muni <- st_read(dsn=task_gdb, layer='nlcd2014_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2014 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_res_acres_2014) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_res_acres_2014 = sum(dev_res_acres_2014))
saveRDS(dev.res14.muni,paste0(sp_path,'nlcd2014_dev_res_muni.rds'))

dev.com14.muni <- st_read(dsn=task_gdb, layer='nlcd2014_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2014 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_com_acres_2014) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_com_acres_2014 = sum(dev_com_acres_2014))
saveRDS(dev.com14.muni,paste0(sp_path,'nlcd2014_dev_com_muni.rds'))

dev.res23.muni <- st_read(dsn=task_gdb, layer='nlcd2023_developed_in_residential_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_res_acres_2023 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_res_acres_2023) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_res_acres_2023 = sum(dev_res_acres_2023))
saveRDS(dev.res23.muni,paste0(sp_path,'nlcd2023_dev_res_muni.rds'))

dev.com23.muni <- st_read(dsn=task_gdb, layer='nlcd2023_developed_in_commercial_parcel') %>% 
  st_transform(crs=st_crs(ct20)) %>%
  st_intersection(ct20) %>%
  mutate(
    area_m2 = st_area(.),
    dev_com_acres_2023 = as.numeric(area_m2) * 0.000247105
    ) |>
  st_drop_geometry() |> 
  left_join(msf |> select(muni_id,mpo,cmsbt08), by = 'muni_id') |> 
  select(muni_id, muni_name, cmsbt08, mpo, dev_com_acres_2023) |> 
  group_by(muni_id, muni_name, cmsbt08, mpo) |> 
  summarize(dev_com_acres_2023 = sum(dev_com_acres_2023))
saveRDS(dev.com23.muni,paste0(sp_path,'nlcd2023_dev_com_muni.rds'))


```


# 12 - Parcel LUC Changes
```{r}

p14 <- m14 |> 
  rename(MIN_LUCA_2014 = MIN_LUCA,
         MAX_LUCA_2014 = MAX_LUCA,
         LUCA_DES_2014 = LUCA_DES_L,
         LUCA_SYMB_2014 = LUCA_SYMB,
         parcel_area_2014 = Shape_Area
         ) |> 
  select(LOC_ID,MIN_LUCA_2014,MAX_LUCA_2014,LUCA_DES_2014,LUCA_SYMB_2014,parcel_area_2014) |> 
  setDT()
p22 <- m22 |> 
  rename(MIN_LUCA_2022 = MIN_LUCA,
         MAX_LUCA_2022 = MAX_LUCA,
         LUCA_DES_2022 = LUCA_DES_L,
         LUCA_SYMB_2022 = LUCA_SYMB,
         parcel_area_2022 = Shape_Area
         ) |> 
  select(LOC_ID,MIN_LUCA_2022,MAX_LUCA_2022,LUCA_DES_2022,LUCA_SYMB_2022,parcel_area_2022) |> 
  setDT()
pchg <- left_join(p14, p22, by='LOC_ID')

saveRDS(pchg,paste0(sp_path,'parcel_luc_change_2014_2022_sameLOCID.rds'))

```
