---
title: "UrbanSim Output Processor"
author: "Conor Gately"
format: html
editor: source
---

# 1 - Setup: paths and parameters

```{r, setup, include=FALSE}

pacman::p_load(
  sf,
  data.table,
  tidycensus,
  tidyverse,
  openxlsx,
  tidygeocoder,
  lehdr,
  reticulate,
  imputeTS
)

# reticulate::use_python("C:/Program Files/ArcGIS/Pro/bin/Python/envs/arcgispro-py3/python.exe")

# devtools::install_github('MAPC/mapcdatakeys')
library(mapcdatakeys)

# Paths
if (grepl('VM',Sys.info()[['nodename']])){
  base <- '//data-sync/Public/DataServices'
} else {
  base <- 'S:/Network Shares/K Drive/DataServices'
}

data_path <- file.path(base,'Projects/Current_Projects/Environment/EEA_MLCA/Data')
shop_path <- file.path(data_path, 'Task_4/SHoP_Control_Totals')
project_gdb <- file.path(data_path, 'Spatial/EEA_MLCA.gdb')
proj_gpkg <- file.path(data_path, 'Spatial/EEA_MLCA.gpkg')
urbsim_gpkg <- file.path(data_path, 'Spatial/urbansim_geographies.gpkg')
parcel_mb_gpkg <- file.path(data_path, 'Spatial/parcels_with_massbuilds_projects.gpkg')
massbuilds_gpkg <- file.path(data_path, 'Spatial/all_massbuilds.gpkg')
task3_gdb = paste0(data_path, '/Task_3/Spatial/LandUse_Efficiency_Metrics.gdb')

task4_gdb = 'C:/Users/cgate/Desktop/EEA_Task_4/EEA_Task_4.gdb'

years=c(2019, 2024, 2029, 2034, 2039, 2049)

# Municipalities
munis <- all_muni_data_keys |>
  select(muni_id, rpa_acr, mpo) |>
  setDT()

# 2010 Blocks
blk_2010 <- geog_xw_2010 |>
  left_join(munis, by = 'muni_id') |>
  select(bl10_id, muni_id, muni_name, rpa_acr, mpo) |>
  setDT()

mapc_blks <- blk_2010 |> filter(rpa_acr == 'MAPC') |> pull(bl10_id)

blk10_sf <- mapcdatakeys::block_sf(2010) |>
  left_join(munis, by = 'muni_id')

mpos <- all_muni_data_keys |> pull(mpo) |> unique()

sun <- function(x){sort(unique(x))}
lun <- function(x){length(unique(x))}
# Existing setup (non-scenario)
# scen = 'Scen_0'
# years=c(2010, 2019, 2024, 2029, 2034)
# MAPCM_run <- 'run_228'
# SWM_run <- 'state_run_135'

# Scenario 1 (Where We Were)
# scen <- 'Scenario_1'
# MAPCM_run <- 's103'
# SWM_run <- 's62'
# ctps_run <- paste0(SWM_run,'-',MAPCM_run)

# Scenario 2 (Current Policies)
# scen <- 'Scenario_2'
# MAPCM_run <- 's104'
# SWM_run <- 's63'
# ctps_run <- paste0(SWM_run,'-',MAPCM_run)

# Scenario 3 ()
# scen <- 'Scenario_3'
# MAPCM_run <- 'run_247'
# SWM_run <- 's148'
# ctps_run <- paste0(SWM_run,'-',MAPCM_run)

# Scenario 4 ()
scen <- 'Scenario_4'
MAPCM_run <- 'run_246'
SWM_run <- 's149'
ctps_run <- paste0(SWM_run,'-',MAPCM_run)

# Conservation Test
# years <- c(2011,2012,2013,2014)
# scen='Conservation_test'
# MAPCM_run <- 'run_234'
# scen_gdb <- paste0(base, '/Projects/Current_Projects/Environment/EEA_MLCA/Data/UrbanSim/',scen,'/',scen,'.gdb')

scen_path <- file.path(data_path,'Task_4',scen,'UrbanSim_Outputs')

knitr::opts_knit$set(root.dir = data_path)
```

# 2 - Population Post-Processor

# 2.1 - Post-processor 01 - process raw microsimulated outputs

```{r}


pums13 <-
  readRDS(file.path(scen_path,'pums2013_microsim_households_short.rds'))

for (run in c(SWM_run)) {
# for (run in c(MAPCM_run)) {
  for (yr in years) {
    hh <- fread(file.path(scen_path, run, yr, 'households.csv'))
    hh[, serialno := as.character(serialno)]
    setorder(hh, serialno, block_id)
    hh[, row := .I]
    hh[, hid := paste(serialno, row, sep = '_')]
    hh <-
      hh[, .(serialno,
             block_id,
             hid,
             income,
             workers,
             persons,
             tenure,
             cars)]
    setkey(hh, serialno)
    
    # Join person-level PUMS data for each household in UrbanSim output
    hh <- merge(hh, pums13, all.x = T, allow.cartesian = T)
    hh[, is_worker := 0]
    hh[emp_status %in% c(1, 2, 3, 4, 5), is_worker := 1]
    hh[, blockgroup_id := substr(block_id, 1, 12)]
    hh[, year := yr]
    hh[is.na(tot_inc_person) | tot_inc_person < 0, tot_inc_person := 0]
    #Removed the inc_adj arithmetic, income is already in 2013USD.
    #Still reassigned the names for each variable to maintain the workflow. 
    hh[, pinc2013 := as.numeric(tot_inc_person)]
    hh[, hhinc2013 := as.numeric(hh_inc)]
    
    # Create 2010USD income variables
    # Removed the inc_adj arithmetic, income is already in 2013USD.
    hh[, inc2010 := as.integer(round(as.numeric(tot_inc_person) * (237.446 / 251.139)))]
    hh[, hhinc2010 := as.integer(round(as.numeric(hh_inc) * (237.446 / 251.139)))]
    
    swm <- hh |>
      select(
        hid,
        serialno,
        person_num,
        block_id,
        year,
        age,
        is_worker,
        relationship_to_sp1,
        pinc2013,
        hhinc2013,
        inc_adj,
        inc2010,
        hhinc2010,
        persons,
        workers,
        tenure,
        person_num
      ) |>
      mutate(
        #Create child variable
        #A child is anyone under the age of 18
        child = case_when(age < 18 ~ 1,
                          age >= 18 ~ 0),
        
        income_grp = cut(
          hhinc2013,
          breaks = c(-Inf, 35000, 75000, 125000, 225000, Inf),
          labels = c(
            "<=35000",
            "35000-75000",
            "75000-125000",
            "125000-225000",
            ">225000"
          )
        )
      ) |> setDT()
    
    # Join to muni and mpo names
    swm[, bl10_id := as.numeric(block_id)]
    setkey(swm, bl10_id)
    setkey(blk_2010, bl10_id)
    swm <- blk_2010[swm]
    
    # Count children and adults in each household
    swm[, children := sum(child, na.rm = T), hid]
    swm[child == 0, adult := 1][is.na(adult), adult := 0]
    swm[, adults := sum(adult, na.rm = T), hid]
    
    # HHtype
    swm[children > 0, HHtype := 1]
    swm[children == 0 & adults > 1, HHtype := 3]
    swm[children == 0 & adults == 1, HHtype := 5]
    
    # Persons by 5-yr age categories
    swm[age < 5 , ageCAT6 := 1]
    swm[age %in% 5:9 , ageCAT6 := 2]
    swm[age %in% 10:14 , ageCAT6 := 3]
    swm[age %in% 15:19 , ageCAT6 := 4]
    swm[age %in% 20:24 , ageCAT6 := 5]
    swm[age %in% 25:29 , ageCAT6 := 6]
    swm[age %in% 30:34 , ageCAT6 := 7]
    swm[age %in% 35:39 , ageCAT6 := 8]
    swm[age %in% 40:44 , ageCAT6 := 9]
    swm[age %in% 45:49 , ageCAT6 := 10]
    swm[age %in% 50:54 , ageCAT6 := 11]
    swm[age %in% 55:59 , ageCAT6 := 12]
    swm[age %in% 60:64 , ageCAT6 := 13]
    swm[age %in% 65:69 , ageCAT6 := 14]
    swm[age %in% 70:74 , ageCAT6 := 15]
    swm[age %in% 75:79 , ageCAT6 := 16]
    swm[age %in% 80:84 , ageCAT6 := 17]
    swm[age >= 85 , ageCAT6 := 18]
    
    # Age of HHder
    swm[person_num == 1, ageHH := age]
    swm[, ageHHder := max(ageHH, na.rm = T), hid]
    swm[, pid := paste0(hid, '_', person_num)]
    swm[, ageHH := NULL]
    
    # Count of persons by ageCAT6 in each HHd
    swm[ageCAT6 == 1, age1 := .N, hid]
    swm[ageCAT6 == 2, age2 := .N, hid]
    swm[ageCAT6 == 3, age3 := .N, hid]
    swm[ageCAT6 == 4, age4 := .N, hid]
    swm[ageCAT6 == 5, age5 := .N, hid]
    swm[ageCAT6 == 6, age6 := .N, hid]
    swm[ageCAT6 == 7, age7 := .N, hid]
    swm[ageCAT6 == 8, age8 := .N, hid]
    swm[ageCAT6 == 9, age9 := .N, hid]
    swm[ageCAT6 == 10, age10 := .N, hid]
    swm[ageCAT6 == 11, age11 := .N, hid]
    swm[ageCAT6 == 12, age12 := .N, hid]
    swm[ageCAT6 == 13, age13 := .N, hid]
    swm[ageCAT6 == 14, age14 := .N, hid]
    swm[ageCAT6 == 15, age15 := .N, hid]
    swm[ageCAT6 == 16, age16 := .N, hid]
    swm[ageCAT6 == 17, age17 := .N, hid]
    swm[ageCAT6 == 18, age18 := .N, hid]
    
    if(run==MAPCM_run){
      swm <- swm[mpo=='MAPC']
    }
    if(run==SWM_run){
      swm <- swm[mpo!='MAPC']
    }
    
    saveRDS(swm,
            paste0(scen_path, '/', run, 
              '/urbansim_',
              run,
              '_microhouseholds_',
              yr,
              '_unrevised.rds'
            ))
    message(paste0(run,' ',yr,' saved'))
  }
}

```

# 2.2 - Post-processor 02 - population difference calculation

```{r}


# 2010 SF-1 comparison data
mc <- fread(file.path(scen_path, 'pop_by_age_muni_reweighter_validate_v2.csv'))
setkey(mc, muni_id)
setkey(munis, muni_id)
mc <- munis[mc]
rmc <- mc[, lapply(.SD, sum, na.rm = T), .(mpo, ageCAT6), .SDcols = c('rw_pop', 'sf1_pop')]
mmc <- mc[, lapply(.SD, sum, na.rm = T), .(rpa_acr, ageCAT6), .SDcols =
            c('rw_pop', 'sf1_pop')]
setnames(mmc, 'rpa_acr', 'mpo')

# Load UMDI HHpop comparison data
# Rebel towns are in OCPC
for (run in c(MAPCM_run, SWM_run)) {
# for (run in c(MAPCM_run)) {
  if (run == MAPCM_run) {
    umdi <- fread(file.path(shop_path, 'pop.in.hhds.MAPC101_by.RPA.ageCAT6.csv')) |>
      select(year, rpa_acr, ageCAT6, hhpop) |>
      rename(umdi_hh_pop = hhpop, mpo = rpa_acr) |>
      setDT()
  }
  
  if (run == SWM_run) {
    umdi <- fread(file.path(shop_path, 'pop.in.hhds.MAPC97_by.RPA.ageCAT6.csv')) |>
      select(year, mpo, ageCAT6, hhpop) |>
      rename(umdi_hh_pop = hhpop) |>
      setDT()
  }
  
  umdi <-
    umdi[, lapply(.SD, sum, na.rm = T), .(year, mpo, ageCAT6), .SDcols = c('umdi_hh_pop')]
  umdi[, umdi_hh_pop := round(umdi_hh_pop)]
  umdi[year > 2010, year := year - 1]
  
  umdi <- umdi[year > 2000]
  setkey(umdi, year, mpo, ageCAT6)
  
  # Create annual data table to interpolate UMDI population by age for all years 2010-2050
  
  ufill <- data.table(
    year = sort(rep(seq(2010, 2049), 13 * 18)),
    mpo = rep(sort(rep(sun(
      umdi$mpo
    ), 18)), 40),
    ageCAT6 = rep(seq(1, 18), 13 * 40)
  )
  setkey(ufill, year, mpo, ageCAT6)
  
  umerge <- umdi[ufill]
  setorder(umerge, mpo, ageCAT6)
  
  umdi <- umerge |>
    group_by(mpo, ageCAT6) |>
    mutate(umdi_hh_pop = round(imputeTS::na_interpolation(umdi_hh_pop))) |>
    setDT()
  setkey(umdi, mpo, year, ageCAT6)
  
  fs <- list.files(
    path = file.path(scen_path, run),
    pattern = 'unrevised.rds',
    full.names = TRUE
  )
  
  if (run == SWM_run) {
    readComp <- function(x) {
      dt <- readRDS(x)
      dt[, pop := 1]
      dt[muni_id %in% c(82, 122, 231, 285), mpo := 'OCPC']
      dt <-
        dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
    }
    urbsim <- lapply(fs, readComp)
    uspop <- rbindlist(urbsim)
    uspop <- uspop[!is.na(mpo) & mpo != 'MAPC']
    setkey(uspop, mpo, year, ageCAT6)
    adiff <- umdi[uspop]
    adiff[, diff := pop - umdi_hh_pop]
    setnames(adiff, 'pop', 'urbansim_hh_pop')
    fwrite(
      adiff,
      paste0(
        scen_path,
        '/',
        run,
        '/urbansim_',
        run,
        '_umdi_pop_age_difference.csv'
      )
    )
  }
 
  if (run == MAPCM_run) {
    readComp <- function(x) {
      dt <- readRDS(x)
      dt[, pop := 1]
      dt[muni_id %in% c(82, 122, 231, 285), mpo := 'MAPC']
      dt <- dt[]
      dt <-
        dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
    }
    urbsim <- lapply(fs, readComp)
    uspop <- rbindlist(urbsim)
    uspop <- uspop[!is.na(mpo)]
    setkey(uspop, mpo, year, ageCAT6)
    adiff <- umdi[uspop]
    adiff[, diff := pop - umdi_hh_pop]
    setnames(adiff, 'pop', 'urbansim_hh_pop')
    fwrite(adiff,
           paste0(
             scen_path,'/',
             run,
             '/urbansim_',
             run,
             '_umdi_pop_age_difference.csv'
           ))
    
  }
}
```

# 2.3 - Person duplication function

```{r}
#| eval: false

# Function that creates subsets of households to sample from for replicate/delete adjustments
cx <- function(x) {
  sub1 <- x[ageCAT6 == 1  & children > 1 & ageHHder>17]
  add1 <- x[ageCAT6 == 1  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub2 <- x[ageCAT6 == 2  & children > 1 & ageHHder>17]
  add2 <- x[ageCAT6 == 2  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub3 <- x[ageCAT6 == 3  & children > 1 & ageHHder>17]
  add3 <- x[ageCAT6 == 3  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub4 <-
    x[(ageCAT6 == 4  &
         HHT == 3 &
         persons > 2 & person_num > 1) | (ageCAT6 == 4 &
                            HHT == 1 & children > 1 & person_num > 1 & ageHHder>17)]
  add4 <-
    x[(ageCAT6 == 4  &
         HHT == 1 &
         children < 4 &
         adults > 1 &
         ageHHder %in% 25:50 &
         person_num > 1) |
        (ageCAT6 == 4  & ageHHder < 44 & HHT == 3 & persons < 6 &
           person_num > 1)]
  
  sub5 <-
    x[(ageCAT6 == 5  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) | (ageCAT6 == 5 &
                              HHT == 1 & adults >= 2 &
                              person_num > 1)]
  add5 <- x[ageCAT6 == 5  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub6 <-
    x[(ageCAT6 == 6  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 6  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add6 <- x[ageCAT6 == 6  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub7 <-
    x[(ageCAT6 == 7  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 7  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add7 <- x[ageCAT6 == 7  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub8 <-
    x[(ageCAT6 == 8  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 8  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add8 <- x[ageCAT6 == 8  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 24:49]
  
  sub9 <-
    x[(ageCAT6 == 9  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) | (ageCAT6 == 9 &
                              HHT == 1 & adults >= 2 &
                              person_num > 1)]
  add9 <- x[ageCAT6 == 9  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 30:54]
  
  sub10 <-
    x[(ageCAT6 == 10 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 10 & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add10 <- x[ageCAT6 == 10 &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 40:59]
  
  sub11 <-
    x[(ageCAT6 == 11 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 11 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add11 <- x[ageCAT6 == 11 &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 40:64]
  sub12 <-
    x[(ageCAT6 == 12 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 12 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add12 <- x[ageCAT6 == 12  &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 45:69]
  sub13 <-
    x[(ageCAT6 == 13 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 13  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add13 <- x[ageCAT6 == 13  &
               HHT == 3 & persons >= 2 & ageHHder %in% 50:74]
  
  sub14 <-
    x[(ageCAT6 == 14  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 14  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add14 <- x[ageCAT6 == 14 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub15 <-
    x[(ageCAT6 == 15 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 15  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add15 <- x[ageCAT6 == 15 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub16 <-
    x[(ageCAT6 == 16 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 16 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add16 <- x[ageCAT6 == 16 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub17 <-
    x[(ageCAT6 == 17 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 17 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add17 <- x[ageCAT6 == 17 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub18 <-
    x[(ageCAT6 == 18 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 18 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add18 <- x[ageCAT6 == 18 &
               HHT > 1 & persons >= 2 & persons<8 ]
  
  add.sample <<-
    list(
      add1,
      add2,
      add3,
      add4,
      add5,
      add6,
      add7,
      add8,
      add9,
      add10,
      add11,
      add12,
      add13,
      add14,
      add15,
      add16,
      add17,
      add18
    )
  sub.sample <<-
    list(
      sub1,
      sub2,
      sub3,
      sub4,
      sub5,
      sub6,
      sub7,
      sub8,
      sub9,
      sub10,
      sub11,
      sub12,
      sub13,
      sub14,
      sub15,
      sub16,
      sub17,
      sub18
    )
  
}
```

# 2.4 - Post-processor 04 - run person duplication/subtraction

```{r}
for (run in SWM_run) {
# for (run in MAPCM_run) {
  for (yr in years) {
    flaglist <- data.table()
    
    adiff <- fread(paste0(
      scen_path,
      '/',
      run,
      '/urbansim_',
      run,
      '_umdi_pop_age_difference.csv'
    ))
    adiff[, diff := round(diff)]
    
    dt <- readRDS(
      paste0(
        scen_path,
        '/',
        run,
        '/urbansim_',
        run,
        '_microhouseholds_',
        yr,
        '_unrevised.rds'
      )
    ) |>
      mutate(child = case_when(age < 18 ~ 1, age >= 18 ~ 0)) |>
      setDT()
    
    dt[, AGEP := age]
    dt[, children := sum(child), hid]
    dt[child == 0, adult := 1]
    dt[, adults := sum(adult, na.rm = T), hid]
    dt[ageCAT6 == 1, age1 := .N, hid]
    dt[ageCAT6 == 2, age2 := .N, hid]
    dt[ageCAT6 == 3, age3 := .N, hid]
    dt[ageCAT6 == 4, age4 := .N, hid]
    dt[ageCAT6 == 5, age5 := .N, hid]
    dt[ageCAT6 == 6, age6 := .N, hid]
    dt[ageCAT6 == 7, age7 := .N, hid]
    dt[ageCAT6 == 8, age8 := .N, hid]
    dt[ageCAT6 == 9, age9 := .N, hid]
    dt[ageCAT6 == 10, age10 := .N, hid]
    dt[ageCAT6 == 11, age11 := .N, hid]
    dt[ageCAT6 == 12, age12 := .N, hid]
    dt[ageCAT6 == 13, age13 := .N, hid]
    dt[ageCAT6 == 14, age14 := .N, hid]
    dt[ageCAT6 == 15, age15 := .N, hid]
    dt[ageCAT6 == 16, age16 := .N, hid]
    dt[ageCAT6 == 17, age17 := .N, hid]
    dt[ageCAT6 == 18, age18 := .N, hid]
    dt[, SPORDER := person_num]
    dt[SPORDER == 1, ageHH := age]
    dt[, ageHHder := max(ageHH, na.rm = T), hid]
    dt[, persons := max(SPORDER), hid]
    dt[SPORDER == 1, ht := HHtype]
    dt[, HHT := max(ht, na.rm = T), hid]
    
    
    addon <- data.table()
    delflag <- data.table()
    
    if (run == MAPCM_run) {
      msub <- 'MAPC'
    }
    if (run == SWM_run) {
      msub <- mpos[mpos != 'MAPC']
    }
    for (m in msub) {
      swm <- dt[mpo == m]
      cx(swm)
      
      for (i in 1:3) {
        diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
        if (diff < 0) {
          sN <- abs(diff)
          m1 <- add.sample[[i]]
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i  & children < 4 &
                        adults >= 1 & ageHHder %in% 20:60]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i  & children <= 4]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i]
          }
          hhs <- sample(unique(m1$hid), min(lun(m1$hid), sN))
          if (lun(m1$hid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i
              )
            )
          }
          samp <- m1[hid %in% hhs & ageCAT6 == i]
          samp1 <- samp[get(paste0('age', i)) == 1]
          samp2 <-
            samp[get(paste0('age', i)) > 1, .SD[sample(.N, 1)], hid]
          addon <- rbind(addon, samp1, samp2)
        }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          m1 <- m1[!hid %in% delflag$hid]
          m1[, maxsp := max(SPORDER), hid]
          m1 <- m1[SPORDER == maxsp]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for subtraction ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for subtraction ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
        }
      }
      
      for (i in 4) {
        diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
        if (diff < 0) {
          sN <- abs(diff)
          m1 <- add.sample[[i]]
          hhs <- sample(unique(m1$hid), min(lun(m1$hid), sN))
          if (lun(m1$hid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i
              )
            )
          }
          samp1 <-
            m1[hid %in% hhs &
                 HHT == 1 & ageCAT6 == i, .SD[sample(.N, 1)], hid]
          samp1[, age := sample(15:18, 1)]
          samp2 <-
            m1[hid %in% hhs &
                 HHT == 3 &
                 ageCAT6 == i, .SD[sample(.N, 1)], hid]
          samp2[, age := 19][, ageCAT6 := 4]
          addon <- rbind(addon, samp1, samp2)
        }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          m1 <- m1[!hid %in% delflag$hid]
          m1[, maxsp := max(SPORDER), hid]
          m1 <- m1[SPORDER == maxsp]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for subtraction ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for subtraction ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
        }
      }
      for (i in 5:10) {
        diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
        if (diff < 0) {
          sN <- abs(diff)
          if (m %in% c('CCC',
                       'FRCOG',
                       'MRPC',
                       'MVC',
                       'NPEDC',
                       'NMCOG',
                       'OCPC',
                       'PVPC')) {
            m1 <- swm[ageCAT6 == i & HHT == 3]
            if (lun(m1$pid) < sN) {
              m1 <- swm[ageCAT6 == i & HHT < 5]
            }
            hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
            samp <- m1[pid %in% hhs]
            if (lun(m1$pid) < sN) {
              flaglist <-
                c(flaglist, data.table(
                  flag = paste0(
                    'WARNING - Insufficient sample size for duplication ',
                    m,
                    ' ',
                    yr,
                    ' - ',
                    'ageCAT == ',
                    i,
                    ' --- diff = ',
                    diff,
                    ' --- sample = ',
                    dim(m1)[1]
                  )
                ))
              message(
                paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              )
            }
          }
          if (!m %in% c('CCC',
                        'FRCOG',
                        'MRPC',
                        'MVC',
                        'NPEDC',
                        'NMCOG',
                        'OCPC',
                        'PVPC')) {
            m1 <- add.sample[[i]]
            hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
            if (lun(m1$pid) < sN) {
              flaglist <-
                c(flaglist, data.table(
                  flag = paste0(
                    'WARNING - Insufficient sample size for duplication ',
                    m,
                    ' ',
                    yr,
                    ' - ',
                    'ageCAT == ',
                    i,
                    ' --- diff = ',
                    diff,
                    ' --- sample = ',
                    dim(m1)[1]
                  )
                ))
              message(
                paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              )
            }
            samp <- m1[pid %in% hhs]
          }
          addon <- rbind(addon, samp)
        }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          m1 <- m1[!hid %in% delflag$hid]
          m1[, maxsp := max(SPORDER), hid]
          m1 <- m1[SPORDER == maxsp]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for subtraction ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for subtraction ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
        }
      }
      
      
      for (i in 11:18) {
        diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
        if (diff < 0) {
          sN <- abs(diff)
          m1 <- add.sample[[i]]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i & HHT == 3]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i & HHT < 5]
          }
          # if (lun(m1$pid) < sN) {
          #   m1 <- swm[ageCAT6 == i]
          # }
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          addon <- rbind(addon, samp)
        }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          m1 <- m1[!hid %in% delflag$hid]
          m1[, maxsp := max(SPORDER), hid]
          m1 <- m1[SPORDER == maxsp]
          # if (lun(m1$pid) < sN){
          #   m1 <- swm[ageCAT6 == i & person_num > 1]
          # }
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for subtraction ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for subtraction ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
        }
      }
      message(paste0(m, ' complete ', yr))
      rm(add.sample, sub.sample)
    }
    
    # Join IDs of persons to delete back to main data.table and then delete those rows
    setkey(delflag, pid)
    setkey(dt, pid)
    dt2 <- left_join(dt, delflag, by = c('pid', 'hid')) |> setDT()
    
    dt2 <- dt2[is.na(delete)]
    
    # Join person records to append to data.table
    addon[, addflag := 1]
    dt3 <- rbind(dt2, addon, fill = T)
    
    # Recalculate person and household characteristics
    setorder(dt3, hid, person_num)
    dt3[, person_num := 1:.N, hid]
    dt3[, SPORDER := person_num]
    dt3[, pid := paste0(hid, '_', person_num)]
    dt3[, persons := .N, hid]
    dt3[, workers := sum(is_worker), hid]
    dt3[pinc2013 < 0, pinc2013 := 0]
    dt3[inc2010 < 0, inc2010 := 0]
    dt3[, hhinc2013 := sum(pinc2013), hid]
    dt3[, hhinc2010 := sum(inc2010), hid]
    dt3[, HHsize := max(SPORDER), hid]
    dt3 <- dt3 |>
      mutate(
        child = case_when(age < 18 ~ 1, age >= 18 ~ 0),
        income_grp = cut(
          hhinc2013,
          breaks = c(-Inf, 35000, 75000, 125000, 225000, Inf),
          labels = c(
            "<=35000",
            "35000-75000",
            "75000-125000",
            "125000-225000",
            ">225000"
          )
        )
      ) |> setDT()
    # Count children and adults in each household
    dt3[, children_in_HH := sum(child, na.rm = T), hid]
    dt3[, children := sum(child, na.rm = T), hid]
    dt3[child == 0, adult := 1][is.na(adult), adult := 0]
    dt3[, adults := sum(adult, na.rm = T), hid]
    dt3[, size2 := HHsize]
    dt3[size2 > 7, size2 := 7]
    
    # Persons by 5-yr age categories
    dt3[age < 5 , ageCAT6 := 1]
    dt3[age %in% 5:9 , ageCAT6 := 2]
    dt3[age %in% 10:14 , ageCAT6 := 3]
    dt3[age %in% 15:19 , ageCAT6 := 4]
    dt3[age %in% 20:24 , ageCAT6 := 5]
    dt3[age %in% 25:29 , ageCAT6 := 6]
    dt3[age %in% 30:34 , ageCAT6 := 7]
    dt3[age %in% 35:39 , ageCAT6 := 8]
    dt3[age %in% 40:44 , ageCAT6 := 9]
    dt3[age %in% 45:49 , ageCAT6 := 10]
    dt3[age %in% 50:54 , ageCAT6 := 11]
    dt3[age %in% 55:59 , ageCAT6 := 12]
    dt3[age %in% 60:64 , ageCAT6 := 13]
    dt3[age %in% 65:69 , ageCAT6 := 14]
    dt3[age %in% 70:74 , ageCAT6 := 15]
    dt3[age %in% 75:79 , ageCAT6 := 16]
    dt3[age %in% 80:84 , ageCAT6 := 17]
    dt3[age >= 85 , ageCAT6 := 18]
    dt3[, HHtype := HHT]
    
    dt3 <- dt3 |> mutate(HHtype.chk = case_when(
      children >= 1 & persons > 1 ~ 1,
      children == 0 & persons > 1 ~ 3,
      persons == 1 ~ 5
    )) |> setDT()
    
    bad <- dt3[HHT != HHtype.chk]
    
    saveRDS(
      dt3,
      paste0(
        scen_path,
        '/',
        run,
        '/urbansim_',
        run,
        '_microhouseholds_',
        yr,
        '_pop_corrected.rds'
      )
    )
    
    message(paste0('Population correction complete ', run, ' ', yr))
  }
  
  fm <- data.table(unlist(flaglist))
  fm <- melt.data.table(fm)
  if (dim(fm)[1] > 0) {
    fwrite(
      fm,
      paste0(
        scen_path,
        '/',
        run,
        '/',
        run,
        '_ageCAT_with_insufficient_sample_sizes_for_adjustment.csv'
      )
    )
  }
  # Recalculate age differences to UMDI
  fs <-
    list.files(path = paste0(scen_path, '/', run),
               pattern = "pop_corrected.rds",
               full.names = TRUE)
  
  if (run == SWM_run) {
    readComp <- function(x) {
      dt <- readRDS(x)
      dt[, pop := 1]
      dt <-
        dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
    }
    urbsim <- lapply(fs, readComp)
    uspop <- rbindlist(urbsim)
    uspop <- uspop[!is.na(mpo)]
    setkey(uspop, mpo, year, ageCAT6)
    
    adiff <- umdi[uspop]
    adiff[, diff := pop - umdi_hh_pop]
    setnames(adiff, 'pop', 'urbansim_hh_pop')
    adiff <- adiff[year %in% years]
    fwrite(
      adiff,
      paste0(
        scen_path,
        '/',
        run,
        '/urbansim_',
        run,
        '_pop_corrected_umdi_pop_age_difference.csv'
      )
    )
  }
  if (run == MAPCM_run) {
    readComp <- function(x) {
      dt <- readRDS(x)
      dt[, pop := 1]
      dt <-
        dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
    }
    urbsim <- lapply(fs, readComp)
    uspop <- rbindlist(urbsim)
    uspop <- uspop[!is.na(mpo)]
    setkey(uspop, mpo, year, ageCAT6)
    
    adiff <- umdi[uspop]
    adiff[, diff := pop - umdi_hh_pop]
    setnames(adiff, 'pop', 'urbansim_hh_pop')
    adiff <- adiff[year %in% years]
    fwrite(
      adiff,
      paste0(
        scen_path,
        '/',
        run,
        '/urbansim_',
        run,
        '_pop_corrected_umdi_pop_age_difference.csv'
      )
    )
  }
}

```

# 2.5 - Post-processor 05 - CTPS microsim household export

```{r}
#| eval: false


dir <- scen_path

for (yr in years) {
  myear = yr
  syear = yr
  
  hh <- readRDS(
    paste0(
      scen_path,
      '/',
      SWM_run,
      '/urbansim_',
      SWM_run,
      '_microhouseholds_',
      syear,
      '_pop_corrected.rds'
    )
  )
  mh <- readRDS(
    paste0(
      scen_path,
      '/',
      MAPCM_run,
      '/urbansim_',
      MAPCM_run,
      '_microhouseholds_',
      myear,
      '_pop_corrected.rds'
    )
  )
  
  munis <- mapcdatakeys::all_muni_data_keys |> select(muni_id, muni_name, mpo)

  #Full list of 2010 Blocks
  keys_blocks <- mapcdatakeys::geog_xw_2010 |>
    select(bl10_id, muni_name) |>
    mutate(across(where(is.double), as.character)) |> left_join(mapcdatakeys::all_muni_data_keys, by = c("muni_name")) |>
    select(c(bl10_id, muni_name, mpo, rpa_acr))
  
  #Rebel towns block list
  rt_bl10 <- keys_blocks |>
    filter(muni_name %in% c("Duxbury", "Hanover", "Pembroke", "Stoughton")) |>
    pull(bl10_id)
  
  #MAPC 97 town Block List
  mapc97_bl10 <- keys_blocks |>
    filter(mpo == "MAPC") |>
    pull(bl10_id)
  
  #Subset MAPC to 97 towns
  mp <- mh |>
    mutate(bl10_id = as.character(bl10_id)) |>
    dplyr::filter(bl10_id %in% mapc97_bl10)
  
  #Removes MAPC97 from statewide data
  hp <- hh |>
    mutate(bl10_id = as.character(bl10_id)) |>
    filter(!bl10_id %in% mapc97_bl10)
  
  new <- rbind(mp, hp)
  new <- new[block_id != -1]
  setnames(new, 'inc2010', 'wage_inc')
  setnames(new, 'hhinc2010', 'hh_inc')
  
  if (yr < 2020) {
    fwrite(
      new,
      paste0(
        scen_path,
        '/urbansim_complete_population_',
        ctps_run,
        '_',
        yr,
        '.csv'
      )
    )
  } else {
    fwrite(
      new,
      paste0(
        scen_path,
        '/urbansim_complete_population_',
        ctps_run,
        '_',
        yr + 1,
        '.csv'
      )
    )
  }
  blk.hh <- new[, lun(hid), block_id]
  setnames(blk.hh, 'V1', 'total_households')
  setkey(blk.hh, block_id)
  
  pop_exp_df <- new[, .(
    hid,
    block_id,
    hh_inc,
    persons,
    workers,
    children,
    person_num,
    age,
    is_worker,
    wage_inc
  )]
  
  # stamp <- gsub('-', '', lubridate::today())
  
  if (yr < 2020) {
    fwrite(
      pop_exp_df,
      paste0(
        scen_path,
        '/ma_population_',
        ctps_run,
        '_',
        yr,
        '.csv'
      )
    )
  } else {
    fwrite(
      pop_exp_df,
      paste0(
        scen_path,
        '/ma_population_',
        ctps_run,
        '_',
        yr + 1,
        '.csv'
      )
    )
  }
  
  semp <- fread(paste0(scen_path, '/', SWM_run, '/', syear, '/jobs.csv'))
  sew <- dcast.data.table(semp,
                          block_id ~ aggr_sector_id,
                          value.var = 'aggr_sector_id',
                          fun.aggregate = length)
  setnames(sew, c('block_id', paste0('emp_naics_', 1:10)))
  sew[, total_emp := emp_naics_1 + emp_naics_2 + emp_naics_3 + emp_naics_4 +
        emp_naics_5 + emp_naics_6 + emp_naics_7 + emp_naics_8 + emp_naics_9 + emp_naics_10]
  
  memp <- fread(paste0(scen_path, '/', MAPCM_run, '/', myear, '/jobs.csv'))
  mew <- dcast.data.table(memp,
                          block_id ~ aggr_sector_id,
                          value.var = 'aggr_sector_id',
                          fun.aggregate = length)
  setnames(mew, c('block_id', paste0('emp_naics_', 1:10)))
  mew[, total_emp := emp_naics_1 + emp_naics_2 + emp_naics_3 + emp_naics_4 +
        emp_naics_5 + emp_naics_6 + emp_naics_7 + emp_naics_8 + emp_naics_9 + emp_naics_10]
  
  sew <- sew |>
    mutate(bl10_id = as.character(block_id)) |>
    left_join(keys_blocks, by = "bl10_id")
  
  mew <- mew |>
    mutate(bl10_id = as.character(block_id)) |>
    left_join(keys_blocks, by = "bl10_id")
  
  sew <- sew[!bl10_id %in% mapc97_bl10]
  mew <- mew[!bl10_id %in% rt_bl10]
  sew <- rbind(sew, mew)
  setnames(sew, 'total_emp', 'total_jobs')
  setkey(sew, block_id)
  setkey(blk.hh, block_id)
  sew.exp <- merge(sew, blk.hh, all = T, by = 'block_id') |>
    select(
      emp_naics_1,
      emp_naics_2,
      emp_naics_3,
      emp_naics_4,
      emp_naics_5,
      emp_naics_6,
      emp_naics_7,
      emp_naics_8,
      emp_naics_9,
      emp_naics_10,
      total_jobs,
      total_households,
      block_id
    )
  sew.exp[is.na(sew.exp)] <- 0
  
  setnames(sew.exp, (
    c(
      '1_constr',
      '2_eduhlth',
      '3_finance',
      '4_public',
      '5_info',
      '6_ret_leis',
      '7_manu',
      '8_other',
      '9_profbus',
      '10_ttu',
      'total_jobs',
      'total_households',
      'block_id'
    )
  ))
  if (yr < 2020) {
    fwrite(
      sew.exp,
      paste0(
        scen_path,
        '/',
        'ma_employment_',
        ctps_run,
        '_',
        yr,
        '.csv'
      )
    )
  } else {
    fwrite(
      sew.exp,
      paste0(
        scen_path,
        '/',
        'ma_employment_',
        ctps_run,
        '_',
        yr + 1,
        '.csv'
      )
    )
  }
  
}
  
```

# 3 - UrbanSim Block Indicator Compile

```{r}

swm.hh <- blk10_sf |> filter(!bl10_id %in% mapc_blks)

swm.blk <- blk10_sf |> filter(!bl10_id %in% mapc_blks)

dev.raw <- fread(paste0(data_path,'/Spatial/block_developable_acres.csv')) |> 
  mutate(developable_acres = SHAPE_Area * 0.000247105) |> 
  select(bl10_id, block_acres, developable_acres) |> 
  mutate(dev_pct = developable_acres / block_acres,
         bl10_id = as.character(bl10_id)) |> 
  filter(dev_pct>=.15 & !is.na(dev_pct) & !is.infinite(dev_pct)) |>
  setDT()

dev <- blk10_sf %>%
  mutate(area_m2 = st_area(.)) |> 
  mutate(acres = area_m2 * 0.000247105) |> 
  st_drop_geometry() |> 
  select(bl10_id, acres) |> 
  left_join(dev.raw, by='bl10_id') |> 
  setDT()
dev[is.na(block_acres), block_acres:=acres]
dev[is.na(developable_acres), developable_acres:=0]
dev[is.na(dev_pct), dev_pct:=0]
dev[,acres:=NULL]


y2 <- c(2010,2019,2024,2034,2049)


for (run in SWM_run) {
  for (yr in y2) {
    dt <- fread(
      paste0(
        scen_path,
        '/',
        run,
        '/Indicators/block_indicators_',
        yr,
        '.csv'
      )
    ) |> 
      filter(!block_id %in% mapc_blks) |>
      mutate(bl10_id = as.character(block_id)) |> 
      left_join(dev, by='bl10_id') |> 
      mutate(
        year = case_when(
          yr >= 2019 ~ yr + 1,
          .default = yr)
        ) |> 
      setDT()
    
    
    dt[is.na(developable_acres), developable_acres:=0]
    dt[is.na(block_acres), block_acres:=acres]
    
    dt <- dt[, .(
      year,
      bl10_id,
      block_acres,
      developable_acres,
      dev_pct,
      muni_id,
      total_households,
      hh_owners,
      hh_renters,
      total_residential_units,
      single_fam,
      multi_fam,
      own_fam,
      rent_fam,
      own_single_fam,
      own_multi_fam,
      rent_single_fam,
      rent_multi_fam,
      sum_with_child,
      median_income,
      median_rent,
      median_value,
      income_35000l,
      income_35001_75000,
      income_75001_125000,
      income_125001_200000,
      income_200001p,
      workers_0,
      workers_1,
      workers_2,
      workers_3,
      workers_4p,
      total_jobs,
      density_jobs,
      jobs_aggr_sector_1,
      jobs_aggr_sector_2,
      jobs_aggr_sector_3,
      jobs_aggr_sector_4,
      jobs_aggr_sector_5,
      jobs_aggr_sector_6,
      jobs_aggr_sector_7,
      jobs_aggr_sector_8,
      jobs_aggr_sector_9,
      jobs_aggr_sector_10
    )]
    
    setnames(
      dt,
      collapse::.c(
        year,
        bl10_id,
        block_acres,
        developable_acres,
        dev_pct,
        muni_id,
        total_households,
        owner_households,
        renter_households,
        total_residential_units,
        single_fam_units,
        multi_fam_units,
        owner_units,
        renter_units,
        owner_single_fam_units,
        owner_multi_fam_units,
        renter_single_fam_units,
        renter_multi_fam_units,
        households_with_children,
        median_household_income,
        median_household_rent,
        median_home_value,
        households_income_under35000,
        households_income_35001_75000,
        households_income_75001_125000,
        households_income_125001_200000,
        households_income_over200001,
        households_with_workers_0,
        households_with_workers_1,
        households_with_workers_2,
        households_with_workers_3,
        households_with_workers_4plus,
        total_jobs,
        jobs_per_acre,
        jobs_aggr_sector_1,
        jobs_aggr_sector_2,
        jobs_aggr_sector_3,
        jobs_aggr_sector_4,
        jobs_aggr_sector_5,
        jobs_aggr_sector_6,
        jobs_aggr_sector_7,
        jobs_aggr_sector_8,
        jobs_aggr_sector_9,
        jobs_aggr_sector_10
      )
    )

    dt[, residential_units_per_acre := round(total_residential_units / developable_acres, 3)]
    dt[, jobs_per_acre := round(total_jobs / developable_acres, 3)]
    dt[, median_household_income := round(median_household_income)]
    dt[, median_household_rent := round(median_household_rent)]
    dt[, median_home_value := round(median_home_value)]
    setorder(dt, bl10_id)
    dt[is.na(dt)] <- 0
    for (j in seq_len(ncol(dt))) {
  set(dt, which(is.infinite(dt[[j]])), j, 0)
}
    
    fwrite(dt,
           paste0(
             scen_path,
             '/',
             scen,
             '_SWM_block_indicators_',
             yr,
             '.csv'
           ))

    dt.hh <- dt |>
      select(
        bl10_id,
        total_households,
        total_residential_units,
        total_jobs,
        residential_units_per_acre,
        jobs_per_acre,
        owner_households,
        renter_households,
        single_fam_units,
        multi_fam_units,
        owner_single_fam_units,
        owner_multi_fam_units,
        renter_single_fam_units,
        renter_multi_fam_units,
        jobs_aggr_sector_1,
        jobs_aggr_sector_2,
        jobs_aggr_sector_3,
        jobs_aggr_sector_4,
        jobs_aggr_sector_5,
        jobs_aggr_sector_6,
        jobs_aggr_sector_7,
        jobs_aggr_sector_8,
        jobs_aggr_sector_9,
        jobs_aggr_sector_10
      ) |>
      setDT()

    if(yr<2019){
  yf <- yr
} else {
    yf <- yr+1
  }
setnames(dt.hh,names(dt.hh)[2:ncol(dt.hh)],paste0(names(dt.hh)[2:ncol(dt.hh)],'_',yf)) 
swm.hh <- left_join(swm.hh, dt.hh, by='bl10_id')

  }
}

# Run MAPC 101 data

mapcm.hh <- blk10_sf |> filter(bl10_id %in% mapc_blks)

sub.blk <- blk10_sf |> filter(bl10_id %in% mapc_blks)

for (run in MAPCM_run) {
  for (yr in y2) {
    dt <- fread(
      paste0(
        scen_path,
        '/',
        run,
        '/Indicators/block_indicators_',
        yr,
        '.csv'
      )
    ) |> 
      mutate(bl10_id = as.character(block_id)) |> 
      left_join(dev, by='bl10_id') |> 
      mutate(
        year = case_when(
          yr >= 2019 ~ yr + 1,
          .default = yr)
        ) |> 
      setDT()

    dt[is.na(block_acres),block_acres:=acres]
    dt[is.na(developable_acres),developable_acres:=0]
    
    dt <- dt[, .(
      year,
      bl10_id,
      block_acres,
      developable_acres,
      dev_pct,
      muni_id,
      total_households,
      hh_owners,
      hh_renters,
      total_residential_units,
      single_fam,
      multi_fam,
      own_fam,
      rent_fam,
      own_single_fam,
      own_multi_fam,
      rent_single_fam,
      rent_multi_fam,
      sum_with_child,
      median_income,
      median_rent,
      median_value,
      income_35000l,
      income_35001_75000,
      income_75001_125000,
      income_125001_200000,
      income_200001p,
      workers_0,
      workers_1,
      workers_2,
      workers_3,
      workers_4p,
      total_jobs,
      density_jobs,
      jobs_aggr_sector_1,
      jobs_aggr_sector_2,
      jobs_aggr_sector_3,
      jobs_aggr_sector_4,
      jobs_aggr_sector_5,
      jobs_aggr_sector_6,
      jobs_aggr_sector_7,
      jobs_aggr_sector_8,
      jobs_aggr_sector_9,
      jobs_aggr_sector_10
    )]
    
    setnames(
      dt,
      collapse::.c(
        year,
        bl10_id,
        block_acres,
        developable_acres,
        dev_pct,
        muni_id,
        total_households,
        owner_households,
        renter_households,
        total_residential_units,
        single_fam_units,
        multi_fam_units,
        owner_units,
        renter_units,
        owner_single_fam_units,
        owner_multi_fam_units,
        renter_single_fam_units,
        renter_multi_fam_units,
        households_with_children,
        median_household_income,
        median_household_rent,
        median_home_value,
        households_income_under35000,
        households_income_35001_75000,
        households_income_75001_125000,
        households_income_125001_200000,
        households_income_over200001,
        households_with_workers_0,
        households_with_workers_1,
        households_with_workers_2,
        households_with_workers_3,
        households_with_workers_4plus,
        total_jobs,
        jobs_per_acre,
        jobs_aggr_sector_1,
        jobs_aggr_sector_2,
        jobs_aggr_sector_3,
        jobs_aggr_sector_4,
        jobs_aggr_sector_5,
        jobs_aggr_sector_6,
        jobs_aggr_sector_7,
        jobs_aggr_sector_8,
        jobs_aggr_sector_9,
        jobs_aggr_sector_10
      )
    )
    dt[, residential_units_per_acre := round(total_residential_units / developable_acres, 3)]
    dt[, jobs_per_acre := round(total_jobs / developable_acres, 3)]
    dt[, median_household_income := round(median_household_income)]
    dt[, median_household_rent := round(median_household_rent)]
    dt[, median_home_value := round(median_home_value)]
    setorder(dt, bl10_id)
    dt[is.na(dt)] <- 0
    for (j in seq_len(ncol(dt))) {
  set(dt, which(is.infinite(dt[[j]])), j, 0)
}
    fwrite(dt,
           paste0(
             scen_path,
             '/',
             scen,
             '_MAPCM_block_indicators_',
             yr + 1,
             '.csv'
           ))

dt.hh <- dt |>
      select(
        bl10_id,
        total_households,
        total_residential_units,
        total_jobs,
        residential_units_per_acre,
        jobs_per_acre,
        owner_households,
        renter_households,
        single_fam_units,
        multi_fam_units,
        owner_single_fam_units,
        owner_multi_fam_units,
        renter_single_fam_units,
        renter_multi_fam_units,
        jobs_aggr_sector_1,
        jobs_aggr_sector_2,
        jobs_aggr_sector_3,
        jobs_aggr_sector_4,
        jobs_aggr_sector_5,
        jobs_aggr_sector_6,
        jobs_aggr_sector_7,
        jobs_aggr_sector_8,
        jobs_aggr_sector_9,
        jobs_aggr_sector_10
      ) |>
      setDT()

if(yr<2019){
  yf <- yr
} else {
    yf <- yr+1
  }
setnames(dt.hh,names(dt.hh)[2:ncol(dt.hh)],paste0(names(dt.hh)[2:ncol(dt.hh)],'_',yf)) 
mapcm.hh <- left_join(mapcm.hh, dt.hh, by='bl10_id')

  }
}

setDT(swm.hh)
setDT(mapcm.hh)

m2 <- rbind(swm.hh,mapcm.hh)
m2[is.na(m2)] <- 0
setorder(m2,bl10_id)

m2[,blockgroup2010_geoid:=substr(bl10_id,1,12)]
m2[,tract2010_geoid:=substr(bl10_id,1,11)]
setnames(m2,'bl10_id','block2010_geoid')
ads <- mapcdatakeys::all_muni_data_keys |> select(muni_id,cmsbt08)

d2 <- dev |> 
  rename(block2010_geoid = bl10_id)

m3 <- m2 |> 
  left_join(ads,by='muni_id') |> 
  left_join(d2, by='block2010_geoid') |> 
  setDT()


pop.df <- data.table()
for (yr in years) {
  if (yr<2020){
    y=yr
  } else {
    y = yr+1
  }
  pop <- fread(paste0(scen_path, '/ma_population_', ctps_run, '_', y, '.csv')) |>
    filter(person_num == 1) |>
    group_by(block_id) |>
    summarize(pop = sum(persons)) |>
    ungroup() |>
    mutate(year = y)
  pop.df <- rbind(pop.df, pop)
}

pw <- pivot_wider(pop.df, id_cols = 'block_id', names_from = 'year', names_prefix = 'pop_',values_from = 'pop') |> 
  mutate(block_id=as.character(block_id)) |> 
  setDT()
pw[is.na(pw)] <- 0
setnames(pw,'pop_2019','pop_2020')

setnames(m3,'block2010_geoid','block_id')
m4 <- m3 |> 
  st_drop_geometry() |> 
  select(-geometry) |>
  left_join(pw, by='block_id') |> 
  relocate(c(blockgroup2010_geoid, tract2010_geoid, cmsbt08),.after=block_id) |> 
  setDT()

m4[is.na(m4)] <- 0
for (j in seq_len(ncol(m4))) {
  set(m4, which(is.infinite(m4[[j]])), j, 0)
}

fwrite(m4,paste0(scen_path,'/',scen,'_full_indicators_block.csv'))


m5 <- m4 |>
  select(-dev_pct) |> 
  relocate(c(block_acres,developable_acres),.after=mpo)

cs <- names(m5)[9:ncol(m5)]

state.indicators <- m5[,lapply(.SD,sum),.SDcols=(cs)] |> 
  mutate(residential_units_per_acre_2020 = total_residential_units_2020 / developable_acres,
         residential_units_per_acre_2025 = total_residential_units_2025 / developable_acres,
         residential_units_per_acre_2035 = total_residential_units_2035 / developable_acres,
         residential_units_per_acre_2050 = total_residential_units_2050 / developable_acres,
         jobs_per_acre_2020 = total_jobs_2020 / developable_acres,
         jobs_per_acre_2025 = total_jobs_2025 / developable_acres,
         jobs_per_acre_2035 = total_jobs_2035 / developable_acres,
         jobs_per_acre_2050 = total_jobs_2050 / developable_acres,
         household_change_2025_2035 = total_households_2035 - total_households_2025,
         household_change_2025_2050 = total_households_2050 - total_households_2025,
         total_units_change_2025_2035 = total_residential_units_2035 - total_residential_units_2025,
         total_units_change_2025_2050 = total_residential_units_2050 - total_residential_units_2025,
         sf_units_change_2025_2035 = single_fam_units_2035 - single_fam_units_2025,
         sf_units_change_2025_2050 = single_fam_units_2050 - single_fam_units_2025,
         mf_units_change_2025_2035 = multi_fam_units_2035 - multi_fam_units_2025,
         mf_units_change_2025_2050 = multi_fam_units_2050 - multi_fam_units_2025,
         units_per_acre_change_2025_2035 = residential_units_per_acre_2035 - residential_units_per_acre_2025,
         units_per_acre_change_2025_2050 = residential_units_per_acre_2050 - residential_units_per_acre_2025,
         jobs_per_acre_change_2025_2035 = jobs_per_acre_2035 - jobs_per_acre_2025,
         jobs_per_acre_change_2025_2050 = jobs_per_acre_2050 - jobs_per_acre_2025,
         pop_change_2025_2035 = pop_2035 - pop_2025,
         pop_change_2025_2050 = pop_2050 - pop_2025
  ) |> 
  mutate(mpo = 'STATEWIDE') |> 
  relocate(mpo) |>
  mutate(Scenario = scen) |> 
  relocate(Scenario) |> 
  setDT()
         
muni.indicators <- m5[,lapply(.SD,sum),.(muni_id,muni_name),.SDcols=(cs)] |> 
  mutate(residential_units_per_acre_2020 = total_residential_units_2020 / developable_acres,
         residential_units_per_acre_2025 = total_residential_units_2025 / developable_acres,
         residential_units_per_acre_2035 = total_residential_units_2035 / developable_acres,
         residential_units_per_acre_2050 = total_residential_units_2050 / developable_acres,
         jobs_per_acre_2020 = total_jobs_2020 / developable_acres,
         jobs_per_acre_2025 = total_jobs_2025 / developable_acres,
         jobs_per_acre_2035 = total_jobs_2035 / developable_acres,
         jobs_per_acre_2050 = total_jobs_2050 / developable_acres,
         household_change_2025_2035 = total_households_2035 - total_households_2025,
         household_change_2025_2050 = total_households_2050 - total_households_2025,
         total_units_change_2025_2035 = total_residential_units_2035 - total_residential_units_2025,
         total_units_change_2025_2050 = total_residential_units_2050 - total_residential_units_2025,
         sf_units_change_2025_2035 = single_fam_units_2035 - single_fam_units_2025,
         sf_units_change_2025_2050 = single_fam_units_2050 - single_fam_units_2025,
         mf_units_change_2025_2035 = multi_fam_units_2035 - multi_fam_units_2025,
         mf_units_change_2025_2050 = multi_fam_units_2050 - multi_fam_units_2025,
         units_per_acre_change_2025_2035 = residential_units_per_acre_2035 - residential_units_per_acre_2025,
         units_per_acre_change_2025_2050 = residential_units_per_acre_2050 - residential_units_per_acre_2025,
         jobs_per_acre_change_2025_2035 = jobs_per_acre_2035 - jobs_per_acre_2025,
         jobs_per_acre_change_2025_2050 = jobs_per_acre_2050 - jobs_per_acre_2025,
         pop_change_2025_2035 = pop_2035 - pop_2025,
         pop_change_2025_2050 = pop_2050 - pop_2025
  ) |> 
  mutate(Scenario = scen) |> 
  relocate(Scenario) |> 
  setorder(muni_id) |> 
  setDT()

cmtype.indicators <- m5[,lapply(.SD,sum),.(cmsbt08),.SDcols=(cs)] |> 
  mutate(residential_units_per_acre_2020 = total_residential_units_2020 / developable_acres,
         residential_units_per_acre_2025 = total_residential_units_2025 / developable_acres,
         residential_units_per_acre_2035 = total_residential_units_2035 / developable_acres,
         residential_units_per_acre_2050 = total_residential_units_2050 / developable_acres,
         jobs_per_acre_2020 = total_jobs_2020 / developable_acres,
         jobs_per_acre_2025 = total_jobs_2025 / developable_acres,
         jobs_per_acre_2035 = total_jobs_2035 / developable_acres,
         jobs_per_acre_2050 = total_jobs_2050 / developable_acres,
         household_change_2025_2035 = total_households_2035 - total_households_2025,
         household_change_2025_2050 = total_households_2050 - total_households_2025,
         total_units_change_2025_2035 = total_residential_units_2035 - total_residential_units_2025,
         total_units_change_2025_2050 = total_residential_units_2050 - total_residential_units_2025,
         sf_units_change_2025_2035 = single_fam_units_2035 - single_fam_units_2025,
         sf_units_change_2025_2050 = single_fam_units_2050 - single_fam_units_2025,
         mf_units_change_2025_2035 = multi_fam_units_2035 - multi_fam_units_2025,
         mf_units_change_2025_2050 = multi_fam_units_2050 - multi_fam_units_2025,
         units_per_acre_change_2025_2035 = residential_units_per_acre_2035 - residential_units_per_acre_2025,
         units_per_acre_change_2025_2050 = residential_units_per_acre_2050 - residential_units_per_acre_2025,
         jobs_per_acre_change_2025_2035 = jobs_per_acre_2035 - jobs_per_acre_2025,
         jobs_per_acre_change_2025_2050 = jobs_per_acre_2050 - jobs_per_acre_2025,
         pop_change_2025_2035 = pop_2035 - pop_2025,
         pop_change_2025_2050 = pop_2050 - pop_2025
  ) |> 
  mutate(Scenario = scen) |> 
  relocate(Scenario) |> 
  setorder(cmsbt08) |> 
  setDT()

mpo.indicators <- m5[,lapply(.SD,sum),.(mpo),.SDcols=(cs)] |> 
  mutate(residential_units_per_acre_2020 = total_residential_units_2020 / developable_acres,
         residential_units_per_acre_2025 = total_residential_units_2025 / developable_acres,
         residential_units_per_acre_2035 = total_residential_units_2035 / developable_acres,
         residential_units_per_acre_2050 = total_residential_units_2050 / developable_acres,
         jobs_per_acre_2020 = total_jobs_2020 / developable_acres,
         jobs_per_acre_2025 = total_jobs_2025 / developable_acres,
         jobs_per_acre_2035 = total_jobs_2035 / developable_acres,
         jobs_per_acre_2050 = total_jobs_2050 / developable_acres,
         household_change_2025_2035 = total_households_2035 - total_households_2025,
         household_change_2025_2050 = total_households_2050 - total_households_2025,
         total_units_change_2025_2035 = total_residential_units_2035 - total_residential_units_2025,
         total_units_change_2025_2050 = total_residential_units_2050 - total_residential_units_2025,
         sf_units_change_2025_2035 = single_fam_units_2035 - single_fam_units_2025,
         sf_units_change_2025_2050 = single_fam_units_2050 - single_fam_units_2025,
         mf_units_change_2025_2035 = multi_fam_units_2035 - multi_fam_units_2025,
         mf_units_change_2025_2050 = multi_fam_units_2050 - multi_fam_units_2025,
         units_per_acre_change_2025_2035 = residential_units_per_acre_2035 - residential_units_per_acre_2025,
         units_per_acre_change_2025_2050 = residential_units_per_acre_2050 - residential_units_per_acre_2025,
         jobs_per_acre_change_2025_2035 = jobs_per_acre_2035 - jobs_per_acre_2025,
         jobs_per_acre_change_2025_2050 = jobs_per_acre_2050 - jobs_per_acre_2025,
         pop_change_2025_2035 = pop_2035 - pop_2025,
         pop_change_2025_2050 = pop_2050 - pop_2025
  ) |> 
  mutate(Scenario = scen) |> 
  relocate(Scenario) |> 
  setorder(mpo) |> 
  setDT()


xl.indicators <- createWorkbook()
boldStyle <- createStyle(textDecoration = "Bold")
italicStyle <- createStyle(textDecoration = "Italic")

addWorksheet(xl.indicators, sheetName = 'MPO')
addWorksheet(xl.indicators, sheetName = 'Community Type')
addWorksheet(xl.indicators, sheetName = 'Municipal')
writeData(xl.indicators, sheet = 'MPO', x = rbind(mpo.indicators,state.indicators))
writeData(xl.indicators, sheet = 'Community Type', x = cmtype.indicators)
writeData(xl.indicators, sheet = 'Municipal', x = muni.indicators)


meta <- data.table(
  employment_aggregation = seq(1, 10),
  sector_description = c(
    'Construction',
    'Education and Health Services',
    'Financial Activities',
    'Government',
    'Information',
    'Leisure and Hospitality',
    'Manufacturing',
    'Mining and Logging / Other Services',
    'Professional and Business Services',
    'Trade, Transportation, Utilities'
  )
)

addWorksheet(xl.indicators, sheetName = 'employment sector metadata')
writeData(xl.indicators, sheet = 'employment sector metadata', x = meta)

# Styles for sheet
setColWidths(
  xl.indicators,
  cols=1:ncol(meta),
  sheet = 'employment sector metadata',
  widths = "auto"
)
addStyle(
  xl.indicators,
  sheet = 'employment sector metadata',
  style = boldStyle,
  cols=1:ncol(meta),
  rows = 1,
  gridExpand = TRUE
)

setColWidths(
  xl.indicators,
  cols=1:139,
  sheet = 'MPO',
  widths = "auto"
)
setColWidths(
  xl.indicators,
  cols=1:139,
  sheet = 'Community Type',
  widths = "auto"
)
setColWidths(
  xl.indicators,
  cols=1:126,
  sheet = 'Municipal',
  widths = "auto"
)
addStyle(
  xl.indicators,
  sheet = 'MPO',
  style = boldStyle,
  cols=1:139,
  rows = 1,
  gridExpand = TRUE
)
addStyle(
  xl.indicators,
  sheet = 'Community Type',
  style = boldStyle,
  cols=1:139,
  rows = 1,
  gridExpand = TRUE
)
addStyle(
  xl.indicators,
  sheet = 'Municipal',
  style = boldStyle,
  cols=1:139,
  rows = 1,
  gridExpand = TRUE
)
addStyle(
  xl.indicators,
  sheet = 'MPO',
  style = italicStyle,
  cols=1:139,
  rows = 15,
  gridExpand = TRUE
)

saveWorkbook(xl.indicators, paste0(scen_path,'/',scen,'_urbansim_summary_indicators.xlsx'), overwrite = T)

setorder(m5,block_id)
fwrite(m5, paste0(scen_path,'/',scen,'_urbansim_block_indicators.csv'))



```

# 3.1 - Scenario 1 outputs For HF LULC Change Analysis

```{r}

s1 <- fread(paste0(data_path,'/Task_4/Scenario_1/UrbanSim_Outputs/Scenario_1_full_indicators_block.csv')) |> 
  select(block_id,
         block_acres,
         developable_acres,
         total_households_2010,
         total_households_2020,
         total_households_2025,
         total_households_2035,
         total_households_2050,
         total_residential_units_2010,
         total_residential_units_2020,
         total_residential_units_2025,
         total_residential_units_2035,
         total_residential_units_2050,
         single_fam_units_2010,
         single_fam_units_2020,
         single_fam_units_2025,
         single_fam_units_2035,
         single_fam_units_2050,
         multi_fam_units_2010,
         multi_fam_units_2020,
         multi_fam_units_2025,
         multi_fam_units_2035,
         multi_fam_units_2050,
         total_jobs_2010,
         total_jobs_2020,
         total_jobs_2025,
         total_jobs_2035,
         total_jobs_2050) |> 
  mutate(
    HH_change_2010_2020 = total_households_2020 - total_households_2010,
    HH_change_2020_2050 = total_households_2050 - total_households_2020,
    unit_change_2010_2020 = total_residential_units_2020 - total_residential_units_2010,
    unit_change_2020_2050 = total_residential_units_2050 - total_residential_units_2020,
    sf_unit_change_2010_2020 = single_fam_units_2020 - single_fam_units_2010,     
    sf_unit_change_2020_2050 = single_fam_units_2050 - single_fam_units_2020,
    mf_unit_change_2010_2020 = multi_fam_units_2020 - multi_fam_units_2010,     
    mf_unit_change_2020_2050 = multi_fam_units_2050 - multi_fam_units_2020,
    emp_change_2010_2020 = total_jobs_2020 - total_jobs_2010,     
    emp_change_2020_2050 = total_jobs_2050 - total_jobs_2020,
    scenario = 'Scenario_1',
    block_id = as.character(block_id)
         ) |> 
  relocate(scenario) |>
  right_join(blk10_sf, by=c('block_id'='bl10_id')) |> 
  relocate(c(muni_id,muni_name,rpa_acr,mpo),.after=block_id)

st_write(s1, paste0(data_path,'/Task_4/Scenario_1/UrbanSim_Outputs/Scenario_1_block_indicators.gpkg'),layer='Scenario_1_block_indicators', append=F)

ns <- names(s1)[8:ncol(s1)-1]
setDT(s1)
s1[,blockgroup_id := substr(block_id, 1, 12)]
s1.bg <- s1[,lapply(.SD,sum),blockgroup_id,.SDcols=(ns)] 
s1.bge <- blockgroup_sf(2010) |> 
  left_join(s1.bg, by=c('bg10_id'='blockgroup_id'))

st_write(s1.bge, dsn=paste0(data_path,'/Task_4/Scenario_1/UrbanSim_Outputs/Scenario_1_blockgroup_indicators.gpkg'),layer='Scenario_1_blockgroup_indicators', append=F)

```

# 3.2 - Scenario 2 outputs For HF LULC Change Analysis
```{r}
# Scenario 2

s2 <- fread(paste0(data_path,'/Task_4/Scenario_2/UrbanSim_Outputs/Scenario_2_full_indicators_block.csv')) |> 
  select(block_id,
         block_acres,
         developable_acres,
         total_households_2010,
         total_households_2020,
         total_households_2025,
         total_households_2035,
         total_households_2050,
         total_residential_units_2010,
         total_residential_units_2020,
         total_residential_units_2025,
         total_residential_units_2035,
         total_residential_units_2050,
         single_fam_units_2010,
         single_fam_units_2020,
         single_fam_units_2025,
         single_fam_units_2035,
         single_fam_units_2050,
         multi_fam_units_2010,
         multi_fam_units_2020,
         multi_fam_units_2025,
         multi_fam_units_2035,
         multi_fam_units_2050,
         total_jobs_2010,
         total_jobs_2020,
         total_jobs_2025,
         total_jobs_2035,
         total_jobs_2050) |> 
  mutate(
    HH_change_2010_2020 = total_households_2020 - total_households_2010,
    HH_change_2020_2050 = total_households_2050 - total_households_2020,
    unit_change_2010_2020 = total_residential_units_2020 - total_residential_units_2010,
    unit_change_2020_2050 = total_residential_units_2050 - total_residential_units_2020,
    sf_unit_change_2010_2020 = single_fam_units_2020 - single_fam_units_2010,     
    sf_unit_change_2020_2050 = single_fam_units_2050 - single_fam_units_2020,
    mf_unit_change_2010_2020 = multi_fam_units_2020 - multi_fam_units_2010,     
    mf_unit_change_2020_2050 = multi_fam_units_2050 - multi_fam_units_2020,
    emp_change_2010_2020 = total_jobs_2020 - total_jobs_2010,     
    emp_change_2020_2050 = total_jobs_2050 - total_jobs_2020,
    scenario = 'Scenario_2',
    block_id = as.character(block_id)
         ) |> 
  relocate(scenario) |>
  right_join(blk10_sf, by=c('block_id'='bl10_id')) |> 
  relocate(c(muni_id,muni_name,rpa_acr,mpo),.after=block_id)

st_write(s2, paste0(data_path,'/Task_4/Scenario_2/UrbanSim_Outputs/Scenario_2_block_indicators.gpkg'),layer='Scenario_2_block_indicators', append=F)

ns <- names(s2)[8:ncol(s2)-1]
setDT(s2)
s2[,blockgroup_id := substr(block_id, 1, 12)]
s2.bg <- s2[,lapply(.SD,sum),blockgroup_id,.SDcols=(ns)] 
s2.bge <- blockgroup_sf(2010) |> 
  left_join(s2.bg, by=c('bg10_id'='blockgroup_id'))

st_write(s2.bge, dsn=paste0(data_path,'/Task_4/Scenario_2/UrbanSim_Outputs/Scenario_2_blockgroup_indicators.gpkg'),layer='Scenario_2_blockgroup_indicators', append=F)

```

# 3.3 - Scenario 3 outputs For HF LULC Change Analysis
```{r}
# Scenario 3

res.sim <- fread(paste0(data_path,'/Task_4/Scenario_3/Inputs/swm_residential_development_simulated_EEA_Scenario_3_and_4.csv')) |> 
  select(block_id, residential_units)

s3 <- fread(
  paste0(
    data_path,
    '/Task_4/Scenario_3/UrbanSim_Outputs/Scenario_3_full_indicators_block.csv'
  )
) |>
  # left_join(res.sim, by = 'block_id') |>
  # _[!is.na(residential_units), `:=`(
  #   single_fam_units_2035 = single_fam_units_2035 - residential_units,
  #   multi_fam_units_2035 = multi_fam_units_2035 + residential_units,
  #   single_fam_units_2050 = single_fam_units_2050 - residential_units,
  #   multi_fam_units_2050 = multi_fam_units_2050 + residential_units
  # )] |>
  # select(-residential_units) |> 
  mutate(
    scenario = 'Scenario_3',
    block_id = as.character(block_id),
    blockgroup_id = substr(block_id, 1, 12)
  ) |>
  select(
    scenario,
    block_id,
    blockgroup_id,
    block_acres,
    developable_acres,
    total_households_2010,
    total_households_2020,
    total_households_2025,
    total_households_2035,
    total_households_2050,
    total_residential_units_2010,
    total_residential_units_2020,
    total_residential_units_2025,
    total_residential_units_2035,
    total_residential_units_2050,
    single_fam_units_2010,
    single_fam_units_2020,
    single_fam_units_2025,
    single_fam_units_2035,
    single_fam_units_2050,
    multi_fam_units_2010,
    multi_fam_units_2020,
    multi_fam_units_2025,
    multi_fam_units_2035,
    multi_fam_units_2050,
    total_jobs_2010,
    total_jobs_2020,
    total_jobs_2025,
    total_jobs_2035,
    total_jobs_2050
  ) |>
  mutate(
    HH_change_2010_2020 = total_households_2020 - total_households_2010,
    HH_change_2020_2050 = total_households_2050 - total_households_2020,
    unit_change_2010_2020 = total_residential_units_2020 - total_residential_units_2010,
    unit_change_2020_2050 = total_residential_units_2050 - total_residential_units_2020,
    sf_unit_change_2010_2020 = single_fam_units_2020 - single_fam_units_2010,
    sf_unit_change_2020_2050 = single_fam_units_2050 - single_fam_units_2020,
    mf_unit_change_2010_2020 = multi_fam_units_2020 - multi_fam_units_2010,
    mf_unit_change_2020_2050 = multi_fam_units_2050 - multi_fam_units_2020,
    emp_change_2010_2020 = total_jobs_2020 - total_jobs_2010,
    emp_change_2020_2050 = total_jobs_2050 - total_jobs_2020
  ) |>
  right_join(blk10_sf, by = c('block_id' = 'bl10_id')) |>
  relocate(muni_id, muni_name, rpa_acr, mpo, .after = block_id)

st_write(
  s3,
  dsn = paste0(scen_path, '/Scenario_3_urbansim_block_indicators.gpkg'),
  layer = 'Scenario_3_block_indicators',
  append = F
)

# st_write(s3, dsn = 'C:/Users/cgate/Desktop/Scenario_3_urbansim_outputs.gpkg', layer = 'Scenario_3_block_indicators', append=F)


ns <- names(s3)[8:(ncol(s3)-1)]

s3.bg <- s3[,lapply(.SD,sum),blockgroup_id,.SDcols=(ns)] 
s3.bg.exp <- left_join(blockgroup_sf(2010),s3.bg,by=c('bg10_id'='blockgroup_id'))

st_write(s3.bg.exp, dsn = paste0(scen_path,'/Scenario_3_urbansim_blockgroup_indicators.gpkg'), layer = 'Scenario_3_blockgroup_indicators', append=F)

```

# 3.4 - Scenario 4 outputs For HF LULC Change Analysis
```{r}

res.sim <- fread(paste0(data_path,'/Task_4/Scenario_3/Inputs/swm_residential_development_simulated_EEA_Scenario_3_and_4.csv')) |> 
  select(block_id, residential_units)

# Scenario 4

s4 <- fread(paste0(data_path,'/Task_4/Scenario_4/UrbanSim_Outputs/Scenario_4_full_indicators_block.csv')) |> 
  # left_join(res.sim, by = 'block_id') |>
  # _[!is.na(residential_units), `:=`(
  #   single_fam_units_2035 = single_fam_units_2035 - residential_units,
  #   multi_fam_units_2035 = multi_fam_units_2035 + residential_units,
  #   single_fam_units_2050 = single_fam_units_2050 - residential_units,
  #   multi_fam_units_2050 = multi_fam_units_2050 + residential_units
  # )] |>
  # select(-residential_units) |> 
  mutate(scenario = 'Scenario_4',
         block_id = as.character(block_id),
         blockgroup_id = substr(block_id, 1, 12)
         )|> 
  select(scenario,
         block_id,
         blockgroup_id,
         block_acres,
         developable_acres,
         total_households_2010,
         total_households_2020,
         total_households_2025,
         total_households_2035,
         total_households_2050,
         total_residential_units_2010,
         total_residential_units_2020,
         total_residential_units_2025,
         total_residential_units_2035,
         total_residential_units_2050,
         single_fam_units_2010,
         single_fam_units_2020,
         single_fam_units_2025,
         single_fam_units_2035,
         single_fam_units_2050,
         multi_fam_units_2010,
         multi_fam_units_2020,
         multi_fam_units_2025,
         multi_fam_units_2035,
         multi_fam_units_2050,
         total_jobs_2010,
         total_jobs_2020,
         total_jobs_2025,
         total_jobs_2035,
         total_jobs_2050) |> 
  mutate(
    HH_change_2010_2020 = total_households_2020 - total_households_2010,
    HH_change_2020_2050 = total_households_2050 - total_households_2020,
    unit_change_2010_2020 = total_residential_units_2020 - total_residential_units_2010,
    unit_change_2020_2050 = total_residential_units_2050 - total_residential_units_2020,
    sf_unit_change_2010_2020 = single_fam_units_2020 - single_fam_units_2010,     
    sf_unit_change_2020_2050 = single_fam_units_2050 - single_fam_units_2020,
    mf_unit_change_2010_2020 = multi_fam_units_2020 - multi_fam_units_2010,     
    mf_unit_change_2020_2050 = multi_fam_units_2050 - multi_fam_units_2020,
    emp_change_2010_2020 = total_jobs_2020 - total_jobs_2010,     
    emp_change_2020_2050 = total_jobs_2050 - total_jobs_2020
         ) |> 
  right_join(blk10_sf, by=c('block_id'='bl10_id')) |> 
  relocate(muni_id,muni_name,rpa_acr,mpo, .after=block_id)

st_write(s4, dsn = paste0(scen_path,'/Scenario_4_urbansim_block_indicators.gpkg'), layer = 'Scenario_4_block_indicators', append=F)

# st_write(s3, dsn = 'C:/Users/cgate/Desktop/Scenario_4_urbansim_outputs.gpkg', layer = 'Scenario_4_block_indicators', append=F)


ns <- names(s4)[8:(ncol(s4)-1)]

s4.bg <- s4[,lapply(.SD,sum),blockgroup_id,.SDcols=(ns)] 
s4.bg.exp <- left_join(blockgroup_sf(2010),s4.bg,by=c('bg10_id'='blockgroup_id'))

st_write(s4.bg.exp, dsn = paste0(scen_path,'/Scenario_4_urbansim_blockgroup_indicators.gpkg'), layer = 'Scenario_4_blockgroup_indicators', append=F)
```



# 3.5 - Scenario comparison

```{r}

rpa <- mapcdatakeys::all_muni_data_keys |> select(muni_id,mpo)

s1 <- fread(paste0(data_path,'/Task_4/Scenario_1/UrbanSim_Outputs/Scenario_1_full_indicators_block.csv')) |> 
  select(block_id,
         block_acres,
         developable_acres,
         total_households_2020,
         total_households_2050,
         total_residential_units_2020,
         total_residential_units_2050,
         single_fam_units_2020,
         single_fam_units_2050,
         multi_fam_units_2020,
         multi_fam_units_2050,
         total_jobs_2020,
         total_jobs_2050) |> 
  mutate(block_id = as.character(block_id)) 
names(s1) <- c('bl10_id','block_acres','developable_acres',paste0('scen1_',names(s1)[4:ncol(s1)]))


s2 <- fread(paste0(data_path,'/Task_4/Scenario_2/UrbanSim_Outputs/Scenario_2_full_indicators_block.csv')) |> 
  select(block_id,
         total_households_2020,
         total_households_2050,
         total_residential_units_2020,
         total_residential_units_2050,
         single_fam_units_2020,
         single_fam_units_2050,
         multi_fam_units_2020,
         multi_fam_units_2050,
         total_jobs_2020,
         total_jobs_2050) |> 
  mutate(block_id = as.character(block_id))
names(s2) <- c('bl10_id',paste0('scen2_',names(s2)[2:ncol(s2)]))

res.sim <- fread(paste0(data_path,'/Task_4/Scenario_3/Inputs/swm_residential_development_simulated_EEA_Scenario_3_and_4.csv')) |> 
  select(block_id, residential_units)

s3 <- fread(paste0(data_path,'/Task_4/Scenario_3/UrbanSim_Outputs/Scenario_3_full_indicators_block.csv')) |>
  # left_join(res.sim, by='block_id') |> 
  # _[!is.na(residential_units),`:=`(single_fam_units_2050 = single_fam_units_2050 - residential_units, multi_fam_units_2050 = multi_fam_units_2050 + residential_units)] |> 
  select(block_id,
         total_households_2020,
         total_households_2050,
         total_residential_units_2020,
         total_residential_units_2050,
         single_fam_units_2020,
         single_fam_units_2050,
         multi_fam_units_2020,
         multi_fam_units_2050,
         total_jobs_2020,
         total_jobs_2050) |> 
  mutate(block_id = as.character(block_id))
names(s3) <- c('bl10_id',paste0('scen3_',names(s3)[2:ncol(s3)]))

s4 <- fread(paste0(data_path,'/Task_4/Scenario_4/UrbanSim_Outputs/Scenario_4_full_indicators_block.csv')) |> 
  # left_join(res.sim, by='block_id') |> 
  # _[!is.na(residential_units),`:=`(single_fam_units_2050 = single_fam_units_2050 - residential_units, multi_fam_units_2050 = multi_fam_units_2050 + residential_units)] |> 
  select(block_id,
         total_households_2020,
         total_households_2050,
         total_residential_units_2020,
         total_residential_units_2050,
         single_fam_units_2020,
         single_fam_units_2050,
         multi_fam_units_2020,
         multi_fam_units_2050,
         total_jobs_2020,
         total_jobs_2050) |> 
  mutate(block_id = as.character(block_id))
names(s4) <- c('bl10_id',paste0('scen4_',names(s4)[2:ncol(s4)]))



ms <- geog_xw_2010 |> 
  select(bl10_id,muni_id,muni_name) |> 
  unique() |> 
  left_join(rpa, by='muni_id') |> 
  mutate(bl10_id = as.character(bl10_id))

comp <- block_sf(2010) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  select(bl10_id,geometry) |> 
  left_join(ms, by='bl10_id') |> 
  left_join(s1, by='bl10_id') |> 
  left_join(s2, by='bl10_id') |> 
  left_join(s3, by='bl10_id') |> 
  left_join(s4, by='bl10_id') |> 
  mutate(
    hh2020_per_acre_scen1 = scen1_total_households_2020 / developable_acres,
    hh2020_per_acre_scen2 = scen2_total_households_2020 / developable_acres,
    hh2020_per_acre_scen3 = scen3_total_households_2020 / developable_acres,
    hh2020_per_acre_scen4 = scen4_total_households_2020 / developable_acres,
    hh2050_per_acre_scen1 = scen1_total_households_2050 / developable_acres,
    hh2050_per_acre_scen2 = scen2_total_households_2050 / developable_acres,
    hh2050_per_acre_scen3 = scen3_total_households_2050 / developable_acres,
    hh2050_per_acre_scen4 = scen4_total_households_2050 / developable_acres,
    chg_hh_per_acre_scen1 = hh2050_per_acre_scen1 - hh2020_per_acre_scen1,
    chg_hh_per_acre_scen2 = hh2050_per_acre_scen2 - hh2020_per_acre_scen2,
    chg_hh_per_acre_scen3 = hh2050_per_acre_scen3 - hh2020_per_acre_scen3,
    chg_hh_per_acre_scen4 = hh2050_per_acre_scen4 - hh2020_per_acre_scen4,
    household_growth_scen1 = scen1_total_households_2050 - scen1_total_households_2020, 
    household_growth_scen2 = scen2_total_households_2050 - scen2_total_households_2020,
    household_growth_scen3 = scen3_total_households_2050 - scen3_total_households_2020,
    household_growth_scen4 = scen4_total_households_2050 - scen4_total_households_2020,
    households_diff_2050_s2_s1 = scen2_total_households_2050 - scen1_total_households_2050,
    households_diff_2050_s3_s1 = scen3_total_households_2050 - scen1_total_households_2050,
    households_diff_2050_s4_s1 = scen4_total_households_2050 - scen1_total_households_2050,
    households_diff_2050_s3_s2 = scen3_total_households_2050 - scen2_total_households_2050,
    households_diff_2050_s4_s2 = scen4_total_households_2050 - scen2_total_households_2050,
    household_growth_diff_s2_s1 = household_growth_scen2 - household_growth_scen1,
    household_growth_diff_s3_s1 = household_growth_scen3 - household_growth_scen1,
    household_growth_diff_s4_s1 = household_growth_scen4 - household_growth_scen1,
    household_growth_diff_s3_s2 = household_growth_scen3 - household_growth_scen2,
    household_growth_diff_s4_s2 = household_growth_scen4 - household_growth_scen2,
    unit2020_per_acre_scen1 = scen1_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen2 = scen2_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen3 = scen3_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen4 = scen4_total_residential_units_2020 / developable_acres,
    unit2050_per_acre_scen1 = scen1_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen2 = scen2_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen3 = scen3_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen4 = scen4_total_residential_units_2050 / developable_acres,
    chg_unit_per_acre_scen1 = unit2050_per_acre_scen1 - unit2020_per_acre_scen1,
    chg_unit_per_acre_scen2 = unit2050_per_acre_scen2 - unit2020_per_acre_scen2,
    chg_unit_per_acre_scen3 = unit2050_per_acre_scen3 - unit2020_per_acre_scen3,
    chg_unit_per_acre_scen4 = unit2050_per_acre_scen4 - unit2020_per_acre_scen4,
    unit_growth_scen1 = scen1_total_residential_units_2050 - scen1_total_residential_units_2020, 
    unit_growth_scen2 = scen2_total_residential_units_2050 - scen2_total_residential_units_2020,
    unit_growth_scen3 = scen3_total_residential_units_2050 - scen3_total_residential_units_2020,
    unit_growth_scen4 = scen4_total_residential_units_2050 - scen4_total_residential_units_2020,
    SF_unit_growth_scen1 = scen1_single_fam_units_2050 - scen1_single_fam_units_2020, 
    SF_unit_growth_scen2 = scen2_single_fam_units_2050 - scen2_single_fam_units_2020,
    SF_unit_growth_scen3 = scen3_single_fam_units_2050 - scen3_single_fam_units_2020,
    SF_unit_growth_scen4 = scen4_single_fam_units_2050 - scen4_single_fam_units_2020,
    MF_unit_growth_scen1 = scen1_multi_fam_units_2050 - scen1_multi_fam_units_2020, 
    MF_unit_growth_scen2 = scen2_multi_fam_units_2050 - scen2_multi_fam_units_2020,
    MF_unit_growth_scen3 = scen3_multi_fam_units_2050 - scen3_multi_fam_units_2020,
    MF_unit_growth_scen4 = scen4_multi_fam_units_2050 - scen4_multi_fam_units_2020,
    unit_diff_2050_s2_s1 = scen2_total_residential_units_2050 - scen1_total_residential_units_2050,
    unit_diff_2050_s3_s1 = scen3_total_residential_units_2050 - scen1_total_residential_units_2050,
    unit_diff_2050_s4_s1 = scen4_total_residential_units_2050 - scen1_total_residential_units_2050,
    unit_diff_2050_s3_s2 = scen3_total_residential_units_2050 - scen2_total_residential_units_2050,
    unit_diff_2050_s4_s2 = scen4_total_residential_units_2050 - scen2_total_residential_units_2050,
    unit_growth_diff_s2_s1 = unit_growth_scen2 - unit_growth_scen1,
    unit_growth_diff_s3_s1 = unit_growth_scen3 - unit_growth_scen1,
    unit_growth_diff_s4_s1 = unit_growth_scen4 - unit_growth_scen1,
    unit_growth_diff_s3_s2 = unit_growth_scen3 - unit_growth_scen2,
    unit_growth_diff_s4_s2 = unit_growth_scen4 - unit_growth_scen2,
    SF_unit_diff_2050_s2_s1 = scen2_single_fam_units_2050 - scen1_single_fam_units_2050,
    SF_unit_diff_2050_s3_s1 = scen3_single_fam_units_2050 - scen1_single_fam_units_2050,
    SF_unit_diff_2050_s4_s1 = scen4_single_fam_units_2050 - scen1_single_fam_units_2050,
    SF_unit_diff_2050_s3_s2 = scen3_single_fam_units_2050 - scen2_single_fam_units_2050,
    SF_unit_diff_2050_s4_s2 = scen4_single_fam_units_2050 - scen2_single_fam_units_2050,
    SF_unit_growth_diff_s2_s1 = SF_unit_growth_scen2 - SF_unit_growth_scen1,
    SF_unit_growth_diff_s3_s1 = SF_unit_growth_scen3 - SF_unit_growth_scen1,
    SF_unit_growth_diff_s4_s1 = SF_unit_growth_scen4 - SF_unit_growth_scen1,
    SF_unit_growth_diff_s3_s2 = SF_unit_growth_scen3 - SF_unit_growth_scen2,
    SF_unit_growth_diff_s4_s2 = SF_unit_growth_scen4 - SF_unit_growth_scen2,
    MF_unit_diff_2050_s2_s1 = scen2_multi_fam_units_2050 - scen1_multi_fam_units_2050,
    MF_unit_diff_2050_s3_s1 = scen3_multi_fam_units_2050 - scen1_multi_fam_units_2050,
    MF_unit_diff_2050_s4_s1 = scen4_multi_fam_units_2050 - scen1_multi_fam_units_2050,
    MF_unit_diff_2050_s3_s2 = scen3_multi_fam_units_2050 - scen2_multi_fam_units_2050,
    MF_unit_diff_2050_s4_s2 = scen4_multi_fam_units_2050 - scen2_multi_fam_units_2050,
    MF_unit_growth_diff_s2_s1 = MF_unit_growth_scen2 - MF_unit_growth_scen1,
    MF_unit_growth_diff_s3_s1 = MF_unit_growth_scen3 - MF_unit_growth_scen1,
    MF_unit_growth_diff_s4_s1 = MF_unit_growth_scen4 - MF_unit_growth_scen1,
    MF_unit_growth_diff_s3_s2 = MF_unit_growth_scen3 - MF_unit_growth_scen2,
    MF_unit_growth_diff_s4_s2 = MF_unit_growth_scen4 - MF_unit_growth_scen2,
    chg_jobs_scen1 = scen1_total_jobs_2050 - scen1_total_jobs_2020,
    chg_jobs_scen2 = scen2_total_jobs_2050 - scen2_total_jobs_2020,
    chg_jobs_scen3 = scen3_total_jobs_2050 - scen3_total_jobs_2020,
    chg_jobs_scen4 = scen4_total_jobs_2050 - scen4_total_jobs_2020
  )

st_write(comp, dsn = task4_gdb, layer = 'Task_4_Scenario_Comparisons_block', append=F)

cs <- names(comp)[7:ncol(comp)]

mc <- st_read(dsn = task4_gdb, layer = 'MassGIS_Munis_Coastal_Clip') |> 
  st_union()

comp.bg <- comp |>
  mutate(bg10_id = substr(bl10_id,1,12)) |>
  setDT() |>
  _[,lapply(.SD,sum),.(bg10_id),.SDcols=(cs)] |> 
  mutate(
    hh2020_per_acre_scen1 = scen1_total_households_2020 / developable_acres,
    hh2020_per_acre_scen2 = scen2_total_households_2020 / developable_acres,
    hh2020_per_acre_scen3 = scen3_total_households_2020 / developable_acres,
    hh2020_per_acre_scen4 = scen4_total_households_2020 / developable_acres,
    hh2050_per_acre_scen1 = scen1_total_households_2050 / developable_acres,
    hh2050_per_acre_scen2 = scen2_total_households_2050 / developable_acres,
    hh2050_per_acre_scen3 = scen3_total_households_2050 / developable_acres,
    hh2050_per_acre_scen4 = scen4_total_households_2050 / developable_acres,
    chg_hh_per_acre_scen1 = hh2050_per_acre_scen1 - hh2020_per_acre_scen1,
    chg_hh_per_acre_scen2 = hh2050_per_acre_scen2 - hh2020_per_acre_scen2,
    chg_hh_per_acre_scen3 = hh2050_per_acre_scen3 - hh2020_per_acre_scen3,
    chg_hh_per_acre_scen4 = hh2050_per_acre_scen4 - hh2020_per_acre_scen4,
    unit2020_per_acre_scen1 = scen1_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen2 = scen2_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen3 = scen3_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen4 = scen4_total_residential_units_2020 / developable_acres,
    unit2050_per_acre_scen1 = scen1_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen2 = scen2_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen3 = scen3_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen4 = scen4_total_residential_units_2050 / developable_acres,
    chg_unit_per_acre_scen1 = unit2050_per_acre_scen1 - unit2020_per_acre_scen1,
    chg_unit_per_acre_scen2 = unit2050_per_acre_scen2 - unit2020_per_acre_scen2,
    chg_unit_per_acre_scen3 = unit2050_per_acre_scen3 - unit2020_per_acre_scen3,
    chg_unit_per_acre_scen4 = unit2050_per_acre_scen4 - unit2020_per_acre_scen4
  )

bg <- blockgroup_sf(2010) |>
  st_intersection(mc) |>
  st_make_valid() |>
  st_cast('MULTIPOLYGON') |>
  mutate(bg10_id = as.character(bg10_id)) |>
  left_join(comp.bg,by='bg10_id')

st_write(bg, dsn = task4_gdb, layer = 'Task_4_Scenario_Comparisons_blockgroup', append=F)

# Tract
comp.ct <- comp |> 
  mutate(ct10_id = substr(bl10_id,1,11)) |> 
  setDT() |> 
  _[,lapply(.SD,sum),.(ct10_id),.SDcols=(cs)] |> 
  mutate(
    hh2020_per_acre_scen1 = scen1_total_households_2020 / developable_acres,
    hh2020_per_acre_scen2 = scen2_total_households_2020 / developable_acres,
    hh2020_per_acre_scen3 = scen3_total_households_2020 / developable_acres,
    hh2020_per_acre_scen4 = scen4_total_households_2020 / developable_acres,
    hh2050_per_acre_scen1 = scen1_total_households_2050 / developable_acres,
    hh2050_per_acre_scen2 = scen2_total_households_2050 / developable_acres,
    hh2050_per_acre_scen3 = scen3_total_households_2050 / developable_acres,
    hh2050_per_acre_scen4 = scen4_total_households_2050 / developable_acres,
    chg_hh_per_acre_scen1 = hh2050_per_acre_scen1 - hh2020_per_acre_scen1,
    chg_hh_per_acre_scen2 = hh2050_per_acre_scen2 - hh2020_per_acre_scen2,
    chg_hh_per_acre_scen3 = hh2050_per_acre_scen3 - hh2020_per_acre_scen3,
    chg_hh_per_acre_scen4 = hh2050_per_acre_scen4 - hh2020_per_acre_scen4,
    unit2020_per_acre_scen1 = scen1_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen2 = scen2_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen3 = scen3_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen4 = scen4_total_residential_units_2020 / developable_acres,
    unit2050_per_acre_scen1 = scen1_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen2 = scen2_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen3 = scen3_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen4 = scen4_total_residential_units_2050 / developable_acres,
    chg_unit_per_acre_scen1 = unit2050_per_acre_scen1 - unit2020_per_acre_scen1,
    chg_unit_per_acre_scen2 = unit2050_per_acre_scen2 - unit2020_per_acre_scen2,
    chg_unit_per_acre_scen3 = unit2050_per_acre_scen3 - unit2020_per_acre_scen3,
    chg_unit_per_acre_scen4 = unit2050_per_acre_scen4 - unit2020_per_acre_scen4
  )

ct <- tract_sf(2010) |> 
  st_intersection(mc) |> 
  st_make_valid() |> 
  st_cast('MULTIPOLYGON') |> 
  mutate(ct10_id = as.character(ct10_id)) |> 
  left_join(comp.ct, by='ct10_id')

st_write(ct, dsn = task4_gdb, layer = 'Task_4_Scenario_Comparisons_tract', append=F)

cmt <- mapcdatakeys::all_muni_data_keys |> 
  select(muni_id,cmsbt08)

cmt.sf <- st_read(dsn = task4_gdb, layer = 'MassGIS_Munis_Coastal_Clip') |>
  select(TOWN_ID,TOWN,Shape) |> 
  rename(muni_id=TOWN_ID, muni_name=TOWN) |> 
  left_join(cmt, by='muni_id')
st_write(cmt.sf, dsn = task4_gdb, layer='muni_cmsbt_coast_clip', append=F)

mun <- comp |> 
  left_join(cmt, by='muni_id') |> 
  setDT() |> 
  _[,lapply(.SD,sum),.(muni_id,muni_name,mpo,cmsbt08),.SDcols=(cs)] |> 
  mutate(
    hh2020_per_acre_scen1 = scen1_total_households_2020 / developable_acres,
    hh2020_per_acre_scen2 = scen2_total_households_2020 / developable_acres,
    hh2020_per_acre_scen3 = scen3_total_households_2020 / developable_acres,
    hh2020_per_acre_scen4 = scen4_total_households_2020 / developable_acres,
    hh2050_per_acre_scen1 = scen1_total_households_2050 / developable_acres,
    hh2050_per_acre_scen2 = scen2_total_households_2050 / developable_acres,
    hh2050_per_acre_scen3 = scen3_total_households_2050 / developable_acres,
    hh2050_per_acre_scen4 = scen4_total_households_2050 / developable_acres,
    chg_hh_per_acre_scen1 = hh2050_per_acre_scen1 - hh2020_per_acre_scen1,
    chg_hh_per_acre_scen2 = hh2050_per_acre_scen2 - hh2020_per_acre_scen2,
    chg_hh_per_acre_scen3 = hh2050_per_acre_scen3 - hh2020_per_acre_scen3,
    chg_hh_per_acre_scen4 = hh2050_per_acre_scen4 - hh2020_per_acre_scen4,
    unit2020_per_acre_scen1 = scen1_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen2 = scen2_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen3 = scen3_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen4 = scen4_total_residential_units_2020 / developable_acres,
    unit2050_per_acre_scen1 = scen1_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen2 = scen2_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen3 = scen3_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen4 = scen4_total_residential_units_2050 / developable_acres,
    chg_unit_per_acre_scen1 = unit2050_per_acre_scen1 - unit2020_per_acre_scen1,
    chg_unit_per_acre_scen2 = unit2050_per_acre_scen2 - unit2020_per_acre_scen2,
    chg_unit_per_acre_scen3 = unit2050_per_acre_scen3 - unit2020_per_acre_scen3,
    chg_unit_per_acre_scen4 = unit2050_per_acre_scen4 - unit2020_per_acre_scen4
  )

msf <- muni_sf(2020) |> 
  select(muni_id,geometry) |> 
  left_join(mun, by='muni_id')

st_write(msf, dsn = task4_gdb, layer = 'Task_4_Scenario_Comparisons_muni', append=F)

cmt.tab <- msf |> 
  st_drop_geometry() |> 
  setDT() |> 
  _[,lapply(.SD,sum),.(cmsbt08),.SDcols=(cs)] |> 
  mutate(
    hh2020_per_acre_scen1 = scen1_total_households_2020 / developable_acres,
    hh2020_per_acre_scen2 = scen2_total_households_2020 / developable_acres,
    hh2020_per_acre_scen3 = scen3_total_households_2020 / developable_acres,
    hh2020_per_acre_scen4 = scen4_total_households_2020 / developable_acres,
    hh2050_per_acre_scen1 = scen1_total_households_2050 / developable_acres,
    hh2050_per_acre_scen2 = scen2_total_households_2050 / developable_acres,
    hh2050_per_acre_scen3 = scen3_total_households_2050 / developable_acres,
    hh2050_per_acre_scen4 = scen4_total_households_2050 / developable_acres,
    chg_hh_per_acre_scen1 = hh2050_per_acre_scen1 - hh2020_per_acre_scen1,
    chg_hh_per_acre_scen2 = hh2050_per_acre_scen2 - hh2020_per_acre_scen2,
    chg_hh_per_acre_scen3 = hh2050_per_acre_scen3 - hh2020_per_acre_scen3,
    chg_hh_per_acre_scen4 = hh2050_per_acre_scen4 - hh2020_per_acre_scen4,
    unit2020_per_acre_scen1 = scen1_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen2 = scen2_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen3 = scen3_total_residential_units_2020 / developable_acres,
    unit2020_per_acre_scen4 = scen4_total_residential_units_2020 / developable_acres,
    unit2050_per_acre_scen1 = scen1_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen2 = scen2_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen3 = scen3_total_residential_units_2050 / developable_acres,
    unit2050_per_acre_scen4 = scen4_total_residential_units_2050 / developable_acres,
    chg_unit_per_acre_scen1 = unit2050_per_acre_scen1 - unit2020_per_acre_scen1,
    chg_unit_per_acre_scen2 = unit2050_per_acre_scen2 - unit2020_per_acre_scen2,
    chg_unit_per_acre_scen3 = unit2050_per_acre_scen3 - unit2020_per_acre_scen3,
    chg_unit_per_acre_scen4 = unit2050_per_acre_scen4 - unit2020_per_acre_scen4
  )

fwrite(cmt.tab, paste0(data_path,'/Task_4/Scenario_comparisons_community_subtype.csv'))

```


# All Code Below is Deprecated
#
#

# 5 - Arcpy erase undevelopable land

```{python}
import arcpy
import os
from arcpy import env

env.overwriteOutput = True
env.workspace = r.task3_gdb 

# Add block area
arcpy.management.CalculateGeometryAttributes(
    in_features="blocks_2010",
    geometry_property="block_area_m2 AREA",
    length_unit="",
    area_unit="SQUARE_METERS",
    coordinate_system=None,
    coordinate_format="SAME_AS_INPUT"
)

arcpy.analysis.PairwiseErase(
    in_features="blocks_2010",
    erase_features="main_state_wide_urbansim_undevelopable_land_constraint",
    out_feature_class="block_developable",
    cluster_tolerance=None
)

arcpy.management.AddField(
    in_table="block_developable",
    field_name="block_developable_area_m2",
    field_type="DOUBLE",
    field_precision=None,
    field_scale=None,
    field_length=None,
    field_alias="",
    field_is_nullable="NULLABLE",
    field_is_required="NON_REQUIRED",
    field_domain=""
)

arcpy.management.CalculateField(
    in_table="block_developable",
    field="block_developable_area_m2",
    expression="!SHAPE_Area!",
    expression_type="PYTHON3",
    code_block="",
    field_type="TEXT",
    enforce_domains="NO_ENFORCE_DOMAINS"
)

arcpy.conversion.ExportTable(
    in_table="block_developable",
    out_table=r.data_path + "/Task_3/Spatial/block_developable_area.csv",
    where_clause="",
    use_field_alias_as_name="NOT_USE_ALIAS"
)


```

# 6 - Calculate change in efficiency metrics

```{r}

dev <- fread(paste0(data_path,"/Task_3/Spatial/block_developable_area.csv")) |> 
  select(bl10_id, block_developable_area_m2) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  left_join(blk.comp, by='bl10_id')

cx <- blk_2010 |>
  select(bl10_id) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  left_join(dev, by='bl10_id') |> 
  left_join(empx, by='bl10_id') |> 
  setDT()
cx[is.na(cx)] <- 0

dx <- cx |> 
  mutate(bl10_id = substr(bl10_id,1,12),
         ct10_id = substr(bl10_id,1,11),
         dev_acres = block_developable_area_m2 * 0.000247105,
         unit_acre_2000 = total_units_2000 / dev_acres,
         unit_acre_2010 = total_units_2010 / dev_acres,
         unit_acre_2020 = total_units_2020 / dev_acres,
         pop_acre_2000 = total_pop_2000 / dev_acres,
         pop_acre_2010 = total_pop_2010 / dev_acres,
         pop_acre_2020 = total_pop_2020 / dev_acres,
         emp_acre_2011 = emp2011 / dev_acres,
         emp_acre_2021 = emp2021 / dev_acres,
         emp_acre_2023 = emp2023 / dev_acres,
         delta_unit_acre_00_10 = unit_acre_2010 - unit_acre_2000,
         delta_unit_acre_10_20 = unit_acre_2020 - unit_acre_2010,
         delta_unit_acre_00_20 = unit_acre_2020 - unit_acre_2000,
         delta_pop_acre_00_10 = pop_acre_2010 - pop_acre_2000,
         delta_pop_acre_10_20 = pop_acre_2020 - pop_acre_2010,
         delta_pop_acre_00_20 = pop_acre_2020 - pop_acre_2000,
         delta_emp_acre_11_21 = emp_acre_2021 - emp_acre_2011,
         delta_emp_acre_21_23 = emp_acre_2023 - emp_acre_2021,
         delta_emp_acre_11_23 = emp_acre_2023 - emp_acre_2011
         )

bgx <- dx |> 
  group_by(bl10_id) |> 
  summarize(total_units_2000 = sum(total_units_2000),
            total_units_2010 = sum(total_units_2010),
            total_units_2020 = sum(total_units_2020),
            total_pop_2000 = sum(total_pop_2000),
            total_pop_2010 = sum(total_pop_2010),
            total_pop_2020 = sum(total_pop_2020),
            emp2011 = sum(emp2011),
            emp2021 = sum(emp2021),
            emp2023 = sum(emp2023),
            dev_acres = sum(dev_acres)) |> 
  ungroup() |> 
  mutate(
         unit_acre_2000 = total_units_2000 / dev_acres,
         unit_acre_2010 = total_units_2010 / dev_acres,
         unit_acre_2020 = total_units_2020 / dev_acres,
         pop_acre_2000 = total_pop_2000 / dev_acres,
         pop_acre_2010 = total_pop_2010 / dev_acres,
         pop_acre_2020 = total_pop_2020 / dev_acres,
         emp_acre_2011 = emp2011 / dev_acres,
         emp_acre_2021 = emp2021 / dev_acres,
         emp_acre_2023 = emp2023 / dev_acres,
         delta_unit_acre_00_10 = unit_acre_2010 - unit_acre_2000,
         delta_unit_acre_10_20 = unit_acre_2020 - unit_acre_2010,
         delta_unit_acre_00_20 = unit_acre_2020 - unit_acre_2000,
         delta_pop_acre_00_10 = pop_acre_2010 - pop_acre_2000,
         delta_pop_acre_10_20 = pop_acre_2020 - pop_acre_2010,
         delta_pop_acre_00_20 = pop_acre_2020 - pop_acre_2000,
         delta_emp_acre_11_21 = emp_acre_2021 - emp_acre_2011,
         delta_emp_acre_21_23 = emp_acre_2023 - emp_acre_2021,
         delta_emp_acre_11_23 = emp_acre_2023 - emp_acre_2011
         )

munix <- dx |> 
  group_by(muni_id,muni_name) |> 
  summarize(total_units_2000 = sum(total_units_2000),
            total_units_2010 = sum(total_units_2010),
            total_units_2020 = sum(total_units_2020),
            total_pop_2000 = sum(total_pop_2000),
            total_pop_2010 = sum(total_pop_2010),
            total_pop_2020 = sum(total_pop_2020),
            emp2011 = sum(emp2011),
            emp2021 = sum(emp2021),
            emp2023 = sum(emp2023),
            dev_acres = sum(dev_acres)) |> 
  ungroup() |> 
  mutate(
         unit_acre_2000 = total_units_2000 / dev_acres,
         unit_acre_2010 = total_units_2010 / dev_acres,
         unit_acre_2020 = total_units_2020 / dev_acres,
         pop_acre_2000 = total_pop_2000 / dev_acres,
         pop_acre_2010 = total_pop_2010 / dev_acres,
         pop_acre_2020 = total_pop_2020 / dev_acres,
         emp_acre_2011 = emp2011 / dev_acres,
         emp_acre_2021 = emp2021 / dev_acres,
         emp_acre_2023 = emp2023 / dev_acres,
         delta_unit_acre_00_10 = unit_acre_2010 - unit_acre_2000,
         delta_unit_acre_10_20 = unit_acre_2020 - unit_acre_2010,
         delta_unit_acre_00_20 = unit_acre_2020 - unit_acre_2000,
         delta_pop_acre_00_10 = pop_acre_2010 - pop_acre_2000,
         delta_pop_acre_10_20 = pop_acre_2020 - pop_acre_2010,
         delta_pop_acre_00_20 = pop_acre_2020 - pop_acre_2000,
         delta_emp_acre_11_21 = emp_acre_2021 - emp_acre_2011,
         delta_emp_acre_21_23 = emp_acre_2023 - emp_acre_2021,
         delta_emp_acre_11_23 = emp_acre_2023 - emp_acre_2011
  )

cmtyp <- all_muni_data_keys |> select(muni_id,cmtyp08)

cmtypx <- dx |> 
  left_join(cmtyp, by = 'muni_id') |> 
  group_by(cmtyp08) |> 
  summarize(total_units_2000 = sum(total_units_2000),
            total_units_2010 = sum(total_units_2010),
            total_units_2020 = sum(total_units_2020),
            total_pop_2000 = sum(total_pop_2000),
            total_pop_2010 = sum(total_pop_2010),
            total_pop_2020 = sum(total_pop_2020),
            emp2011 = sum(emp2011),
            emp2021 = sum(emp2021),
            emp2023 = sum(emp2023),
            dev_acres = sum(dev_acres)) |> 
  ungroup() |> 
  mutate(
         unit_acre_2000 = total_units_2000 / dev_acres,
         unit_acre_2010 = total_units_2010 / dev_acres,
         unit_acre_2020 = total_units_2020 / dev_acres,
         pop_acre_2000 = total_pop_2000 / dev_acres,
         pop_acre_2010 = total_pop_2010 / dev_acres,
         pop_acre_2020 = total_pop_2020 / dev_acres,
         emp_acre_2011 = emp2011 / dev_acres,
         emp_acre_2021 = emp2021 / dev_acres,
         emp_acre_2023 = emp2023 / dev_acres,
         delta_unit_acre_00_10 = unit_acre_2010 - unit_acre_2000,
         delta_unit_acre_10_20 = unit_acre_2020 - unit_acre_2010,
         delta_unit_acre_00_20 = unit_acre_2020 - unit_acre_2000,
         delta_pop_acre_00_10 = pop_acre_2010 - pop_acre_2000,
         delta_pop_acre_10_20 = pop_acre_2020 - pop_acre_2010,
         delta_pop_acre_00_20 = pop_acre_2020 - pop_acre_2000,
         delta_emp_acre_11_21 = emp_acre_2021 - emp_acre_2011,
         delta_emp_acre_21_23 = emp_acre_2023 - emp_acre_2021,
         delta_emp_acre_11_23 = emp_acre_2023 - emp_acre_2011
         )

rpax <- dx |> 
  group_by(rpa_acr) |> 
  summarize(total_units_2000 = sum(total_units_2000),
            total_units_2010 = sum(total_units_2010),
            total_units_2020 = sum(total_units_2020),
            total_pop_2000 = sum(total_pop_2000),
            total_pop_2010 = sum(total_pop_2010),
            total_pop_2020 = sum(total_pop_2020),
            emp2011 = sum(emp2011),
            emp2021 = sum(emp2021),
            emp2023 = sum(emp2023),
            dev_acres = sum(dev_acres)) |> 
  ungroup() |> 
  mutate(
         unit_acre_2000 = total_units_2000 / dev_acres,
         unit_acre_2010 = total_units_2010 / dev_acres,
         unit_acre_2020 = total_units_2020 / dev_acres,
         pop_acre_2000 = total_pop_2000 / dev_acres,
         pop_acre_2010 = total_pop_2010 / dev_acres,
         pop_acre_2020 = total_pop_2020 / dev_acres,
         emp_acre_2011 = emp2011 / dev_acres,
         emp_acre_2021 = emp2021 / dev_acres,
         emp_acre_2023 = emp2023 / dev_acres,
         delta_unit_acre_00_10 = unit_acre_2010 - unit_acre_2000,
         delta_unit_acre_10_20 = unit_acre_2020 - unit_acre_2010,
         delta_unit_acre_00_20 = unit_acre_2020 - unit_acre_2000,
         delta_pop_acre_00_10 = pop_acre_2010 - pop_acre_2000,
         delta_pop_acre_10_20 = pop_acre_2020 - pop_acre_2010,
         delta_pop_acre_00_20 = pop_acre_2020 - pop_acre_2000,
         delta_emp_acre_11_21 = emp_acre_2021 - emp_acre_2011,
         delta_emp_acre_21_23 = emp_acre_2023 - emp_acre_2021,
         delta_emp_acre_11_23 = emp_acre_2023 - emp_acre_2011
         )

swx <- dx |> 
  summarize(total_units_2000 = sum(total_units_2000),
            total_units_2010 = sum(total_units_2010),
            total_units_2020 = sum(total_units_2020),
            total_pop_2000 = sum(total_pop_2000),
            total_pop_2010 = sum(total_pop_2010),
            total_pop_2020 = sum(total_pop_2020),
            emp2011 = sum(emp2011),
            emp2021 = sum(emp2021),
            emp2023 = sum(emp2023),
            dev_acres = sum(dev_acres)) |> 
  mutate(
         unit_acre_2000 = total_units_2000 / dev_acres,
         unit_acre_2010 = total_units_2010 / dev_acres,
         unit_acre_2020 = total_units_2020 / dev_acres,
         pop_acre_2000 = total_pop_2000 / dev_acres,
         pop_acre_2010 = total_pop_2010 / dev_acres,
         pop_acre_2020 = total_pop_2020 / dev_acres,
         emp_acre_2011 = emp2011 / dev_acres,
         emp_acre_2021 = emp2021 / dev_acres,
         emp_acre_2023 = emp2023 / dev_acres,
         delta_unit_acre_00_10 = unit_acre_2010 - unit_acre_2000,
         delta_unit_acre_10_20 = unit_acre_2020 - unit_acre_2010,
         delta_unit_acre_00_20 = unit_acre_2020 - unit_acre_2000,
         delta_pop_acre_00_10 = pop_acre_2010 - pop_acre_2000,
         delta_pop_acre_10_20 = pop_acre_2020 - pop_acre_2010,
         delta_pop_acre_00_20 = pop_acre_2020 - pop_acre_2000,
         delta_emp_acre_11_21 = emp_acre_2021 - emp_acre_2011,
         delta_emp_acre_21_23 = emp_acre_2023 - emp_acre_2021,
         delta_emp_acre_11_23 = emp_acre_2023 - emp_acre_2011
         )

# Export to Excel Workbook
# Prep output excel workbook
wrapStyle <- createStyle(wrapText = TRUE)
boldStyle <- createStyle(textDecoration = "Bold")
eff <- createWorkbook()

addWorksheet(eff, sheetName = 'Blockgroup')
writeData(eff, sheet = 'Blockgroup', x = bgx)
addWorksheet(eff, sheetName = 'Municipality')
writeData(eff, sheet = 'Municipality', x = munix)
addWorksheet(eff, sheetName = 'Community Type')
writeData(eff, sheet = 'Community Type', x = cmtypx)
addWorksheet(eff, sheetName = 'RPA')
writeData(eff, sheet = 'RPA', x = rpax)
addWorksheet(eff, sheetName = 'Statewide')
writeData(eff, sheet = 'Statewide', x = swx)

# Styles for sheet
setColWidths(
  eff,
  cols = 1:ncol(bgx),
  sheet = 'Blockgroup',
  widths = "auto"
)
setColWidths(
  eff,
  cols = 1:ncol(munix),
  sheet = 'Municipality',
  widths = "auto"
)
setColWidths(
  eff,
  cols = 1:ncol(cmtypx),
  sheet = 'Community Type',
  widths = "auto"
)
setColWidths(
  eff,
  cols = 1:ncol(rpax),
  sheet = 'RPA',
  widths = "auto"
)
setColWidths(
  eff,
  cols = 1:ncol(swx),
  sheet = 'Statewide',
  widths = "auto"
)

saveWorkbook(eff, paste0(data_path,'/Task_3/Deliverables/Task_3_Census_Land_Use_Efficiency_Metrics.xlsx'), overwrite = T)

```



# Existing Capacity Constraints 2019
```{r}
# Base year existing housing unit capacity levels

for (yr in c(2019)) {
  dt.mapc <- fread(paste0(
    data_path,
    '/UrbanSim/Current/MAPC/results_mapc_',
    MAPCM_run,
    '_block_indicators_',yr,'.csv'
  )) |> select(
    block_id,
    residential_unit_capacity,
    total_residential_units,
    vacant_residential_units,
    ru_vacancy_rate,
    total_households,
    density_households,
    unavailable_units,
    acres
  ) |>
    mutate(
      year = yr,
      bl10_id = as.character(block_id),
      residential_unit_capacity = round(residential_unit_capacity),
      available_unit_capacity = residential_unit_capacity - total_residential_units,
      cap_pct = round(100 * total_residential_units / residential_unit_capacity, 2)
    ) |>
    mutate(cap_pct = case_when(is.infinite(cap_pct) ~ 100, 
                               is.na(cap_pct) ~ 0,
                               .default = cap_pct)) |>
    left_join(blk.sf, by = 'bl10_id')

  # Statewide Model Outputs
  
    dt.swm <- fread(paste0(
    data_path,
    '/UrbanSim/Current/SWM/results_mapc_',
    SWM_run,
    '_block_indicators_',yr,'.csv'
  )) |> select(
    block_id,
    residential_unit_capacity,
    total_residential_units,
    vacant_residential_units,
    ru_vacancy_rate,
    total_households,
    density_households,
    unavailable_units,
    acres
  ) |>
    mutate(
      year = yr,
      bl10_id = as.character(block_id),
      residential_unit_capacity = round(residential_unit_capacity),
      available_unit_capacity = residential_unit_capacity - total_residential_units,
      cap_pct = round(100 * total_residential_units / residential_unit_capacity, 2)
    ) |>
    mutate(cap_pct = case_when(is.infinite(cap_pct) ~ 100, 
                               is.na(cap_pct) ~ 0,
                               .default = cap_pct)) |>
    left_join(blk.sf, by = 'bl10_id') |> 
    filter(rpa_acr!='MAPC')
    
  dt.comb <- rbind(dt.mapc,dt.swm)
  st_write(dt.comb, 
           dsn = project_gdb, 
           layer = paste0('urbansim_current_block_developed_capacities_',yr))
}

# Export block / block group / tract geographies for Harvard Forest
blk <- st_read(dsn=scratch, layer='Urbansim_households_by_block')
bg <- st_read(dsn=scratch, layer='Urbansim_households_by_blockgroup')
tract <- st_read(dsn=scratch, layer='Urbansim_households_by_tract')

st_write(blk, dsn = urbsim_gpkg, layer ='urbansim_households_by_block')
st_write(bg, dsn = urbsim_gpkg, layer ='urbansim_households_by_blockgroup')
st_write(tract, dsn = urbsim_gpkg, layer ='urbansim_households_by_tract')

st_write(blk, dsn = project_gdb, layer ='urbansim_households_by_block')
st_write(bg, dsn = project_gdb, layer ='urbansim_households_by_blockgroup')
st_write(tract, dsn = project_gdb, layer ='urbansim_households_by_tract')

```

# 4 - Excel format template for UrbanSim block-level outputs

```{r}
# Prep output excel workbook
wrapStyle <- createStyle(wrapText = TRUE)
boldStyle <- createStyle(textDecoration = "Bold")
blk.hh <- createWorkbook()

for (yr in years) {
    mp <- fread(
      paste0(
             data_path,
             '/UrbanSim/',
             scen,
             '/',
             scen,
             '_MAPCM_block_indicators_',
             yr + 1,
             '.csv'
           ))
    sw <- fread(
      paste0(
             data_path,'/UrbanSim/',
             '/',
             scen,
             '/',
             scen,
             '_SWM_block_indicators_',
             yr + 1,
             '.csv'
           ))
    dt <- rbind(mp,sw)
    
addWorksheet(blk.hh, sheetName = paste(scen, yr+1, sep = '_'))
writeData(blk.hh, sheet = paste(scen, yr+1, sep = '_'), x = dt)

}

# Styles for sheet
setColWidths(
  blk.hh,
  cols = 1:ncol(dt),
  sheet = paste(scen, yr + 1, sep = '_'),
  widths = "auto"
)
addStyle(
  blk.hh,
  sheet = paste(scen, yr + 1, sep = '_'),
  style = boldStyle,
  cols = 1:ncol(dt),
  rows = 1,
  gridExpand = TRUE
)

# Employment Sector Aggregation Metadata
meta <- data.table(
  employment_aggregation = seq(1, 10),
  sector_description = c(
    'Construction',
    'Education and Health Services',
    'Financial Activities',
    'Government',
    'Information',
    'Leisure and Hospitality',
    'Manufacturing',
    'Mining and Logging / Other Services',
    'Professional and Business Services',
    'Trade, Transportation, Utilities'
  )
)
# 
addWorksheet(blk.hh, sheetName = 'employment sector metadata')
writeData(blk.hh, sheet = 'employment sector metadata', x = meta)

# Styles for sheet
setColWidths(
  blk.hh,
  cols=1:ncol(meta),
  sheet = 'employment sector metadata',
  widths = "auto"
)
addStyle(
  blk.hh,
  sheet = 'employment sector metadata',
  style = boldStyle,
  cols=1:ncol(meta),
  rows = 1,
  gridExpand = TRUE
)

saveWorkbook(blk.hh, paste0(data_path,'/UrbanSim/',scen,'/',scen, '_block_indicators.xlsx'), overwrite = T)


```
# 4 - Census-based Land Use Efficiency Metrics
```{r}


var00 <- load_variables(2000,'sf1') |> setDT()
var00[grepl('GROUP QUARTERS',concept),sun(concept)]
var00[grepl('UNITS',concept),sun(concept)]
var00[grepl('POPULATION',concept),sun(concept)]

vars.pull <- var00[concept %in% c('HOUSING UNITS [1]', 'TOTAL POPULATION IN OCCUPIED HOUSING UNITS[1]') | (concept == 'GROUP QUARTERS POPULATION BY GROUP QUARTERS TYPE [9]' & label =='Total!!Noninstitutionalized population!!College dormitories (includes college quarters off campus)')] |> pull(name)

# Use NHGIS crosswalk to convert counts in 2000 vintage geographies to 2010 census blocks

xw00 <- fread(paste0(data_path,'/nhgis_blk2000_blk2010_25.csv')) |> 
  select(blk2000ge, blk2010ge, weight) |> 
  rename(GEOID = blk2000ge,
         block2010_geoid = blk2010ge) |> 
  mutate(GEOID = as.character(GEOID))

d00 <- get_decennial(
  state='MA',
  geography='block',
  year=2000,
  variables=vars.pull,
  output='wide'
) |> 
  rename(units2000 = vars.pull[1], 
         pop2000 = vars.pull[2],
         dorm_pop2000 = vars.pull[3]) |> 
  left_join(xw00, by='GEOID') |> 
  mutate(units00_10 = units2000 * weight,
         pop00_10 = pop2000 * weight,
         dorm_pop00_10 = dorm_pop2000 * weight) |> 
  group_by(block2010_geoid) |> 
  summarize(total_units_2000 = round(sum(units00_10)),
            total_pop_2000 = round(sum(pop00_10)),
            total_dorm_pop_2000 = round(sum(dorm_pop00_10))
            ) |> 
  mutate(block2010_geoid = as.character(block2010_geoid)) |> 
  setDT()

# 2010 Housing Units and Population

var10 <- load_variables(2010,'sf1') |> setDT()
var10[grepl('GROUP QUARTERS',concept),sun(concept)]
var10[grepl('UNITS',concept),sun(concept)]
var10[grepl('POPULATION',concept),sun(concept)]

vars.pull <- var10[concept %in% c('HOUSING UNITS', 'TOTAL POPULATION IN OCCUPIED HOUSING UNITS') | (concept== 'GROUP QUARTERS POPULATION IN COLLEGE/UNIVERSITY STUDENT HOUSING BY SEX BY AGE' & label == 'Total (501)')] |> pull(name)

cntys <- mapcdatakeys::geog_xw_2010 |> pull(countyfp) |> unique()

d10 <- data.table()

for (i in cntys){
dtmp <- get_decennial(
  state='MA',
  geography='block',
  county=i,
  year=2010,
  variables=vars.pull,
  output='wide'
) |> 
  rename(total_units_2010 = vars.pull[1], 
         total_pop_2010 = vars.pull[2],
         total_dorm_pop_2010 = vars.pull[3]) |> 
  mutate(block2010_geoid = as.character(GEOID)) |> 
  setDT()
d10 <- rbind(d10,dtmp)
}

d10 <- d10 |> select(-GEOID,-NAME)

# 2020 housing units and population data

var20 <- load_variables(2020,'dhc') |> setDT()
var20[grepl('GROUP QUARTERS',concept),sun(concept)]
var20[grepl('UNITS',concept),sun(concept)]
var20[grepl('POPULATION',concept),sun(concept)]

vars.pull <- var20[concept %in% c('HOUSING UNITS', 'TOTAL POPULATION IN OCCUPIED HOUSING UNITS') | name=='PCO8_001N'] |> pull(name)
print(var20[name %in% vars.pull,.(name,label,concept)])

xw20 <- fread(paste0(data_path,'/nhgis_blk2020_blk2010_25.csv')) |> 
  select(blk2020ge, blk2010ge, weight) |> 
  rename(GEOID = blk2020ge,
         block2010_geoid = blk2010ge) |> 
  mutate(GEOID = as.character(GEOID)) |> 
  setDT()

d20 <- get_decennial(
  state='MA',
  geography='block',
  year=2020,
  sumfile = 'dhc',
  variables=vars.pull,
  output='wide'
) |> 
  rename(units2020 = vars.pull[1], 
         pop2020 = vars.pull[2],
         dorm_pop2020 = vars.pull[3]) |> 
  left_join(xw20, by='GEOID') |> 
  mutate(units20_10 = units2020 * weight,
         pop20_10 = pop2020 * weight,
         dorm_pop20_10 = dorm_pop2020 * weight) |> 
  group_by(block2010_geoid) |> 
  summarize(total_units_2020 = round(sum(units20_10)),
            total_pop_2020 = round(sum(pop20_10)),
            total_dorm_pop_2020 = round(sum(dorm_pop20_10))
            ) |> 
  mutate(block2010_geoid = as.character(block2010_geoid)) |> 
  setDT()

comp <- d00 |> 
  left_join(d10, by = 'block2010_geoid') |> 
  left_join(d20, by = 'block2010_geoid') |> 
  select(-total_dorm_pop_2000,-total_dorm_pop_2010,-total_dorm_pop_2020)

blk.comp <- left_join(blk10_sf, comp, by=c('bl10_id' = 'block2010_geoid'))
st_write(blk.comp, 
         dsn = paste0(data_path, '/Task_3/Spatial/LandUse_Efficiency_Metrics.gdb'), 
         layer = 'block_pop_units_2000_2010_2020',
         overwrite = T)

st_write(blk10_sf |> mutate(block_area_m2 = 0),
         dsn = paste0(data_path, '/Task_3/Spatial/LandUse_Efficiency_Metrics.gdb'), 
         layer = 'blocks_2010',
         append=F)
# Employment

emp11 <- lehdr::grab_lodes(state='ma',year=2011,lodes_type='wac') |>
  select(w_geocode,C000) |>
  setnames(c('GEOID','emp2011')) |> 
  left_join(xw20, by='GEOID') |> 
  mutate(emp2011 = emp2011*weight) |> 
  group_by(block2010_geoid) |> 
  summarize(emp2011 = round(sum(emp2011))) |> 
  ungroup() |> 
  rename(bl10_id = block2010_geoid) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  setDT()

emp21 <- lehdr::grab_lodes(state='ma',year=2021,lodes_type='wac') |>
  select(w_geocode,C000) |>
  setnames(c('GEOID','emp2021')) |> 
  left_join(xw20, by='GEOID') |> 
  mutate(emp2021 = emp2021*weight) |> 
  group_by(block2010_geoid) |> 
  summarize(emp2021 = round(sum(emp2021))) |> 
  ungroup() |> 
  rename(bl10_id = block2010_geoid) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  setDT()

emp23 <- lehdr::grab_lodes(state='ma',year=2023,lodes_type='wac') |>
  select(w_geocode,C000) |>
  setnames(c('GEOID','emp2023')) |>
  left_join(xw20, by='GEOID') |> 
  mutate(emp2023 = emp2023*weight) |> 
  group_by(block2010_geoid) |> 
  summarize(emp2023 = round(sum(emp2023))) |> 
  ungroup() |> 
  rename(bl10_id = block2010_geoid) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  setDT()

empx <- blk_2010 |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  left_join(emp11, by = 'bl10_id') |> 
  left_join(emp21, by='bl10_id') |> 
  left_join(emp23, by='bl10_id') |> 
  setDT()
empx[is.na(empx)] <- 0


```
# Deprecated
```{r}

# Add 2000 and 2010 tract level households and employment
var00 <- load_variables(2000,'sf1') |> setDT()
var00[grepl('HOUSEHOLDS',concept),sun(concept)]

xw <- fread(paste0(data_path,'/nhgis_blk2000_blk2010_25.csv')) |> 
  select(blk2000ge, blk2010ge, weight) |> 
  rename(GEOID = blk2000ge,
         block2010_geoid = blk2010ge) |> 
  mutate(GEOID = as.character(GEOID))

m3 <- get_decennial(
  state='MA',
  geography='block',
  year=2000,
  variables='P015001'
) |> 
  rename(households2000 = value) |> 
  left_join(xw, by='GEOID') |> 
  mutate(hh00_10 = households2000 * weight) |> 
  group_by(block2010_geoid) |> 
  summarize(total_households_2000 = round(sum(hh00_10))) |> 
  mutate(block2010_geoid= as.character(block2010_geoid)) |> 
  right_join(m2, by='block2010_geoid') |> 
  relocate(total_households_2000, .after=mpo) |> 
  setDT()

blk.hh <- st_as_sf(m3,crs=26986)
st_write(m3, dsn = urbsim_gpkg, layer ='urbansim_households_and_employment_by_2010block_2000_2035', append=F)

# Summarize to block group and tract level households and jobs
cols <- names(m3)[7:16]

blkgrp <- m3[,lapply(.SD,sum),blockgroup2010_geoid,.SDcols=(cols)]
blkgrp.sf <- blockgroup_sf(2010) |> 
  left_join(blkgrp, by=c('bl10_id'='blockgroup2010_geoid')) |> 
  setorder(bl10_id) |> 
  rename(blockgroup2010_geoid = bl10_id)

st_write(blkgrp.sf, dsn = urbsim_gpkg, layer ='urbansim_households_and_employment_by_2010blockgroup_2000_2035', append=F)

trct <- m3[,lapply(.SD,sum),tract2010_geoid,.SDcols=(cols)]
trct.sf <- tract_sf(2010) |> 
  left_join(trct, by=c('ct10_id'='tract2010_geoid')) |> 
  setorder(ct10_id) |> 
  rename(tract2010_geoid = ct10_id)

            
st_write(trct.sf, dsn = urbsim_gpkg, layer ='urbansim_households_and_employment_by_2010tract_2000_2035', append=F)

# Add jobs by supersector

mpc <- copy(blk_2010)

for (yr in years) {
  mp <- fread(
    paste0(
      data_path,
      '/UrbanSim/',
      scen,
      '/',
      scen,
      '_MAPCM_block_indicators_',
      yr + 1,
      '.csv'
    )
  ) |>
    mutate(bl10_id = as.numeric(block_id_2010)) |> 
    select(
      bl10_id,
      total_jobs,
      jobs_aggr_sector_1,
      jobs_aggr_sector_2,
      jobs_aggr_sector_3,
      jobs_aggr_sector_4,
      jobs_aggr_sector_5,
      jobs_aggr_sector_6,
      jobs_aggr_sector_7,
      jobs_aggr_sector_8,
      jobs_aggr_sector_9,
      jobs_aggr_sector_10
    ) |>
    setDT()
   if(yr == 2010){
    setnames(mp, c('bl10_id', paste0(names(mp)[2:ncol(mp)], '_', yr)))
  } else {
    setnames(mp, c('bl10_id', paste0(names(mp)[2:ncol(mp)], '_', yr+1)))
  }
  # SWM
  sp <- fread(
    paste0(
      data_path,
      '/UrbanSim/',
      scen,
      '/',
      scen,
      '_SWM_block_indicators_',
      yr + 1,
      '.csv'
    )
  ) |>
    mutate(bl10_id = as.numeric(block_id_2010)) |> 
    filter(!bl10_id %in% mapc_blks) |> 
    select(
      bl10_id,
      total_jobs,
      jobs_aggr_sector_1,
      jobs_aggr_sector_2,
      jobs_aggr_sector_3,
      jobs_aggr_sector_4,
      jobs_aggr_sector_5,
      jobs_aggr_sector_6,
      jobs_aggr_sector_7,
      jobs_aggr_sector_8,
      jobs_aggr_sector_9,
      jobs_aggr_sector_10
    ) |>
    setDT()
  if(yr == 2010){
    setnames(sp, c('bl10_id', paste0(names(sp)[2:ncol(sp)], '_', yr)))
  } else {
    setnames(sp, c('bl10_id', paste0(names(sp)[2:ncol(sp)], '_', yr+1)))
  }
  ds <- rbind(mp,sp)
  
  mpc <- left_join(mpc, ds, by = 'bl10_id')
}
mpc[,blockgroup2010_geoid:=substr(bl10_id,1,12)]
mpc[,tract2010_geoid:=substr(bl10_id,1,11)]

# Summarize to block group and tract level households and jobs
cs <- names(mpc)[6:60]

bg <- mpc[,lapply(.SD,sum),blockgroup2010_geoid,.SDcols=(cs)]
bg.sf <- blockgroup_sf(2010) |> 
  left_join(bg, by=c('bl10_id'='blockgroup2010_geoid')) |> 
  setorder(bl10_id) |> 
  rename(blockgroup2010_geoid = bl10_id)

st_write(bg.sf, dsn = urbsim_gpkg, layer ='urbansim_employment_by_supersector_2010blockgroup_2000_2035', append=F)

ct <- mpc[,lapply(.SD,sum),tract2010_geoid,.SDcols=(cs)]
ct.sf <- tract_sf(2010) |> 
  left_join(ct, by=c('ct10_id'='tract2010_geoid')) |> 
  setorder(ct10_id) |> 
  rename(tract2010_geoid = ct10_id)

            
st_write(ct.sf, dsn = urbsim_gpkg, layer ='urbansim_employment_by_supersector_2010tract_2000_2035', append=F)
```
